### 2012-10-24 inst/examples/compare-cholsky.Rmd
Examples comparing cholsky method, and simulated stock dynamics
Something seems to be going wrong with the Cholsky method.  Results
match fine over certain function spaces but poorly on others.  Funny,
since this is supposed to be numerically *more* stable.

Examples implemented that infer stock dynamics from a stochastic
Beverton Holt and stochastic May (1979) bistable model.  Results
look reasonable, though still need to utilize tuning for the
hyperparameters.

Also added basic package documentation and updated the gp_fit.R
function.  Function currently supports both cholksy and direct metho

`ro comment=NA, message=FALSE, tidy=FALSE or`

## Compare Cholsky and Direct methods

```{r}
gp_fit_direct <- function(obs, X, pars=c(sigma_n=1, l=1)){

  sigma_n <- pars["sigma_n"]
  l <- pars["l"]

  ## Parameterization-specific
  SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
  cov <- function(X, Y) outer(X, Y, SE, l)

  n <- length(obs$x)
  K <- cov(obs$x, obs$x)
  I <-  diag(1, n)

  ## Direct method 
  cov_xx_inv <- solve(K + sigma_n ^ 2 * I)
  Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
  Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
  llik <- -.5 * t(obs$y) %*% cov_xx_inv %*% obs$y - 0.5 * log(det(cov_xx_inv)) - n * log(2 * pi) / 2

  out <- list(mu = Ef, var = Cf, loglik=llik)
  class(out) = "gp"
  out
}
```

```{r}
gp_fit_cholsky <- function(obs, X, pars=c(sigma_n=1, l=1)){

  sigma_n <- pars["sigma_n"]
  l <- pars["l"]

  ## Parameterization-specific
  SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
  cov <- function(X, Y) outer(X, Y, SE, l)

  n <- length(obs$x)
  K <- cov(obs$x, obs$x)
  I <-  diag(1, n)

  ## Cholksy method -- Consider implementing in C
  L <- chol(K + sigma_n ^ 2 * I)
  alpha <- solve(t(L), solve(L, obs$y))
  loglik <- -.5 * t(obs$y) %*% alpha - sum(log(diag(L))) - n * log(2 * pi) / 2
  k_star <- cov(obs$x, X)
  mu <- t(k_star) %*% alpha
  v <- solve(L, k_star)
  var <- diag( cov(X,X) - t(v) %*% v )

  ## Pt by pt following Rassmussen & WIlliams, but there's no reason to do it this way
#  tmp <- lapply(X, function(x_i){  
#    k_star <- cov(obs$x, x_i)
#    Y <- t(k_star) %*% alpha
#    v <- solve(L, k_star)
#    Var <- cov(x_i,x_i) - t(v) %*% v
#    list(Y=Y, Var=Var)
#  })
#  mu <- sapply(tmp,`[[`, "Y")
#  var <- sapply(tmp,`[[`, "Var")

  out <- list(mu = mu, var = var, loglik=loglik)
  class(out) = "gp"
  out
}
```

Run the comparison

```{r}
X <- seq(-5,5,len=50)
obs <- data.frame(x = c(-4, -3, -1,  0,  2),
                  y = c(-2,  0,  1,  2, -1))
system.time(gp <- gp_fit_cholsky(obs, X, c(sigma_n=0.8, l=1)))
system.time(gp_d <- gp_fit_direct(obs, X, c(sigma_n=0.8, l=1)))
```

Plot the results

```{r}
require(reshape2)
df <- data.frame(x=X, y=gp$mu, ymin=(gp$mu-2*sqrt(gp$var)), ymax=(gp$mu+2*sqrt(gp$var)))
df2 <- data.frame(x=X, y=gp_d$mu, ymin=(gp_d$mu-2*sqrt(diag(gp_d$var))), ymax=(gp_d$mu+2*sqrt(diag(gp_d$var))))
df2 <- melt(df2, id="x")
require(ggplot2)
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") + 
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) + 
  geom_line(data=df2, aes(x,value,color=variable))
```

Looks fine for this data, but not for Reed example:

```{r libraries}
require(pdgControl)
require(ggplot2)
f <- BevHolt
p <- c(1.2,.02)
K <- (p[1]-1)/p[2] 
z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
x_grid = seq(0, 1.5 * K, length=100)
T <- 40
sigma_g <- 0.01
x <- numeric(T)
x[1] <- 1
for(t in 1:(T-1))
  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
plot(x)
```

Predict the function over the target grid

```{r}
obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
X <- x_grid
system.time(gp <- gp_fit_cholsky(obs, X, c(sigma_n=0.8, l=1)))
system.time(gp_d <- gp_fit_direct(obs, X, c(sigma_n=0.8, l=1)))
```

Plot the results

```{r}
require(reshape2)
df <- data.frame(x=X, y=gp$mu, ymin=(gp$mu-2*sqrt(gp$var)), ymax=(gp$mu+2*sqrt(gp$var)))
df2 <- data.frame(x=X, y=gp_d$mu, ymin=(gp_d$mu-2*sqrt(diag(gp_d$var))), ymax=(gp_d$mu+2*sqrt(diag(gp_d$var))))
df2 <- melt(df2, id="x")
require(ggplot2)
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") + 
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) + 
  geom_line(data=df2, aes(x,value,color=variable))
```