### 2012-10-25 inst/examples/testing-gp-sdp.Rmd
testing out gp sdp, not sure why I do not valid policy functions

Quick trial SDP approaches with GP function
========================================================

```{r libraries}
require(pdgControl)
require(ggplot2)
opts_knit$set(upload.fun = socialR::flickr.url)
```

### Beverton-Holt function

Simulate some training data under a stochastic growth function with standard parameterization,

``` {r stateeqn}
f <- BevHolt
p <- c(1.5,.05)
K <- (p[1]-1)/p[2] 
```

Noise function 
```{r}
z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
```

Parameter definitions

```{r}
x_grid = seq(0, 1.5 * K, length=100)
T <- 40
sigma_g <- 0.1
x <- numeric(T)
x[1] <- 1
```

Simulation

```{r}
for(t in 1:(T-1))
  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
```

Predict the function over the target grid

```{r}
obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
X <- x_grid
library(nonparametricbayes)
gp <- gp_fit(obs, X, c(sigma_n=1, l=1))
```

Gaussian Process inference from this model.  True model shown in red.

```{r}
df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
true <- data.frame(x=X, y=sapply(X,f, 0, p))
require(ggplot2)
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
  geom_line(data=true, aes(x,y), col='red', lty=2)
```

## Stochastic Dynamic programming solution based on the posterior Gaussian process:

Define a transition matrix $F$ from the Gaussian process, giving the probability of going from state $x_t$ to $x_{t+1}$.
We already have the Gaussian process mean and variance predicted for each point $x$ on our grid, so this is simply:

```{r}
V <- sqrt(diag(gp$Cf))
F <- sapply(x_grid, function(x) dnorm(x, gp$Ef, V))
F <- t(apply(F,1,function(x) x/sum(x))) # normalize
h_grid <- x_grid
```

True $f(x)$

```{r}
mu <- sapply(x_grid, f, 0, p)
F_true <- sapply(x_grid, function(x) dnorm(x, mu, sigma_g))
F_true <- t(apply(F_true,1,function(x) x/sum(x))) # normalize
```

Somewhat silly way to adjust the matrix by harvest level:
```{r}
n <- length(h_grid)
per_harvest <- function(F,n)
  lapply(0:(n-1), function(i){
  d <- diag(1, nrow=(n-i))
  top <- matrix(0, nrow=i, ncol=n)
  top[,1] <- 1
  side <- matrix(0, nrow=(n-i), ncol=i)
  out <- F %*% rbind(top, cbind(d,side))
  t(apply(out,1,function(x) x/sum(x))) # normalize
})

```

True F is okay but not quite right:

```{r}
matrices <- per_harvest(F_true,n)
opt <- find_dp_optim(matrices, x_grid, h_grid, 20, 0, profit, delta=.01)
plot(opt$D[,1])
```

whoops: gp inferred F isn't working:

```{r}
matrices <- per_harvest(F,n)
opt <- find_dp_optim(matrices, x_grid, h_grid, 20, 0, profit, delta=.01)
plot(opt$D[,1])
```

Old-school calculation method:

```{r}
profit <- profit_harvest(price=1, c0 = 0, c1=0)
pdfn <- function(P, s) dlnorm(P, 0, s)
matrices <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g, pdfn)
opt <- find_dp_optim(matrices, x_grid=x_grid, h_grid=h_grid, OptTime=20, xT=0, profit=profit, delta=.05)
plot(opt$D[,1])
```