### 2012-10-25 inst/examples/may79-example.Rmd
Code for running the Reed example with May's bistable model instead of a BH model

GP Example using the May (1979) bistable model
==============================================

Use my custom libraries and the `ggplot` plotting tools

```{r libraries}
require(nonparametric-bayes)
require(pdgControl)
require(ggplot2)
opts_knit$set(upload.fun = socialR::flickr.url)
```

```{r}
f <- May
p <- c(r = .75, k = 10, a=1.3, H=1, Q = 3)
K <- 8 # approx
```

Model dynamics look like this:

```{r}
birth <- function(x) p["r"] * (1-  x / p["k"])
death <- function(x) p["a"] * x ^ (p["Q"] - 1) / (x ^ p["Q"] + p["H"])
df <- data.frame(x = x_grid, b = sapply(x_grid, birth), d = sapply(x_grid, death))
ggplot(df) + geom_line(aes(x, b), col = "blue") + geom_line(aes(x,d), col = "red")
````

Noise function, profit function
```{r}
z_g <- function(sigma_g) rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
profit <- profit_harvest(1,0,0)
```

Parameter definitions

```{r}
x_grid = seq(0, 1.5 * K, length=101)
T <- 40
sigma_g <- 0.1
x <- numeric(T)
x[1] <- 1
```

Simulation

```{r}
x[1] = 2.5
for(t in 1:(T-1))
  x[t+1] = z_g(sigma_g) * f(x[t], h=0, p=p)
plot(x)
```

Predict the function over the target grid

```{r}
obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
X <- x_grid
gp <- gp_fit(obs, X, c(sigma_n=1, l=1))
```

Gaussian Process inference from this model

```{r}
df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
true <- data.frame(x=X, y=sapply(X,f, 0, p))
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
  geom_line(data=true, aes(x,y), col='red', lty=2)
```