### 2012-10-26 inst/examples/testing-gp-sdp.Rmd
exploring hyperparameter estimation, etc

Quick trial SDP approaches with GP function
==============================================

```{r libraries}
require(pdgControl)
require(nonparametricbayes)
require(ggplot2)
opts_knit$set(upload.fun = socialR::flickr.url)
```

### Beverton-Holt function

Simulate some training data under a stochastic growth function with standard parameterization,

``` {r stateeqn}
f <- BevHolt
p <- c(1.5,.05)
K <- (p[1]-1)/p[2] 
```

Parameter definitions

```{r}
x_grid = seq(0, 1.5 * K, length=101)
T <- 40
sigma_g <- 0.1
x <- numeric(T)
x[1] <- 1
```

Noise function, profit function
```{r}
z_g <- function() rlnorm(1, 0, sigma_g) #1+(2*runif(1, 0,  1)-1)*sigma_g #
profit <- profit_harvest(1,0,0)
```

Simulation

```{r}
for(t in 1:(T-1))
  x[t+1] = z_g() * f(x[t], h=0, p=p)
```

Predict the function over the target grid

```{r}
obs <- data.frame(x=x[1:(T-1)],y=x[2:T])
X <- x_grid
library(nonparametricbayes)
gp <- gp_fit(obs, X, c(sigma_n=.5, l=1.5))
```

Gaussian Process inference from this model.  True model shown in red.

```{r}
df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
true <- data.frame(x=X, y=sapply(X,f, 0, p))
require(ggplot2)
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
  geom_line(data=true, aes(x,y), col='red', lty=2)
```

## Optimization over hyperparameters:

```{r}
minusloglik <- function(par){
  gp <- gp_fit(obs, X, par)
  -gp$llik
}
par <- c(sigma_n=1, l=1)
minusloglik(par)
o <- optim(par, minusloglik)
o
hyperpars <- o$par
```

**Yikes.**

let's try 1-D optimization:

```{r}
minusloglik <- function(par){
  gp <- gp_fit(obs, X, c(sigma_n=par, l=1))
  -gp$llik
}
minusloglik(1)
o <- optimize(minusloglik, c(0,150))
o
hyperpars <- c(sigma_n=o$minimum, l=1)
```

**Yikes, that shouldn't happen either**

Plot the optimal hyperparameter solution:

```{r}
gp <- gp_fit(obs, X, hyperpars)
df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
true <- data.frame(x=X, y=sapply(X,f, 0, p))
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
  geom_line(data=true, aes(x,y), col='red', lty=2)
```

### Misc ways to calc harvest matrices

Old-school calculation method:

```{r}
profit <- profit_harvest(price=1, c0 = 0, c1=0)
pdfn <- function(P, s) dlnorm(P, 0, s)
sdp_mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g, pdfn)
opt <- find_dp_optim(sdp_mat, x_grid=x_grid, h_grid=h_grid, OptTime=20, xT=0, profit=profit, delta=.05)
plot(opt$D[,1])
```

Somewhat silly way to adjust the matrix by harvest level:

```{r}
n <- length(h_grid)
silly <- function(F,n)
  lapply(0:(n-1), function(i){
  d <- diag(1, nrow=(n-i))
  top <- matrix(0, nrow=i, ncol=n)
  top[,1] <- 1
  side <- matrix(0, nrow=(n-i), ncol=i)
  out <- F %*% rbind(top, cbind(d,side))
  t(apply(out,1,function(x) x/sum(x))) # normalize
})
```

binning function, not quite right:

```{r}
bin <- function(x, grid_centers){ 
  match <- which.min(abs(grid_centers - x)) 
  out <- integer(length(grid_centers))
  out[match] <- 1
  out
  }
per_harvest <- function(F,n){
  lapply(h_grid, function(h){
    t(sapply(x_grid, function(x) bin(x - h, x_grid)))
  })
}
```