### 2012-11-09 inst/examples/hyperparameter-optimization.Rmd
hyperparameter optimization testing now as a separate file

## Optimization over hyperparameters:

```{r results="hide"}
knit("beverton_holt_data.Rmd")
```

```{r}
minusloglik <- function(par){
  gp <- gp_fit(obs, X, par)
  -gp$llik
}
par <- c(sigma_n=1, l=1)
minusloglik(par)
o <- optim(par, minusloglik)
o
hyperpars <- o$par
```

**Yikes.**

let's try 1-D optimization:

```{r}
minusloglik <- function(par){
  gp <- gp_fit(obs, X, c(sigma_n=par, l=1))
  -gp$llik
}
minusloglik(1)
o <- optimize(minusloglik, c(0,150))
o
hyperpars <- c(sigma_n=o$minimum, l=1)
```

**Yikes, that shouldn't happen either**

Plot the optimal hyperparameter solution:

```{r}
gp <- gp_fit(obs, X, hyperpars)
df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
true <- data.frame(x=X, y=sapply(X,f, 0, p))
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
  geom_line(data=true, aes(x,y), col='red', lty=2)
```