### 2012-11-16 inst/examples/hyperparameter-optimization.Rmd
Updated hyperparameter-optimization example
Shows that we can estimate l appropriately but fails to get sigma_n; instead estimating the upper boundary.

## Optimization over hyperparameters:

```{r results="hide"}
opts_knit$set(upload.fun = socialR::notebook.url)
```

Static definition of gp function

```{r}
gp_fit <- function(obs, X, pars=c(sigma_n=1, l=1)){
  sigma_n <- pars["sigma_n"]
  l <- pars["l"]

  SE <- function(Xi,Xj, l=l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
  cov <- function(X, Y) outer(X, Y, SE, l)

  n <- length(obs$x)
  K <- cov(obs$x, obs$x)
  I <-  diag(1, n)

    cov_xx_inv <- solve(K + sigma_n ^ 2 * I)
    Ef <- cov(X, obs$x) %*% cov_xx_inv %*% obs$y
    Cf <- cov(X, X) - cov(X, obs$x)  %*% cov_xx_inv %*% cov(obs$x, X)
    llik <- -.5 * t(obs$y) %*% cov_xx_inv %*% obs$y - 0.5 * log(det(cov_xx_inv)) - n * log(2 * pi) / 2

  out <- list(Ef = Ef, Cf = Cf, llik = llik, obs=obs, X=X, pars=pars, K=K, inv=cov_xx_inv)
  class(out) = "gpfit"
  out
}

plot.gpfit <- function(gp){
  dat <- data.frame(x=gp$X, y=(gp$Ef), ymin=(gp$Ef-2*sqrt(diag(gp$Cf))), ymax=(gp$Ef+2*sqrt(diag(gp$Cf))))
  ggplot(dat) +
  geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") +
  geom_line(aes(x=x,y=y), size=1) + 
  geom_point(data=gp$obs,aes(x=x,y=y)) + 
  labs(title=paste("llik =", prettyNum(gp$llik)))
}
```

Estimate a GP from data with given hyperparameters, then redraw data from an actual multivariate normal.  Presumably optimal hyperparameters might be near the given values then.

```{r}
require(MASS)
X <- seq(-5,5,len=20)
obs <- data.frame(x = c(-4, -3, -1,  0,  2, 3),
                  y = c(-2,  0,  1,  2, -1, -1))

fit <- gp_fit(obs, X, pars=c("sigma_n"=.3, "l"=.6))
gp_dat <- mvrnorm(1, fit$Ef, fit$Cf)
obs <- data.frame(x=X, y=gp_dat)
X <- seq(-5,5,len=50)
```

Example plot

```{r}
plot.gpfit(gp_fit(obs, X, pars=c("sigma_n"=.3, "l"=.6)))
```

estimates non-boundary solution of `l`:

```{r}
minusloglik <- function(par){
  gp <- gp_fit(obs, X, c(sigma_n=.3, l=par))
  -gp$llik
}
minusloglik(1)
o <- optimize(minusloglik, c(0,40))
o

```

But makes `sigma_n` arbitrarily large (boundary solution).

```{r}
minusloglik <- function(par){
  gp <- gp_fit(obs, X, c(sigma_n=par, l=.6))
  -gp$llik
}
minusloglik(1)
o <- optimize(minusloglik, c(0,40))
o

```

```{r}
gp_k <- gausspr(obs$x, obs$y, kernel="rbfdot")
gp_k
```

Convert to `l`, `sigma=1/(2l^2)`

```{r}
1/sqrt(2*as.numeric(gp_k@kernelf@kpar))
```

Weirdly inconsistent?

```{r}
gp_k <- gausspr(obs$x, obs$y, kernel="rbfdot")
1/sqrt(2*as.numeric(gp_k@kernelf@kpar))
gp_k <- gausspr(obs$x, obs$y, kernel="rbfdot")
1/sqrt(2*as.numeric(gp_k@kernelf@kpar))
```

### a second example, using the beverton holt data:

```{r results="hide"}
knit("beverton_holt_data.Rmd")
```

```{r}
minusloglik <- function(par){
  if(any(par < 0))
    Inf
  else {
  gp <- gp_fit(obs, X, par)
  -gp$llik
  }
}
par <- c(sigma_n=1, l=1)
minusloglik(par)
#o <- optim(par, minusloglik)
o <- optim(par, minusloglik, method="L", lower=c(0,0), upper=c(20,20))
o
hyperpars <- o$par
```

**Yikes.**

let's try 1-D optimization:

```{r}
minusloglik <- function(par){
  gp <- gp_fit(obs, X, c(sigma_n=par, l=5.7))
  -gp$llik
}
minusloglik(1)
o <- optimize(minusloglik, c(0,50))
o
hyperpars <- c(sigma_n=o$minimum, l=1)
```

**Yikes, that shouldn't happen either**

Plot the optimal hyperparameter solution:

```{r}
gp <- gp_fit(obs, X, hyperpars)
df <- data.frame(x=X, y=gp$Ef, ymin=(gp$Ef-2*sqrt(abs(diag(gp$Cf)))), ymax=(gp$Ef+2*sqrt(abs(diag(gp$Cf)))))
true <- data.frame(x=X, y=sapply(X,f, 0, p))
ggplot(df)  + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
  geom_line(aes(x,y)) + geom_point(data=obs, aes(x,y)) +
  geom_line(data=true, aes(x,y), col='red', lty=2)
```