### 2012-11-29 inst/examples/singular-inversion.Rmd
seeing if I get singular matrices in the inversion

testing singular matrix condition

```{r}
require(ggplot2)
require(kernlab)
require(nonparametricbayes)
opts_knit$set(upload.fun = socialR::notebook.url)
opts_chunk$set(dev.args=list(bg="transparent"), comment=NA, tidy=FALSE)
theme_set(theme_bw())
theme_update(panel.background = element_rect(fill = "transparent",colour = NA))
```

Parameterization-specific

```{r}
X <- seq(0,1, length=1000)
obs <- data.frame(x = X,
                  y = sin(2*pi*10*X)+.3*rnorm(length(X)))
```

### Direct method

``` {r}
d <- gp_fit(obs, X, c(sigma_n=1, l=.0001, tau=.3), "direct")
plot.gpfit(d)
```

``` {r}
d <- gp_fit(obs, X, c(sigma_n=1, l=10, tau=.3), "direct")
plot.gpfit(d)
```