### 2012-12-05 inst/examples/mcmc-gp.Rmd
Notes on `tgp` package MCMC

``` {r libs, include=FALSE}
rm(list=ls())   
require(nonparametricbayes)
require(reshape2)
require(ggplot2)
require(data.table)
require(tgp)
```

```{r graphing-options, include=FALSE}
opts_knit$set(upload.fun = socialR::notebook.url)
opts_chunk$set(dev.args=list(bg="transparent"), tidy=FALSE, comment=NA)
theme_notebook <- theme_grey() + 
                  theme(plot.background = element_rect(fill = "transparent",colour = NA),
                        panel.grid.minor=element_line(colour="transparent"))
theme_set(theme_notebook)
```

```{r results="hide"}
knit("~/Documents/code/nonparametric-bayes/inst/examples/beverton_holt_data.Rmd")
true <- data.frame(x = x_grid, y = sapply(x_grid, f, 0, p))
XX <- as.matrix(x_grid)
X <- as.matrix(obs$x)
Z <- as.matrix(obs$y)
```

## Call to `tgp` function `bgp()`

Tweak the hyperparameter prior's parameters (`s2.p`) to influence the GP inferred.

```{r}
gp <- bgp(X=X, XX=XX, Z=Z, meanfn="constant", bprior="b0", BTE=c(1000,6000,2), m0r1=TRUE, verb=4, corr="exp", trace=TRUE, s2.p=c(5,10), tau2.p=c(5,10), s2.lam="fixed", tau2.lam="fixed")
```

## parsing the function call

* `X` is the observed X values (or matrix of appropriate dimension)
* `XX` the desired predicted grid. (will also predict on X)
* `Z` is observed Y values (or matrix of appropriate dimension)
* `m0r1` means scale data to mean zero and unit range
* `bprior` is the prior distribution, defaulting to `bfalt`. `b0` is a Gaussian.  This prior is for the Gaussian process and is not to be confused with the hyperpriors describing the distribution of various parameters.
* `BTE` Burn-in B steps, Terminate after T steps, sample Every E steps. 
* `verb` Very verbose output
* Other options (for adaptive/query learning, etc, igored for now.  Focus on priors).  
* The priors for the hyperparameters are given by `s2.p`, `tau.p` etc (both are Inverse Gammas).  Their parameters can vary hiearchically in general, but we hold them fixed, `s2.lam = 'fixed'`, etc.)

## Hyperparameters

Hyperparameter distributions returned in the trace:

```{r}
names(gp$trace$XX[[1]][2500,])
```

* `s2` is $\sigma^2$ in Gramacy and $\tau$ in Rassmussen and Williams, is the scale of the noise. Comes from an inverse gamma prior.  Set by inverse-gamma parameters `a0` and `g0`.

* `tau2` Also comes from an inverse gamma prior. It is also set by inverse-gamma parameters `a0` and `g0`.

$$Z | \beta, \sigma^2, K \sim N_n(\mathbf{F} \beta, \sigma^2 \mathbf{K}) $$
$$ \beta | \sigma^2, \tau^2, \mathbf{W}, \beta_0 ~ N_m(\beta_0, \sigma^2 \tau^2 \mathbf{W} )  $$
$$\beta_0 | N_m(\mathbf{\mu}, \mathbf{B})$$

The parameters in this run are constant across the trace points (because hyperparameters are fixed).

```{r}
gp$trace$XX[[1]][2500,]
gp$trace$XX[[3]][2500,]
```

```{r}
pars <- melt(gp$trace$XX[[1]], id = "index")
ggplot(pars, aes(index, value)) + geom_line() + facet_wrap(~variable, scales="free_y")
````

It is possible to calculate summary statistics and check out the distribution of these fellows.

```{r}
mean(gp$trace$XX[[1]][,"nug"])
ggplot(pars, aes(value)) + geom_histogram() + facet_wrap(~variable, scales="free")
```

* We keep the heirachical priors fixed in the `tau.lm`, `s2.lm`, etc. `nug` fixed by default.  Note this results in the the priors from which these are sampled are fixed: e.g. `gp$trace$hier$s2.a0`, `gp$trace$hier$s2.g0`

* `nug.p` This is the "nugget," the measurement error.  Comes from a mixture of gamma prior parameter (initial values) for the nugget parameter `c(a1,g1,a2,g2)` where `g1` and `g2` are scale (`1/rate`) parameters. _Default reduces to simple exponential prior._ Specifying `nug.p = 0` fixes the nugget parameter to the "starting" value in `gd[1]`, i.e., it is excluded from the MCMC

* `index` is the sampling point (from `BTE`, we see we sample starting at step 1000 and ending at step 6000, recording every 2 steps, so there are $(T-B)/E = 2500$ index points).  
* `lambda` possibly a mixing parameter in the importance sampler temperature??
* `beta0` No mention in the manual
* `d` The parameter for priors using the hierachical exponential distribution for the parameters `a1`, `a2`, `g1`, `g2`.  
* `b` ??

* Where are the hyperparameters for the correlation function (e.g. the length-scale for the Gaussian??)

## Extracting the estimated process

(The **MAP**, maximum a posteriori expected mean and variance).

Extract the posterior Gaussian process mean and the $\pm 2$ standard deviations over the predicted grid from the fit:

```{r}
V <- gp$ZZ.ks2
dat <- data.frame(x   = gp$XX[[1]], 
                 y    = gp$ZZ.km, 
                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
```

Plot the posterior Gaussian Process:

```{r gp-plot}
  p1 <- ggplot(dat) +
    geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
    geom_line(aes(x=x,y=y), size=1) + #MEAN
    geom_point(data=gp$obs,aes(x=x,y=y)) + 
    labs(title=paste("llik =", prettyNum(gp$llik)))
  if(!is.null(true))
    p1 <- p1 + geom_line(data = true, aes(x, y), col = "red", lty = 2) 
p1 + theme_notebook
```

Find the actual optimum policy from the true parametric model

```{r}
x_grid <- dat$x
h_grid <- x_grid
matrices_true <- f_transition_matrix(f, p, x_grid, h_grid)
opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
```

Estimate a policy from the Gaussian Process:

```{r}
rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))
matrices_gp <- lapply(h_grid, function(h){
    S <- dat$y - h
    F_ <- t(sapply(1:length(S), function(i){
      if(S[i]>0) {
        out <- dlnorm(x_grid/S[i], 0, V[i])
      } else {
        out <- numeric(length(x_grid))
        out[1] <- 1
        out
      }
    }))
    F_ <- rownorm(F_)
  })
opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
```

```{r}
matrices_true[[1]][1:10,1:10]
matrices_gp[[1]][1:10,1:10]

```

```{r policy_plot}
policies <- melt(data.frame(stock=x_grid, GP = x_grid[opt_gp$D[,1]], Exact = x_grid[opt_true$D[,1]]), id="stock")
policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) +
  geom_point() + xlab("stock size") + ylab("escapement") 
policy_plot + theme_notebook
```

We can see what happens when we attempt to manage a stock using this:

``` {r othernoise}
z_g <- function() rlnorm(1,0, sigma_g)
````

``` {r  simulate }
set.seed(1)
sim_gp <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit)
set.seed(1)
sim_true <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit)
````

``` {r  simplot}
df <- data.frame(time = sim_gp$time, stock_gp = sim_gp$fishstock, stock_true = sim_true$fishstock, harvest_gp = sim_gp$harvest, havest_true = sim_true$harvest)
df <- melt(df, id="time")
simplot <- ggplot(df) + geom_line(aes(time,value, color=variable))
simplot + theme_notebook
````

Total Profits

```{r}
sum(sim_gp$profit)
sum(sim_true$profit)
```