### 2012-12-21 inst/examples/myers-exploration.Rmd
fixed bug (had Ef and V in gp_transition_matrix)

```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "myers-cache/", warning=FALSE)
read_chunk('gaussian-process-control.R')
library(knitcitations)
```

```{r libraries, include=FALSE}
```
```{r graphing-options, include=FALSE}
```

Fixed priors on hyperparameters, fixed model type.

```{r gp-priors}
```

```{r sdp-pars-fixed}
profit = function(x,h) pmin(x, h)
delta <- 0.01
OptTime = 20
reward = 0
xT <- 0
z_g = function() rlnorm(1, 0, sigma_g)
z_m = function() 1+(2*runif(1, 0,  1)-1) * sigma_m
```

```{r Myer-explore}
f <- Myer_harvest
p <- c(1, 2, 4.5) 
K <- p[1] * p[3] / 2 + sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2
allee <- p[1] * p[3] / 2 - sqrt( (p[1] * p[3]) ^ 2 - 4 * p[3] ) / 2 # allee threshold
e_star <- (p[1] * sqrt(p[3]) - 2) / 2 ## Bifurcation point 
```

```{r sdp-pars-explore}
sigma_g <- 0.05
sigma_m <- 0.2
x_grid <- seq(0, 1.5 * K, length=101)
h_grid <- x_grid
```

With parameters `r p`.

```{r}

sim_obs <- function(Xo, z_g, f, p, Tobs = 40, seed = 1){
  x <- numeric(Tobs)
  x[1] <- Xo
  set.seed(seed)
  for(t in 1:(Tobs-1))
    x[t+1] = z_g() * f(x[t], h=0, p=p)
  data_plot <- qplot(1:Tobs, x) + ggtitle(paste("seed", seed))
  print(data_plot)
  obs <- data.frame(x=c(0,x[1:(Tobs-1)]),y=c(0,x[2:Tobs]))
  obs
}

par_est <- function(obs){
  estf <- function(p){
    mu <- log(obs$x) + p["r"]*(1-obs$x/p["K"])
    llik <- -sum(dlnorm(obs$y, mu, p["s"]), log=TRUE)
    if(!is.numeric(llik) | is.nan(llik) | !(llik < Inf)){
      warning("possible error in llik")
      llik <- 1e30
    }
    llik
  }
  o <- optim(par = c(r=1,K=mean(obs$x),s=1), estf, method="L", lower=c(1e-3,1e-3,1e-3))
  f_alt <- Ricker
  p_alt <- c(o$par['r'], o$par['K'])
  sigma_g_alt <- o$par['s']
  list(f_alt = f_alt, p_alt = p_alt, sigma_g_alt = sigma_g_alt)
}

optimal_policy <- function(gp, f, f_alt, p, p_alt, x_grid, h_grid, sigma_g, sigma_g_alt, delta, xT, profit, reward, OptTime){
  matrices_gp <- gp_transition_matrix(gp$ZZ.km, gp$ZZ.ks2, x_grid, h_grid)
  opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, OptTime, xT, profit, delta, reward=reward)
  matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
  opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
  matrices_estimated <- f_transition_matrix(f_alt, p_alt, x_grid, h_grid, sigma_g_alt)
  opt_estimated <- find_dp_optim(matrices_estimated, x_grid, h_grid, OptTime, xT, profit, delta=delta, reward = reward)
  list(gp_D = opt_gp$D[,1], true_D = opt_true$D[,1], est_D = opt_estimated$D[,1])
}

simulate_opt <- function(OPT, f, p, x_grid, h_grid, K, z_g, profit){
  gp_D <- sapply(1:OptTime, function(i) OPT$gp_D)
  true_D <- sapply(1:OptTime, function(i) OPT$true_D)
  est_D <- sapply(1:OptTime, function(i) OPT$est_D)
  set.seed(1)
  sim_gp <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, gp_D, z_g, profit=profit))
  set.seed(1)
  sim_true <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, true_D, z_g, profit=profit))
  set.seed(1)
  sim_est <- lapply(1:100, function(i) ForwardSimulate(f, p, x_grid, h_grid, K, est_D, z_g, profit=profit))
  dat <- list(GP = sim_gp, Parametric = sim_est, True = sim_true)
  dat <- melt(dat, id=names(dat[[1]][[1]]))
  dt <- data.table(dat)
  setnames(dt, c("L1", "L2"), c("method", "reps")) 
  dt
}

sim_plots <- function(dt, seed=1){
  fish_plot <- ggplot(dt) + 
    geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1)))+
    ggtitle(paste("seed", seed))
  print(fish_plot)
  harvest_plot <- ggplot(dt) +
    geom_line(aes(time, harvest, group=interaction(reps,method), color=method), alpha=.1) +
    scale_colour_manual(values=cbPalette, guide = guide_legend(override.aes = list(alpha = 1))) +
    ggtitle(paste("seed", seed))
  print(harvest_plot)
}

profits_stats <- function(dt){
  profits <- dt[, sum(profit), by = c("reps", "method")]
  means <- profits[, mean(V1), by = method]
  sds <- profits[, sd(V1), by = method]
  yield <- cbind(means, sd = sds$V1)
  yield
}

gp_plot <- function(gp, f, p, f_alt, p_alt, x_grid, obs){
  tgp_dat <- 
  data.frame(  x = gp$XX[[1]], 
               y = gp$ZZ.km, 
             ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
             ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
  true <- sapply(x_grid, f, 0, p)
  est <- sapply(x_grid, f_alt, 0, p_alt)
  models <- data.frame(x=x_grid, GP=tgp_dat$y, Parametric=est, True=true)
  models <- melt(models, id="x")
  names(models) <- c("x", "method", "value")
  plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
    geom_line(data=models, aes(x, value, col=method), lwd=2, alpha=0.8) + 
    geom_point(data=obs, aes(x,y), alpha=0.8) + 
    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
    scale_colour_manual(values=cbPalette)
  print(plot_gp)
}

```

```{r parallel, message=FALSE, warning=FALSE, results='hide'}
require(snowfall)
sfInit(par=TRUE, cpu=8)
sfExportAll()
sfLibrary(pdgControl)
sfLibrary(nonparametricbayes)
sfLibrary(reshape2)
sfLibrary(ggplot2)
sfLibrary(data.table)
sfLibrary(tgp)
sfLibrary(kernlab)
sfLibrary(MCMCpack)
sfLibrary(plyr)
```

```{r }
seed <- round(runif(16) * 1e6)
seed[1] <- 1
seed
yields <- lapply(seed, 
function(seed_i){
  Xo <- allee + x_grid[5]
  obs <- sim_obs(Xo, z_g, f, p, Tobs=40, seed = seed_i)
  est <- par_est(obs)
  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
          m0r1=FALSE, corr="exp", trace=FALSE, 
          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
  gp_plot(gp, f, p, est$f_alt, est$p_alt, x_grid, obs)
  OPT <- optimal_policy(gp, f, est$f_alt, p, est$p_alt, x_grid, h_grid, sigma_g, est$sigma_g_alt, delta, xT, profit, reward, OptTime)
  dt <- simulate_opt(OPT, f, p, x_grid, h_grid, K, z_g, profit)
  sim_plots(dt, seed=seed_i)
  profits_stats(dt)
})
```

```{r}
yields <- melt(yields, id=c("method", "V1", "sd"))
yields         
````

```{r echo=FALSE, results="asis"}
bibliography("html")
```