### 2012-12-28 inst/examples/allee.Rmd
Example of Ricker-Allee in oscillation regime, shows non-trivial dynamics, GP does very well (closes #22)

```{r set-options, echo = FALSE, cache = FALSE, external = TRUE, include = FALSE}
opts_chunk$set(external = TRUE, cache = FALSE, cache.path = "allee/", warning=FALSE)
library(knitcitations)
```

```{r libraries, include=FALSE}
require(pdgControl)
require(nonparametricbayes)
require(reshape2)
require(ggplot2)
require(data.table)
require(tgp)
library(kernlab)
require(MCMCpack)
require(plyr)

```

```{r graphing-options, include=FALSE}
opts_knit$set(upload.fun = socialR::notebook.url)
#opts_chunk$set(dev = 'Cairo_pdf', dev.args=list(""))
opts_chunk$set(dev="png", dev.args=list(bg="transparent"))
opts_chunk$set(comment=NA, tidy=FALSE)
theme_set(theme_bw(base_size=16))
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
#sha <- gsub("^commit ", "", system("git log -n 1", intern=TRUE)[1])
#short_sha <- gsub("(^.{10}).*", "\\1", sha)
#date <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
#opts_chunk$set(fig.path = paste("figure/", date, "-", short_sha, "-", sep=""))

```

Fixed priors on hyperparameters, fixed model type.

```{r gp-priors}
#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
s2.p <- c(5,5)  
tau2.p <- c(5,1)
d.p = c(10, 1/0.1, 10, 1/0.1)
nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
beta0_prior <- function(x, tau) dnorm(x, 0, tau)
beta = c(0)
priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
```

```{r sdp-pars-fixed}
profit = function(x,h) pmin(x, h)
delta <- 0.01
OptTime = 20  # stationarity with unstable models is tricky thing
reward = 0
xT <- 0

```

```{r RickerAllee-exp}
f <- RickerAllee
# c(5, 10, 5) is 2-cycle, c(5.5, 10, 5) is 6 cycle, 5.3 is about 4
p <- c(5.3, 10, 5) 
K <- 10
allee <- 5
```

```{r sdp-pars-explore}
sigma_g <- 0.001
sigma_m <- 0.0
z_g = function() rlnorm(1, 0, sigma_g)
z_m = function() 1+(2*runif(1, 0,  1)-1) * sigma_m
x_grid <- seq(0, 1.5 * K, length=101)
h_grid <- x_grid
```

With parameters `r p`.

```{r }
seed_i <- 1
  Xo <- K # observations start from
  x0 <- Xo # simulation under policy starts from
```

```{r}
  obs <- sim_obs(Xo, z_g, f, p, Tobs=35, nz=15, 
                 harvest = sort(rep(seq(0, .5, length=7), 5)), seed = seed_i)
```

```{r}
  alt <- par_est(obs,  init = c(r=2, K=mean(obs$x), s=1))
  est <- par_est_allee(obs, f, p,  init = c(2, 6, 2, s = sigma_g))
```

Which estimates a Ricker model with $r =$ `r alt$p[1]`, $K =$ `r alt$p[2]`, and the Allen allee model with $r =$ `r est$p[1]`, $K =$ `r est$p[2]` and $C =$ `r est$p[3]`.

```{r}
  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
          meanfn="constant", bprior="b0", BTE=c(2000,16000,2),
          m0r1=FALSE, corr="exp", trace=TRUE, 
          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
  gp_plot(gp, f, p, est$f, est$p, alt$f, alt$p, x_grid, obs, seed_i)
```

```{r}
  posteriors_plot(gp, priors) # needs trace=TRUE!
```

```{r}
  OPT <- optimal_policy(gp, f, est$f, alt$f,
                        p, est$p, alt$p,
                        x_grid, h_grid, sigma_g, 
                        sigma_g, sigma_g, # est$sigma_g, alt$sigma_g, but those ests are poor
                        delta, xT, profit, reward, OptTime)
  plot_policies(x_grid, OPT$gp_D, OPT$est_D, OPT$true_D, OPT$alt_D)
```

```{r}
dt <- simulate_opt(OPT, f, p, x_grid, h_grid, x0, z_g, profit)
sim_plots(dt, seed=seed_i)
profits_stats(dt)
```

```{r echo=FALSE, results="asis"}
bibliography("html")
```