### 2013-03-19 inst/doc/beyond-mechanistic-models.Rmd
add code to beyond-mechanistic-models. adapt text into manuscript form

```{r}
read_chunk("nonparametric-bayes.R")
```
```{r knit-settings, echo=FALSE, include=FALSE, cache=FALSE}
```
```{r libraries, echo=FALSE, include=FALSE}
```
```{r plotting-options, echo=FALSE, include=FALSE}
```
```{r stateeq}
```
```{r sdp-pars, dependson="stateeq"}
```
```{r obs, dependson="sdp-pars"}
```
```{r mle, dependson="obs"}
```
```{r gp-priors}
```
```{r gp, dependson=c("obs", "gp-priors")}
```
```{r opt, dependson=c("gp", "mle")}
```
```{r sim, dependson="opt"}
```

Process-based models built on modern machine learning approaches 
offer a powerful and underappreciated approach for ecological
forecasting and management.  We present a novel example of how
one such example - non-parametric Gaussian processes, can be adapted
to the context of management and decision-making.

While effective machine learning algorithms have speckled the literature, 
most theoretical ecologists in particular have been wary of these 'black-box'
approaches in favor of simple, mechanistic models.  While a blind application
of machine learning approaches would indeed be dangerous grounds to stake 
either ecological understanding or management practice, such concerns are not
grounds to reject such methods out of hand.  Rather, further attention is 
warranted to identify the circumstances in which machine-learning approaches
will most likely outperform existing more mechanistic approaches, or help
crack difficult management questions where ecological or dynamical complexity
has caused traditional approaches to founder.

Here we illustrate how a machine learning model can be implemented in place
of simple mechanistic model in a decision theory (optimal control) problem,
and highlight the features of this system that make the nonparametric approach
particularly effective.

Mechanistic models have long been the gold standard of theoretical
modeling in ecology (e.g. see @Geritz2012 or @Cuddington2013).  Only by
understanding the processes involved can we make reliable long term
predictions and build an knowledge of cause and effect that guides the
hypotheses we make, the data we collect, and the management decisions
we make.  Process-based models, whether expressed in the language of
mathematics or English, identify the connection between mosquitoes and
the spread of malaria, or greenhouses gases and climate change, guiding
our approach to understand and manage these threats. Despite this central
importance, we argue here that ecologists would do well to give greater
attention to the role non-mechanistic models can play in ecological
management and decision making.  The value of these approaches is
greatest in a context where decisions are made over short time horizons,
and updated as new data becomes available.

## The alternative to process-based models: pattern-based modeling

Before we proceed further it would be useful to define our terms and
lay some greater context for the discussion. Throughout, we will distinguish
"mechanistic" or "process-based" models from models that are merely "correlative" 
or "pattern-based". Historically pattern-based modeling meant regression (usually linear
regression) and resided in methodology rooted in statistics departments
and was the primary focus of empirical ecologists, while mechanistic
modeling meant dynamical systems (ODEs, PDEs, SDEs and their discrete
kin), residing in methodology from the mathematics department and was
the primary focus of theoretical ecologists.

We believe these historical divides continue to color much of the
literature today, with theorists more skeptical than empiricists of
statistical methodology and vice versa.  Meanwhile, the ground underneath
has shifted.  Originally pattern-matching approaches could be critiqued
on the grounds of their simplicity (not everything is linear) while
mechanistic approaches could be critiqued on their tenuous connections
to data -- model parameter values such as death rates, birth rates,
etc. would be estimated in advance and then stuck into the model, rather
than estimated in the context of the model itself.  Today both of these
critiques are outdated. Computational power and hierarchical statistical
methods (particularly approximate likelihood or simulation techniques such
as particle filters and approximate Bayesian computing) have brought the
inference richer dynamical systems into their fold, while pattern-based
modeling has spawned an entirely new approach under the banner of machine
learning that has divided the statistics community (See @Bremman2001,
"The Two Cultures").   Machine learning models can represent almost
arbitrarily complex patterns and incorporate learning and decision making
strategies that have earned them the name 'artificial intelligence.'
Yet at the same time they can be far less transparent their linear
regression predecessors, making both astounding successes and startling
biases and failures like their biological namesake. More ecologists,
particularly theorists and modelers, would do well to learn both how to
take advantage of these strengths and recognize their weaknesses.

A simple example comparing the performance of the two modeling approaches
in a quantitative decision problem will help to introduce these issues.

Mechanistic modeling emphasizes the importance of capturing the correct
gross properties of a system over the tracking minute fluctuations.
For instance, in selecting and parameterizing model of a population
of conservation concern, we may be most interested in getting the
long-term behavior correct -- such as identifying if the dynamics support
persistence of the population -- rather than worrying how well they
reflect the year-to-year fluctuations.  We would certainly have good
reason to prefer such a model over alternatives which are irreconcilable
to the most basic biological processes, such as unbounded growth, or
growth curves that do not pass through the origin in a closed system. So
it may come as a surprise to realize that such obviously wrong models
can perform as well or even better than reasonable mechanistic models
in guiding ecological management and decision making.

```{r gp_plot, dependson=c("gp", "mle", "plotting-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.  Graph shows the expected value for the function $f$ under each model.  Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.  The training data is also shown as black points.  (The GP is conditioned on 0,0, shown as a pseudo-data point). "}
```
```{r policies_plot, dependson=c("opt", "plotting-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.  Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979]."
}
```
```{r sim_plot, dependson=c("sim", "plotting-options"), fig.cap="Gaussian process inference outperforms parametric estimates. Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model."
}
```

An example of such a comparison is shown in Figure 1, in which the optimal
management decision is determined by stochastic dynamic programming
algorithm using the biologically (a) plausible and (b) the implausible
model of the population growth.

Three elements contribute to the biologically implausible phenomenological
model performing better in this context:

### 1. Relevant state space

The dynamics occur over a range of state-space in which the biologically
implausible model performs as well or better than the more plausible
model.  This is the most obvious and immediate reason why the shortcomings
that appear to make the model implausible do not make it useless.
This effect is enhanced in our example because the management actions help
drive the system towards rather than away from this region of state-space.

### 2. Predictive accuracy

This application relies only on the predictive accuracy of the model,
not an interpretation of the parameters.  Predictive accuracy is not the
goal of all modeling, as ecologists have been observing for as long as
they made models (perhaps none more memorably than @Levins1969).

Mechanistic modeling is at its most powerful not when it is used to
directly forecast future states but when it provides an understanding
of how to approach a problem.  SIR-type models from the epidemiological
literature are a good example.  While the simplest SIR models have little
predictive power over the outbreak intensity or timing at a particular
location, they provide a powerful understanding of the spread of an
infection in terms of a single, biologically meaningful parameter: $R_0$,
the basic reproductive number.  From the model, it becomes clear that
management need not vaccinate every member of the population to stop
the spread, but rather it suffices to vaccinate a sufficient fraction
of the population to reduce $R_0$ below 1.

### 3. Time scale for new data

In the sequential decision making problem we considered, we are
presented with new data after each action.  The relevant timescale
for the prediction is thus not the long-term dynamics, which would be
wildly divergent, but the dynamics over this much shorter interval.
While ecologists may be hesitant to base continual management on a model
with obviously inaccurate long-term behavior, engineers tend to consider
the problem in frequency space and gravitate to the opposite position --
a good control model should prioritize high-frequency accuracy over low
frequency accuracy.  The differences in intuition may arise from the
timescales at which each profession can typically adjust their control
variables -- much faster for a control system of a chemical plant than
state policy for a natural resource.  Still, the lesson is clear: when
facing repeated management decisions over a short timescale, such as
setting annual harvests of a fishery, it may be more valuable to use a
machine learning algorithm that makes accurate year-ahead predictions
that capture some of the high-frequency fluctuations that appear only
as noise in a mechanistic model of the long-term population dynamics.

## Advantages of non-mechanistic models

1. Can better use all of the data available.

Incorporating various sources of information into mechanistic models
can be an immensely difficult due to the increased complexity involved.
Only thanks to tandem advances in increasing computational power and
hierarchical statistical methodology have we been able to tackle such
intuitively important complexity (and the potentially new available data
that accompanies it) such as spatial distribution, heterogeneities of
space, time, and individuals, to shift to ecosystem-based approaches
from single-species based approaches.  Without the need to formulate
mechanisms, many modern machine learning algorithms can leverage potential
information from all available sources of data directly.  The algorithms
can recognize unanticipated or subtle patterns in large data sets
that enable more accurate predictions than mechanistic models that are
formulated at a more macroscopic level.  Such approaches depend critically

2. Better express ambiguity in regions where data is not available

One of the greatest strengths of mechanistic models is their greatest weakness as well.

## Better the devil you know?  The dangers of pattern-based approaches

When all we have identified is a pattern rather than a mechanism, there is
little way to guarantee that the pattern will not change in the future.
Few examples illustrate this danger more broadly than the sub-prime
mortgage crisis of 2008.  Simplifying greatly, the unquenchable demand for
mortgage backed securities had shifted the underlying dynamics on which
the phenomenological models were based.  Historical data indicated that
mortgage-backed securities involved little risk. As demand exhausted
the supply of traditional mortgages, it became profitable to invent
new mortgages that could be sold to higher risk customers, such as
those without proof of income.  The pattern-based models did not have
enough time or data to learn about the higher default rate occurring as
a result, but remained anchored too in the old patterns where default
rates were low.  While a mechanistic formulation may have anticipated the
resulting higher risk, the pattern based model must learn by seeing it.
The economy still struggles to pay off that lesson.

Of course mechanistic models may also fail to predict sudden
structural changes.  It is all too easy to overlook some slowly changing
environmental variable, particularly if temporal changes in a parameter
must be modeled explicitly if they are to be included at all.  Machine
learning approaches are frequently less rigid in their assumptions, but
can also less transparent about them.  This creates a second weakness
in the machine learning approach.

Mechanistic models are not free from this danger either. In fact they can
be more susceptible to them for the very same reasons.  A mechanistic
model has the ability to more convincingly produce out-of-sample
predictions or extrapolation.  This may allow us to determine that a
virus will spread or an endangered species will go extinct even before
we have observed

Dangers of biased estimation.

Parametric models are frequently used phenomenologically, rather than
being derived by plausible underlying mechanisms.  @Geritz2012 condemn
this practice on the basis of the common misinterpretations that result.
These mistakes arise from ignoring or misinterpreting the biological
meaning of the model parameters. The use of delay equations is a
frequent offender -- for instance, one is hard-pressed to describe a
mechanistic model consistent with Hutchinson's 1948 delay-logistic.
Assuming either scramble competition or interference competition gives
rise to two alternative delay equations, both structurally different
from the original.

Three elements contribute to this.

A process-based understanding helps us formulate hypotheses, guide
experimental design and data collection, predict outcomes, and influence
management decisions.  The alternative to a processed-based model is  a
correlative one. Statical models machine learning.

> since the late 1980s there has been some consensus amongst ecologists
that management decisions are best guided by models which are grounded
in ecological theory, and which strike a balance between too much or too
little detail describing the relevant processes (e.g., DeAngelis 1988,
Starfield 1997, Jackson et al. 2000, Carpenter 2003, Nelson et al. 2008)

machine learning algorithms

Prediction without

Example of Statistical model vs process-based model.

@Geritz2012 @Cuddington2013

Managing by rule-of-thumb

Managing by algorithm

```{r posteriors, dependson=c("gp", "plotting-options"), fig.cap = "Histogram of posterior distributions for the estimated Gaussian Process shown in Figure 1.  Prior distributions overlaid."}
```