### 2013-03-21 inst/examples/pomp-example.Rmd
pomp example

```{r}
library(pomp)
```

```{r}
process.sim <- function(x, t, params, delta.t, ...){
# r = 0.75, K = 10, Q = 3, H = 1, a = 1.3 
  r <- params["r"]
  K <- params["K"]
  Q <- params["Q"]
  H <- params["H"]
  a <- params["a"]
  sigma <- params["sigma"]
  X <- x["X"]
  epsilon <- exp(rlnorm(n = 1, mean = 0, sd = sigma))

  c(X = unname(
    X * exp(r * delta.t * (1 - X / K) - a * X ^ (Q - 1) / (X ^ Q + H ^ Q)) * epsilon)) 
}
```

```{r}
measure.sim <- function(x, t, params, ...){
  tau <- params["tau"]
  X <- x["X"]
  y <- c(Y = unname(rlnorm(n = 1, meanlog = log(X), sd=tau)))
}
```

Create a container of class `pomp`

```{r}
may <- pomp(data = data.frame(time = 1:100, Y = NA), times = "time", rprocess = discrete.time.sim(step.fun = process.sim, delta.t = 1), rmeasure = measure.sim, t0 = 0)
```

Parameters and an initial condition

```{r}
theta <- c(r = 0.75, K = 10, Q = 3, H = 1, a = 1.3, 
           sigma = 0.1, tau = 0.1, X.0 = 5) 
```
```

Simulate

```{r}
dat <- simulate(may, params=theta, obs=TRUE, states=TRUE)
library(reshape2)
library(ggplot2)
dat <- melt(dat)[3:5]
names(dat) = c("time", "value", "variable")
ggplot(dat, aes(time, value, lty=variable)) + geom_line()
```

Note that simulate returns an instance of our object, with the data column filled in.  This is not true if we give it arguments such as state or obs = TRUE.

Measurement density

```{r}
measure.dens <- function(y, x, t, params, log, ...) {
    tau <- params["tau"]
    ## state at time t:
    X <- x["X"]
    ## observation at time t:
    Y <- y["Y"]
    ## compute the likelihood of Y|X,tau
    dlnorm(x = Y, meanlog = log(X), sdlog = tau, log = log)
}
```

Add to our container

```{r}
may <- pomp(may, dmeasure = measure.dens)
```

Optionally, work in the transformed variable space instead by providing parameter and inverse parameter transforms.

```{r}
may  <- pomp(may, parameter.transform = function(params, 
    ...) {
    exp(params)
}, parameter.inv.transform = function(params, ...) {
    log(params)
})
```

Run the particle filter

```{r}
pf <- pfilter(may, params = theta, Np = 1000)
logLik(pf)
```

```{r}

```