### 2013-04-19 inst/examples/BUGS/par_uncertainty_sdp.Rmd
parametric bayes intial example

```{r}
f_transition_matrix <- function(f, p, x_grid, h_grid, sigma_g, pardist){
  lapply(h_grid, par_F, f, p, x_grid, sigma_g, pardist)
}

par_F <- function(h, f, p, x_grid, sigma_g, pardist, n_mc = 100){

  # Set up monte carlo sampling 
  d <- dim(pardist)
  indices <- round(runif(n_mc,1, d[1]))

  # compute mean with monte carlo samples
  mu <- sapply(indices, function(i) { # Cols are parameter samples
    p <- unname(pardist[i,c(4,1,5)])          # Only, we need to add the measurement and process noise  
    unname(sapply(x_grid, f, h, p))   # rows are possible values of x_t
  })
  mu <- data.frame(t(mu)) # Let's have cols as possible x_t, rows as parameters, and type as data.frame

  F_true <- 
    sapply(mu, function(cur_state){    # For each x_t

      bypar <- sapply(cur_state, function(m){ # For each parameter value
        if(snap_to_grid(m,x_grid) < x_grid[2]){ # 
          out <- numeric(length(x_grid))
          out[1] <- 1
          out
        } else {
          out <- dlnorm(x_grid/m, 0, sigma_g) # should be using the estimate process noise!
        }})  
      ave_over_pars <- apply(bypar, 1, sum) # collapse by weighted average over possible parameters    
      ave_over_pars / sum(ave_over_pars)
      })

  F_true <- t(F_true)

}

snap_to_grid <- function(x, grid) sapply(x, function(x) grid[which.min(abs(grid - x))])

# internal helper function
rownorm <- function(M) 
  t(apply(M, 1, function(x){ 
    if(sum(x)>0){
      x/sum(x)
    } else {
      out <- numeric(length(x))
      out[1] <- 1
      out
    }
  }))
```

```{r}
knit("allen.Rmd")

pardist <- mcmcall
#pardist[,1] = p[2] + rnorm(100, 0, 0.000001)
#pardist[,4] = p[1] + rnorm(100, 0, 0.000001)
#pardist[,5] = p[3] + rnorm(100, 0, 0.000001)

sdp = f_transition_matrix(f, p, x_grid, h_grid, sigma_g, pardist)
s_opt <- value_iteration(sdp, x_grid, h_grid, OptTime=1000, xT, profit, delta)
#opt <- find_dp_optim(sdp, x_grid, h_grid, OptTime, xT, profit, delta, reward=0)

SDP_Mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g)
pars_fixed <- value_iteration(SDP_Mat, x_grid, h_grid, OptTime=1000, xT, profit, delta)

require(reshape2)
policies <- melt(data.frame(stock=x_grid, pars.uncert = x_grid[s_opt$D], pars.fixed = x_grid[pars_fixed$D]), id="stock")

ggplot(policies, aes(stock, value, color=variable)) + geom_point(alpha=.5) + xlab("stock size") + ylab("harvest")

ggplot(policies, aes(stock, stock - value, color=variable)) + geom_line(alpha=1) + xlab("stock size") + ylab("escapement") 
```