### 2013-04-25 inst/examples/BUGS/2013-04-19-bayesian-parametric-uncertainty-sdp.Rmd
updated examples

---
layout: post
category: ecology
tags:
- nonparametric-bayes

---

```{r libs, include=FALSE, cache=FALSE}
library(pdgControl)
library(nonparametricbayes)
library(reshape2)
library(ggplot2)
library(knitcitations)
```

```{r setupall, echo=FALSE}
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE)
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

```

Adapting the parametric uncertainty represented by the posterior distributions of the Bayesian estimate (see [earlier notes](http://carlboettiger.info/2013/04/12/parametric-bayesian-example.html)) to the stochastic dynamic programming solution for the optimal policy.  Simply requires evaluating the expectation over the distribution, but is computationally intensive given the spread and the three parameters.

```{r}
f_transition_matrix <- function(f, p, x_grid, h_grid, sigma_g, pardist){
  lapply(h_grid, par_F, f, p, x_grid, sigma_g, pardist)
}

par_F <- function(h, f, p, x_grid, sigma_g, pardist, n_mc = 100){

  # Set up monte carlo sampling 
  d <- dim(pardist)
  indices <- round(runif(n_mc,1, d[1]))

  # Compute the matrix
  F_true <- 
    sapply(x_grid, function(x_t){           # For each x_t

      bypar <- sapply(indices, function(i){ # For each parameter value
        p <- unname(pardist[i,c(4,1,5)])    # parameters for the mean, at current sample 
        mu <- f(x_t,h,p)
        est_sigma_g <- pardist[i,2]         # Variance parameter

        if(snap_to_grid(mu,x_grid) < x_grid[2]){ # handle the degenerate case 
          out <- numeric(length(x_grid))
          out[1] <- 1
          out
        } else {
          out <- dlnorm(x_grid/mu, 0, est_sigma_g) 
        }
      })  
      ave_over_pars <- apply(bypar, 1, sum) # collapse by weighted average over possible parameters    
      ave_over_pars / sum(ave_over_pars)
      })
  F_true <- t(F_true)

}

snap_to_grid <- function(x, grid) sapply(x, function(x) grid[which.min(abs(grid - x))])

# internal helper function
rownorm <- function(M) 
  t(apply(M, 1, function(x){ 
    if(sum(x)>0){
      x/sum(x)
    } else {
      out <- numeric(length(x))
      out[1] <- 1
      out
    }
  }))
```

Run the Bayesian analysis to obtain posterior distributions for parameters.

```{r, echo=TRUE, include=FALSE}
knit("allen.Rmd")
```

## Test case using perturbed parameters

First, sanity test. Use the correct parameter values (slightly perturbed).

```{r}
pardist <- mcmcall
pardist[,1] = p[2] + rnorm(100, 0, 0.000001)
pardist[,4] = p[1] + rnorm(100, 0, 0.000001)
pardist[,2] = sigma_g + rnorm(100, 0, 0.000001)
pardist[,5] = p[3] + rnorm(100, 0, 0.000001)
```

Compute optimal policy

```{r}
sdp = f_transition_matrix(f, p, x_grid, h_grid, sigma_g, pardist)
s_opt <- value_iteration(sdp, x_grid, h_grid, OptTime=1000, xT, profit, delta)
```

Compare to the case without parameter uncertainty (growth noise only)

```{r}
SDP_Mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g)
pars_fixed <- value_iteration(SDP_Mat, x_grid, h_grid, OptTime=1000, xT, profit, delta)
```

Plot results

```{r}
require(reshape2)
policies <- melt(data.frame(stock=x_grid, pars.uncert = x_grid[s_opt$D], pars.fixed = x_grid[pars_fixed$D]), id="stock")
ggplot(policies, aes(stock, stock - value, color=variable)) + geom_line(alpha=1) + xlab("stock size") + ylab("escapement") 
```

## Using actual estimates

```{r}
pardist <- mcmcall
```

Transform parameters back

```{r}
pardist[,4] = exp(pardist[,4])
pardist[,5] = exp(pardist[,5])
```

Compute optimal policy

```{r}
sdp = f_transition_matrix(f, p, x_grid, h_grid, sigma_g, pardist)
s_opt <- value_iteration(sdp, x_grid, h_grid, OptTime=1000, xT, profit, delta)
```

Compare to the case without parameter uncertainty (growth noise only)

```{r}
SDP_Mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g)
pars_fixed <- value_iteration(SDP_Mat, x_grid, h_grid, OptTime=1000, xT, profit, delta)
```

Plot results

```{r}
require(reshape2)
policies <- melt(data.frame(stock=x_grid, pars.uncert = x_grid[s_opt$D], pars.fixed = x_grid[pars_fixed$D]), id="stock")
ggplot(policies, aes(stock, stock - value, color=variable)) + geom_line(alpha=1) + xlab("stock size") + ylab("escapement") 
```

## To Do

So far this is just a proof of principle example.

* Needs to be adjusted to account for uncertainty in the estimates of the noise processes as well. 
* Needs to be written as generic and documented functions, add to nonparametric-bayes package routines.