### 2013-04-25 inst/examples/BUGS/par_uncertainty_sdp.Rmd

Run the Bayesian analysis to obtain posterior distributions for parameters.

```{r, echo=TRUE, include=FALSE}
knit("allen.Rmd")
```

```{r}
pardist <- mcmcall
```

Transform parameters back

```{r}
pardist[,4] = exp(pardist[,4])
pardist[,5] = exp(pardist[,5])
```

Compute optimal policy

```{r}
sdp = f_transition_matrix(f, p, x_grid, h_grid, sigma_g, pardist)
s_opt <- value_iteration(sdp, x_grid, h_grid, OptTime=1000, xT, profit, delta)
```

Compare to the case without parameter uncertainty (growth noise only)

```{r}
SDP_Mat <- determine_SDP_matrix(f, p, x_grid, h_grid, sigma_g)
pars_fixed <- value_iteration(SDP_Mat, x_grid, h_grid, OptTime=1000, xT, profit, delta)
```

Plot results

```{r}
require(reshape2)
policies <- melt(data.frame(stock=x_grid, pars.uncert = x_grid[s_opt$D], pars.fixed = x_grid[pars_fixed$D]), id="stock")
ggplot(policies, aes(stock, stock - value, color=variable)) + geom_line(alpha=1) + xlab("stock size") + ylab("escapement") 
```