### 2013-04-30 inst/examples/BUGS/nonparametric.Rmd
exploring comparison cases

## Non-parametric Bayes

```{r gp-priors}
#inv gamma has mean b / (a - 1) (assuming a>1) and variance b ^ 2 / ((a - 2) * (a - 1) ^ 2) (assuming a>2)
s2.p <- c(5,5)  
tau2.p <- c(5,1)
d.p = c(10, 1/0.1, 10, 1/0.1)
nug.p = c(10, 1/0.1, 10, 1/0.1) # gamma mean
s2_prior <- function(x) dinvgamma(x, s2.p[1], s2.p[2])
tau2_prior <- function(x) dinvgamma(x, tau2.p[1], tau2.p[2])
d_prior <- function(x) dgamma(x, d.p[1], scale = d.p[2]) + dgamma(x, d.p[3], scale = d.p[4])
nug_prior <- function(x) dgamma(x, nug.p[1], scale = nug.p[2]) + dgamma(x, nug.p[3], scale = nug.p[4])
beta0_prior <- function(x, tau) dnorm(x, 0, tau)
beta = c(0)
priors <- list(s2 = s2_prior, tau2 = tau2_prior, beta0 = dnorm, nug = nug_prior, d = d_prior, ldetK = function(x) 0)
```

Estimate the Gaussian Process (nonparametric Bayesian fit)

```{r}
  gp <- bgp(X=obs$x, XX=x_grid, Z=obs$y, verb=0,
          meanfn="constant", bprior="b0", BTE=c(20,16000,2),
          m0r1=FALSE, corr="exp", trace=F, 
          beta = beta, s2.p = s2.p, d.p = d.p, nug.p = nug.p, tau2.p = tau2.p,
          s2.lam = "fixed", d.lam = "fixed", nug.lam = "fixed", tau2.lam = "fixed")      
```

Show traces and posteriors against priors

```{r, eval=FALSE}
hyperparameters <- c("index", "s2", "tau2", "beta0", "nug", "d", "ldetK")
posteriors <- melt(gp$trace$XX[[1]][,hyperparameters], id="index")

# Traces 
ggplot(posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)

## A list of functions has been defined in "priors", 
## Here we just make sure to compute priors over the range of values returned by the posteriors
prior_curves <- ddply(posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100)
    data.frame(value = grid, density = priors[[dd$variable[1]]](grid))
})

# Posteriors (easier to read than histograms)
ggplot(posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

(Note that I keep thinking this should be done with `stat_function`, but cannot `facet_wrap` across different functions, see [SO/1376967](http://stackoverflow.com/questions/1376967/using-stat-function-and-facet-wrap-together-in-ggplot2-in-r)