### 2013-04-30 inst/examples/BUGS/setup.Rmd
breaking out sections from allen.Rmd for individual testing.

Plotting and knitr options, (can generally be ignored)

```{r plotting-options}
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA)

library(knitcitations)

library(ggplot2) # plotting 
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=10))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```
Load necessary libraries,

```{r libraries, cache=FALSE, message=FALSE, warning=FALSE}
library(nonparametricbayes) # loads the rest as dependencies
```

### Model and parameters

Uses the model derived in `citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,

```{r}
f <- RickerAllee
p <- c(.2, 7.5, 5)
K <- p[2]
allee <- p[3]
```

Various parameters defining noise dynamics, grid, and policy costs.

```{r sdp-pars, dependson="stateeq"}
sigma_g <- 0.05
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
x_grid <- seq(0, 1.5 * K, length=101)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 20  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
seed_i <- 111
Xo <- K # observations start from
x0 <- Xo # simulation under policy starts from
Tobs <- 35
```

### Sample Data

```{r obs, dependson="sdp-pars"}
  #harvest <- sort(rep(seq(0, .5, length=7), 5))
  x <- numeric(Tobs)
  x[1] <- 6
  nz <- 1
  for(t in 1:(Tobs-1))
    x[t+1] = z_g() * f(x[t], h=0, p=p)
  obs <- data.frame(x = c(rep(0,nz), 
                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
                    y = c(rep(0,nz), 
                          x[2:Tobs]))
write.csv(obs, "obs.txt")
```