### 2013-05-06 inst/examples/BUGS/bayesian-ricker.Rmd
test script for ricker (good for template for extending BUGS)

# Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach

### Parametric Bayes

We initiate the MCMC chain (`init_p`) using the true values of the parameters `p` from the simulation.  While impossible in real data, this gives the parametric Bayesian approach the best chance at succeeding.  `y` is the timeseries (recall `obs` has the $x_t$, $x_{t+1}$ pairs)

```{r}
# a bit unfair to start with the correct values, but anyhow...
init_p = p
names(init_p) = c("r0", "K")
y <- obs$x[-1] 
N=length(y);
```

We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.

```{r}
jags.data <- list("N","y")
n.chains = 1
n.iter = 40000
n.burnin = floor(10000)
n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
```

The actual model is defined in a `model.file` that contains an R function that is automatically translated into BUGS code by *R2WinBUGS*.  The file defines the priors and the model, as seen when read in here

```{r}
cat(readLines(con="ricker-UPrior.txt"), sep="\n")
```

We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.  We use logarithms to maintain strictly positive values of parameters where appropriate.  Because our priors on the noise parameters are inverse gamma distributed.

```{r}
# Uniform priors on standard deviation terms
jags.params=c("K","logr0", "stdQ", "stdR")
jags.inits <- function(){
  list("K"=init_p["K"],"logr0"=log(init_p["r0"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

set.seed(12345)

time_jags <- system.time(       
  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="ricker-UPrior.txt")
)         
time_jags <- unname(time_jags["elapsed"]);
```

#### Convergence diagnostics for parametric bayes

```{r}
jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")

# Traces
ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)

## priors (untransformed variables)
K_prior <- function(x) dunif(x, 0.01, 40)
logr_prior <- function(x) dunif(x, -6, 6)
stdQ_prior <- function(x) dunif(x, 0.001, 100)
stdR_prior <- function(x) dunif(x, 0.001, 100)

par_priors <- list(K = K_prior, deviance = function(x) 0 * x, logr0 = logr_prior, stdQ = stdQ_prior, stdR = stdR_prior)

par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})

# posterior distributions
ggplot(par_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)

```

```{r}
# um, cleaner if we were just be using the long form, par_posterior
mcmc <- as.mcmc(jagsfit)
mcmcall <- mcmc[,-2]
who <- colnames(mcmcall)
who 
mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4])
colnames(mcmcall) <- who
```

```{r}
ricker_pardist <- mcmcall
ricker_pardist[,2] = exp(ricker_pardist[,2]) # transform model parameters back first

bayes_coef <- apply(ricker_pardist,2,mean)
ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
ricker_bayes_pars
```