### 2013-05-07 inst/examples/BUGS/bayesian-ricker.Rmd

```{r}
setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
opts_chunk$set(tidy = FALSE, warning = FALSE, message = FALSE, cache = FALSE, 
    comment = NA, fig.width = 6, fig.height = 4)
library(knitcitations)
library(ggplot2)  # plotting
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size = 10))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA), 
    plot.background = element_rect(fill = "transparent", colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

Load necessary libraries,

```{r}
library(nonparametricbayes) # loads the rest as dependencies
```

### Model and parameters

Uses the model derived in `r citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,

```{r}
f <- RickerAllee
p <- c(2, 8, 5)
K <- p[2]
allee <- p[3]
```

Various parameters defining noise dynamics, grid, and policy costs.

```{r}
sigma_g <- 0.05
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
x_grid <- seq(0, 1.5 * K, length=50)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 30  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
Xo <- K # observations start from
x0 <- Xo # simulation under policy starts from
Tobs <- 40
```

### Sample Data

```{r}
  set.seed(1234)
  #harvest <- sort(rep(seq(0, .5, length=7), 5))
  x <- numeric(Tobs)
  x[1] <- 6
  nz <- 1
  for(t in 1:(Tobs-1))
    x[t+1] = z_g() * f(x[t], h=0, p=p)
  obs <- data.frame(x = c(rep(0,nz), 
                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
                    y = c(rep(0,nz), 
                          x[2:Tobs]))
```

## Write the external bugs file

```{r}
logr0_prior_p <- c(-6.0, 6.0)
logtheta_prior_p <- c(-6.0, 6.0)
logK_prior_p <- c(-6.0, 6.0)
stdQ_prior_p <- c(0.0001, 100)
stdR_prior_p <- c(0.0001, 100)

bugs.model <- 
paste(sprintf(
"model{
  logr0    ~ dunif(%s, %s)
  logtheta    ~ dunif(%s, %s)
  logK    ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)
  stdR ~ dunif(%s, %s)", 
  logr0_prior_p[1], logr0_prior_p[2],
  logtheta_prior_p[1], logtheta_prior_p[2],
  logK_prior_p[1], logK_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2],
  stdR_prior_p[1], stdR_prior_p[2]), 
  "

  iQ <- 1 / (stdQ * stdQ);
  iR <- 1 / (stdR * stdR);

  r0 <- exp(logr0)
  theta <- exp(logtheta)
  K <- exp(logK)

  x[1] ~ dunif(0, 10)

  for(t in 1:(N-1)){
    mu[t] <- r0 * pow(abs(x[t]), theta) / (1 + pow(abs(x[t]), theta) / K)
    x[t+1] ~ dnorm(mu[t], iQ) 
  }

  for(t in 1:(N)){
    y[t] ~ dnorm(x[t], iR)
  }
}")

writeLines(bugs.model, "meyers.bugs")
```

```{r}

## priors (untransformed variables)
logK_prior     <- function(x) dunif(x, logK_prior_p[1], logK_prior_p[2])
logr_prior     <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
stdQ_prior     <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
stdR_prior     <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])

```

### Parametric Bayes based on Meyers model

Contains the

```{r}

init_p = c(1, 1, 1)
names(init_p) = c("r0", "theta", "K")
y <- obs$x[-1] 
N=length(y);
```

We'll be using the JAGS Gibbs sampler, a recent open source BUGS implementation with an R interface that works on most platforms.  We initialize the usual MCMC parameters; see `?jags` for details.

```{r}
jags.data <- list("N","y")
n.chains = 1
n.iter = 40000
n.burnin = floor(10000)
n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
```

We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.  We use logarithms to maintain strictly positive values of parameters where appropriate.  Because our priors on the noise parameters are inverse gamma distributed.

```{r}
jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
jags.inits <- function(){
  list("logr0"=log(init_p["r0"]), "logtheta" = log(init_p["theta"]),  
       "logK"=log(init_p["K"]), "stdQ"=sqrt(0.05), 
       "stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
```

```{r}
set.seed(12345)
time_jags <- system.time(       
  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, 
                  n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
                  n.burnin=n.burnin, model.file="meyers.bugs")
)         
time_jags <- unname(time_jags["elapsed"]);
```

#### Convergence diagnostics for parametric bayes

```{r}
jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")

## Traces
ggplot(par_posteriors) + geom_line(aes(index, value)) +
  facet_wrap(~ variable, scale="free", ncol=1)

```

```{r}
## prior curve functions, must be in order
par_priors <- list( deviance = function(x) 0 * x, logK = logK_prior,
                    logr0 = logr_prior, logtheta = logtheta_prior, 
                    stdQ = stdQ_prior, stdR = stdR_prior)

## calculate points along the prior curve
par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})
```

```{r}
## posterior distributions
ggplot(par_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

```{r}
meyers_pardist <- jags_matrix[2:6]
meyers_pardist[,2] = exp(meyers_pardist[,1]) # transform model parameters back first
meyers_pardist[,2] = exp(meyers_pardist[,2]) # transform model parameters back first
meyers_pardist[,2] = exp(meyers_pardist[,3]) # transform model parameters back first
colnames(meyers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")

bayes_coef <- apply(meyers_pardist,2,mean)
meyers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
```