### 2013-05-13 inst/examples/BUGS/step_ahead_plots.Rmd
plots for step-ahead predictions of each model (from allen.Rmd)

One-step ahead predictor plots

```{r}

step_ahead <- function(x, f, p){
  h = 0
  x_predict <- sapply(x, f, h, p)
  n <- length(x_predict) - 1
  y <- c(x[1], x_predict[1:n])
  y
}

df <- melt(data.frame(time = 1:length(x), stock = x, 
                True = step_ahead(x,f,p)
                ,  
                MLE = step_ahead(x,f,est$p)
                , 
                Parametric.Bayes = step_ahead(x, f, bayes_pars)
                , 
                Ricker = step_ahead(x,alt$f, ricker_bayes_pars)
                , 
                Myers = step_ahead(x, Myer_harvest, myers_bayes_pars)
                 ), id=c("time", "stock"))

ggplot(df) + geom_point(aes(time, stock)) + 
  geom_line(aes(time, value, col=variable)) +
    scale_colour_manual(values=colorkey)

```