### 2013-05-15 inst/examples/BUGS/step_ahead_plots.Rmd
tweak figure layout

One-step ahead predictor plots

```{r plotting-opts}
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=FALSE, comment=NA,
               fig.width=7, fig.height=4)

library(knitcitations)

library(ggplot2) # plotting 
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=12))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}

gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)

step_ahead <- function(x, f, p){
  h = 0
  x_predict <- sapply(x, f, h, p)
  n <- length(x_predict) - 1
  y <- c(x[1], x_predict[1:n])
  y
}

df <- melt(data.frame(time = 1:length(x), stock = x, 
                GP = gp_f_at_obs$E_Ef,
                True = step_ahead(x,f,p),  
                MLE = step_ahead(x,f,est$p), 
                Parametric.Bayes = step_ahead(x, f, bayes_pars), 
                Ricker = step_ahead(x,alt$f, ricker_bayes_pars), 
                Myers = step_ahead(x, Myer_harvest, myers_bayes_pars)
                 ), id=c("time", "stock"))

ggplot(df) + geom_point(aes(time, stock)) + 
  geom_line(aes(time, value, col=variable)) +
    scale_colour_manual(values=colorkey) 
```

```{r}
gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)

ricker_f <- function(x,h,p){
  sapply(x, function(x){ 
    x <- pmax(0, x-h) 
    pmax(0, x * exp(p[2] * (1 - x / p[1] )) )
  })
}

require(MASS)
df_post <- melt(lapply(sample(100), 
  function(i){
    data.frame(time = 1:length(x), stock = x, 
                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
                True = step_ahead(x,f,p),  
                MLE = step_ahead(x,f,est$p), 
                Parametric.Bayes = step_ahead(x, allen_f, pardist[i,]), 
                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
  }), id=c("time", "stock"))

df <- df_post
#df <- subset(df_post, variable %in% c("GP"))
#df <- subset(df_post, variable %in% c("Parametric.Bayes"))
#df <- subset(df_post, variable %in% c("Ricker", "Parametric.Bayes", "Myers"))

fig1 <- ggplot(df) + geom_point(aes(time, stock)) + 
  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
#  geom_point(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1, pch="+") + 
  facet_wrap(~variable) + 
  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) 
fig1
```

```{r}
Tobs <- 8
x <- numeric(Tobs)
x[1] = 5.5
for(t in 1:(Tobs-1))
    x[t+1] = z_g() * f(x[t], h=0, p=p)
y <- x
x <- numeric(Tobs)
x[1] = 4.5
for(t in 1:(Tobs-1))
    x[t+1] = z_g() * f(x[t], h=0, p=p)
x <- c(y,x)
```

```{r}
gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)

ricker_f <- function(x,h,p){
  sapply(x, function(x){ 
    x <- pmax(0, x-h) 
    pmax(0, x * exp(p[2] * (1 - x / p[1] )) )
  })
}

require(MASS)
df_post <- melt(lapply(sample(100), 
  function(i){
    data.frame(time = 1:length(x), stock = x, 
                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
                True = step_ahead(x,f,p),  
                MLE = step_ahead(x,f,est$p), 
                Parametric.Bayes = step_ahead(x, allen_f, pardist[i,]), 
                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
  }), id=c("time", "stock"))

df <- df_post
#df <- subset(df_post, variable %in% c("GP"))
#df <- subset(df_post, variable %in% c("Parametric.Bayes"))
#df <- subset(df_post, variable %in% c("Ricker", "Parametric.Bayes", "Myers"))

fig1 <- ggplot(df) + geom_point(aes(time, stock)) + 
  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
#  geom_point(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1, pch="+") + 
  facet_wrap(~variable) + 
  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) 
fig1
```

```{r}
i <- 15
Tobs <- 5
h_i <- 1

df <- data.frame(
  x = x_grid,  
  GP = matrices_gp[[h_i]][i,], 
  True = matrices_true[[h_i]][i,], 
  MLE = matrices_estimated[[h_i]][i,], 
  Parametric.Bayes = matrices_par_bayes[[h_i]][i,], 
  Ricker = matrices_alt[[h_i]][i,],
  Myers = matrices_myers[[h_i]][i,])

df2 <- data.frame(
  x = x_grid,
  GP = (matrices_gp[[h_i]] %^% Tobs)[i,], 
  True = (matrices_true[[h_i]] %^% Tobs)[i,], 
  MLE = (matrices_estimated[[h_i]] %^% Tobs)[i,],
  Parametric.Bayes = (matrices_par_bayes[[h_i]] %^% Tobs)[i,], 
  Ricker = (matrices_alt[[h_i]] %^% Tobs)[i,], 
  Myers = (matrices_myers[[h_i]] %^% Tobs)[i,])

T2 <- 2 * Tobs

df4 <- data.frame(
  x = x_grid,
  GP = (matrices_gp[[h_i]] %^% T2)[i,], 
  True = (matrices_true[[h_i]] %^% T2)[i,], 
  MLE = (matrices_estimated[[h_i]] %^% T2)[i,],
  Parametric.Bayes = (matrices_par_bayes[[h_i]] %^% T2)[i,], 
  Ricker = (matrices_alt[[h_i]] %^% T2)[i,], 
  Myers = (matrices_myers[[h_i]] %^% T2)[i,])

forecasts <- melt(list(T_start = df, T_mid = df2, T_end = df4), id="x")

ggplot(forecasts) + geom_line(aes(x, value, group=interaction(variable, L1), col=variable, lty=L1)) +
  facet_wrap(~ variable, scale="free_y", ncol=2) + 
  scale_colour_manual(values=colorkey) 
```