### 2013-06-07 inst/examples/BUGS/parametric-vs-nonparametric.Rmd
example set up based on Myers model simulations

# Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach

Plotting and knitr options, (can generally be ignored)

```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
source("~/.knitr_defaults.R")
#opts_knit$set(upload.fun = socialR::flickr.url)
library(knitcitations)
library(nonparametricbayes) 
opts_chunk$set(external=TRUE)
read_chunk("external-chunks.R")
```

```{r posterior-mode}
```

### Model and parameters

```{r stateeq-local}
f <- Myers
p <- c(1, 2, 6)
K <- 5  # approx, a li'l' less
allee <- 1.2 # approx, a li'l' less
```

Various parameters defining noise dynamics, grid, and policy costs.

```{r sdp-pars, dependson="stateeq-local"}
```

### Sample Data

```{r obs, dependson="sdp-pars"}
```

## Maximum Likelihood

```{r mle, dependson="obs"}
```

Mean predictions

```{r mle-output, dependson="mle"}
```

## Non-parametric Bayes

```{r gp-priors}
```

Estimate the Gaussian Process (nonparametric Bayesian fit)

```{r gp, dependson=c("gp-priors", "obs")}
```

Show traces and posteriors against priors

```{r gp_traces_densities, dependson="gp"}
```

```{r gp-output, dependson="gp"}
```

Parametric Bayesian Models
---------------------------

We use the JAGS Gibbs sampler, a recent open source BUGS
implementation with an R interface that works on most platforms.
We initialize the usual MCMC parameters; see `?jags` for details.

All parametric Bayesian estimates use the following basic parameters for the JAGS MCMC:

```{r jags-setup}
```

We will use the same priors for process and observation noise in each model,

```{r common-priors}
```

### Parametric Bayes of correct (Allen) model

We initiate the MCMC chain (`init_p`) using the true values of the
parameters `p` from the simulation.  While impossible in real data, this
gives the parametric Bayesian approach the best chance at succeeding.
`y` is the timeseries (recall `obs` has the $x_t$, $x_{t+1}$ pairs)

The actual model is defined in a `model.file` that contains an R function
that is automatically translated into BUGS code by *R2WinBUGS*.  The file
defines the priors and the model. We write the file from R as follows:

```{r allen-model}
```

Write the priors into a list for later reference

```{r allen-priors, dependson="common-priors"}
```

We define which parameters to keep track of, and set the initial values of
parameters in the transformed space used by the MCMC.  We use logarithms
to maintain strictly positive values of parameters where appropriate.

```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results="hide"}
```

#### Convergence diagnostics for Allen model

R notes: this strips classes from the `mcmc.list` object (so that we have list of matrices; objects that `reshape2::melt` can handle intelligently), and then combines chains into one array. In this array each parameter is given its value at each sample from the posterior (index) for each chain.

```{r allen-traces, dependson="allen-mcmc"}
```

```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
```

Reshape the posterior parameter distribution data, transform back into original space, and calculate the mean parameters and mean function

```{r allen-output, dependson=c("posterior-mode", "allen-traces")}
```

## Parametric Bayes based on the structurally wrong model (Ricker)

```{r ricker-model}
```

Compute prior curves

```{r ricker-priors, dependson="common-priors"}
```

We define which parameters to keep track of, and set the initial values of
parameters in the transformed space used by the MCMC.

```{r ricker-mcmc, dependson="ricker-model"}
```

#### Convergence diagnostics for parametric bayes Ricker model

```{r ricker-traces, dependson="ricker-mcmc"}
```

```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors")}
```

Reshape posteriors data, transform back, calculate mode and corresponding function.

```{r ricker-output, dependson=c("posterior-mode", "ricker-traces")}
```

## Myers Parametric Bayes

```{r myers-model}
```

```{r myers-priors}
```

```{r myers-mcmc, dependson="myers-model"}
```

Convergence diagnostics for parametric bayes

```{r myers-traces, dependson="myers-mcmc"}
```

```{r myers-posteriors, dependson="myers-traces"}
```

```{r myers-output, dependson=c("posterior-mode", "myers-traces")}
```

### Phase-space diagram of the expected dynamics

```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
```

```{r Figure1, dependson="assemble-models"}
```

## Goodness of fit

This shows only the mean predictions.  For the Bayesian cases, we can instead loop over the posteriors of the parameters (or samples from the GP posterior) to get the distribution of such curves in each case.

```{r Figureb, dependson=c("assemble-models", "par-fns")}
```

## Optimal policies by value iteration

Compute the optimal policy under each model using stochastic dynamic programming. We begin with the policy based on the GP model,

```{r gp-opt, dependson="gp-output"}
```

Determine the optimal policy based on the allen and MLE models

```{r mle-opt, dependson="mle-output"}
```

Determine the optimal policy based on Bayesian Allen model

```{r allen-opt, dependson="allen-output"}
```

Bayesian Ricker

```{r ricker-opt, dependson="ricker-output"}
```

Bayesian Myers model

```{r myers-opt, dependson="myers-output"}
```

Assemble the data

```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
```

## Graph of the optimal policies

```{r Figure2, dependson="assemble-opt"}
```

## Simulate 100 realizations managed under each of the policies

```{r sims, dependson="assemble-opt"}
```

```{r Figure3, dependson="sims"}
```

```{r profits, dependson="sims"}
```

```{r totalprofits, dependson="profits"}
```

```{r deviances, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output") }
```