### 2013-06-19 inst/examples/semi-analytic-posteriors.Rmd
update manuscript examples

# Semi-analytic posteriors

```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
source("~/.knitr_defaults.R")
require(reshape2)
require(data.table)
opts_chunk$set(warning=TRUE, message=TRUE, cache=FALSE)
opts_knit$set(upload.fun = socialR::flickr.url)
```

### Generating Model and parameters

Ricker model, parameterized as

$$X_{t+1} = X_t r e^{-\beta X_t + \sigma Z_t}$$

for random unit normal $Z_t$

```{r stateeq}
f <- function(x,h,p)  x * p[1] * exp(-x * p[2]) 
p <- c(exp(1), 1/10)
K <- 10  # approx, a li'l' less
Xo <- 1 # approx, a li'l' less
```

```{r sdp-pars, dependson="stateeq"}
sigma_g <- 0.1
z_g <- function() rlnorm(1,0, sigma_g)
x_grid <- seq(0, 1.5 * K, length=50)
N <- 40
set.seed(123)
```

### Sample Data

```{r obs, dependson="sdp-pars"}
x <- numeric(N)
x[1] <- Xo
for(t in 1:(N-1))
  x[t+1] = z_g() * f(x[t], h=0, p=p)
qplot(1:N, x)
```

## Compute the posterior after marginalizing over $r$ and $\sigma$ parameters:

$$P(\beta | X) $$

```{r}
Mt <- function(t, beta) log(x[t+1]) - log(x[t]) + beta * x[t]
beta_grid = seq(1e-5, 2, by=1e-3)

P_B.X <- sapply(beta_grid, function(beta){
  Mt_vec = sapply(1:(N-1), Mt, beta)
  sum_of_squares <- sum(Mt_vec^2)
  square_of_sums <- sum(Mt_vec)^2
  0.5 ^ (N/2-1) * (sum_of_squares - square_of_sums/(N-1)) ^ (N/2-1) / gamma(N/2-1)
  })

qplot(beta_grid, -log(P_B.X))

beta_grid[which.min(P_B.X)]

```

# Estimating the Allen model:

$$X_{t+1} = X_t \exp\left(r \left(1 - \frac{X_t}{K} \right\left(\frac{X - C}{K}\right)\right) $$

First re-parameterize so we can isolate an additive parameter, using standard quadratic form,

$$X_{t+1} = X_t \exp(c + b X_t + a X_t^2) $$

Where $c = \tfrac{-rC}{K}$, $b = \tfrac{r}{K}\left(\tfrac{C}{K} + 1\right)$, and $a = \tfrac{r}{K^2}$.  Then after integrating out $c$ and $\sigma$ we have

```{r}
Mt <- function(t, a, b)  log(x[t+1]) - log(x[t]) - b * x[t] + a * x[t] ^ 2```

Choosing a sensible grid is a bit more tricky following the transformation, particularly $b$ can be negative or positive, and much larger in magnitude than any of the original parameters.  $a$ on the other hand, can be reasonably restricted:

```{r}
b_grid = seq(0, .1, length=100)
a_grid = seq(0, .1, length=100)
```

We define the probability as a function of the remaining two parameters,

```{r}
prob <- function(a, b){
  Mt_vec = sapply(1:(N-1), Mt, a, b)
  sum_of_squares <- sum(Mt_vec^2)
  square_of_sums <- sum(Mt_vec)^2
0.5 ^ (N/2-1) * (sum_of_squares - square_of_sums/(N-1)) ^ (N/2-1) / gamma(N/2-1)
}
```

and loop over the grid on each

```{r}
P_a_b.X <- sapply(a_grid, function(a)
                sapply(b_grid, function(b) prob(a, b)))

````

We summarize with a contour plot

```{r}
  df = melt(P_a_b.X)
names(df) = c("a", "b", "lik")
ggplot(df, aes(a_grid[a], b_grid[b], z=-log(lik))) + 
  stat_contour(aes(color=..level..), binwidth=1) +
  coord_cartesian(ylim=c(0, .025), xlim=c(0,.05))

```

```{r include=FALSE}
# Conditional Distributions
tmpA <- sapply(a_grid, prob, b=0.1)
tmpB <- sapply(b_grid, function(b) prob(a=0.01, b=b))

# Modes
a_grid[which.min(tmpA)]
b_grid[which.min(tmpB)]

qplot(a_grid, -log(tmpA))
qplot(b_grid, -log(tmpB))
```

# Estimating the Myers model on this data:

$$X_{t+1} = Z_t \frac{r X_t^{\theta}}{1 + X_t^{\theta} / K}$$

With $Z_t$ lognormal, unit mean, std $\sigma$.

```{r}
Mt <- function(t, theta, K) log(x[t+1]) - theta * log(x[t]) + log(1 + x[t] ^ theta / K) 
theta_grid = seq(1e-5, 5, length=100)
K_grid = seq(1e-5, 30, length=100)

prob <- function(theta, K){
  Mt_vec = sapply(1:(N-1), Mt, theta, K)
  sum_of_squares <- sum(Mt_vec^2)
  square_of_sums <- sum(Mt_vec)^2
0.5 ^ (N/2-1) * (sum_of_squares - square_of_sums/(N-1)) ^ (N/2-1) / gamma(N/2-1)
}

P_theta_K.X <- sapply(theta_grid, function(theta)
                sapply(K_grid, function(k) prob(theta, k)))

require(reshape2)
df = melt(P_theta_K.X)
names(df) = c("theta", "K", "lik")
ggplot(df, aes(theta_grid[theta], K_grid[K], z=-log(lik))) + stat_contour(aes(color=..level..), binwidth=3)

```

```{r include=FALSE}
# Conditional Distributions
tmpK <- sapply(K_grid, function(k) prob(1,k))
tmptheta <- sapply(theta_grid, prob, K=10)

# Modes
K_grid[which.min(tmpK)]
theta_grid[which.min(tmptheta)]

# Means
K_grid %*% -log(tmpK) / sum( -log(tmpK) )
theta_grid %*% -log(tmptheta) / sum( -log(tmptheta) )

qplot(theta_grid, -log(tmptheta))
qplot(K_grid, -log(tmpK))
```

## Reconstructing the joint probability distribution

$$P(r, \beta, \sigma | X) = \frac{P(X | r, \beta, \sigma) P(r) P(\beta) P(\sigma) }{\int_r \int_{\beta} \int_{\sigma P(X | r, \beta, \sigma) P(r) P(\beta) P(\sigma)}  $$

Having done the first two integrals analytically and then evaluating the remaining expression over the grid, we simply sum over the grids to complete our calculation of the normalization term from the denominator,

```{r}
N_ricker <- sum(P_B.X) * (beta_grid[2]-beta_grid[1])
```

We then have an analytic function for the numerator,

$$ P(X | r, \beta, \sigma) P(r) P(\beta) P(\sigma) =  \frac{1}{\sqrt{2 \pi \sigma^2}^{T-1}} \exp\left(\frac{\sum_t^{T-1} \left(\log X_{t+1} - \log X_t - \log r + \beta X_t\right)^2 }{2 \sigma^2}\right) \frac{1}{\sigma^2}$$

## Transforming back