### 2013-08-26 inst/examples/sensitivity-allen.Rmd

```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls())
source("~/.knitr_defaults.R")
opts_knit$set(upload.fun = socialR::flickr.url)
library(knitcitations)
library(nonparametricbayes) 
```

### Model and parameters

Various parameters defining noise dynamics, grid, and policy costs.

```{r sdp-pars}
sigma_g <- 0.05
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1
x_grid <- seq(0, 1.5 * 8, length=50)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 50  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
Xo <-  5.7 # observations start from
x0 <- 8 # simulation under policy starts from
Tobs <- 40
MaxT = 1000 # timeout for value iteration convergence
````

```{r stateeq-local}
f <- RickerAllee

```

```{r}
set.seed(12345)
seed <- round(runif(12) * 1e6)
seed
```

```{r parallel}
require(snowfall)
sfInit(parallel=TRUE, cpu=8)
sfLibrary(nonparametricbayes)
sfExportAll()
```

Replicate data-sets from the same model

```{r warning=TRUE}
yields <- sfSapply(seed, 
function(seed_i){
  set.seed(seed_i)

  p <- c(2 + rnorm(1, 0, .1), 8 + rnorm(1, 0, .1), 5 + rnorm(1, 0, .2))

  initial_data <- function(){
    x <- numeric(Tobs)
    x[1] <- Xo
    nz <- 1
    for(t in 1:(Tobs-1))
      x[t+1] = z_g() * f(x[t], h=0, p=p)

    X = c(rep(0,nz), pmax(rep(0,Tobs-1), x[1:(Tobs-1)]))
    Y = c(rep(0,nz), x[2:Tobs])

    out = list(X=X,Y=Y)
    if(x[Tobs-1] < x[1]){
      warning("Training population crashed, trying again")
      out <- initial_data()
    } 
    out
  }

  o <- initial_data()
  X <- o$X; Y <- o$Y
  ## @knitr gp-priors
  s2.p <- c(5,5)  
  d.p = c(10, 1/0.1)

  ## @knitr gp
  gp <- gp_mcmc(X, y=Y, n=1e5, s2.p = s2.p, d.p = d.p)
  gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)

  matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid) 
  opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)

  ## @knitr mle-opt
  matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
  opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

  OPT = data.frame(GP = opt_gp$D, True = opt_true$D)
  policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
  names(policies) <- c("stock", "method", "value")

  sims <- lapply(OPT, function(D){
    set.seed(1)
    lapply(1:100, function(i) 
      ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
    )
  })

  dat <- melt(sims, id=names(sims[[1]][[1]]))
  sims_data <- data.table(dat)
  setnames(sims_data, c("L1", "L2"), c("method", "reps")) 
  # Legend in original ordering please, not alphabetical: 
  sims_data$method = factor(sims_data$method, ordered=TRUE, levels=names(OPT))

  Profit <- sims_data[, sum(profit), by=c("reps", "method")]
  tmp <- dcast(Profit, reps ~ method)
  tmp <- tmp / tmp[,"True"]
  tmp <- melt(tmp[2:dim(tmp)[2]])
  tmp$value[tmp$variable == "GP"]

})
```

```{r}
yields_dat <- melt(yields)
names(yields_dat) <- c("replicate", "simulation", "value")
#class(yields_dat$simulation) = "factor"

col_metadata <- c(replicate = "id of individual replicates using identical settings, relative to the theoretical optimum",
  simulation = "id number for the simulation, which varies the simulated data used",
  value = "total profit achieved by specified run")
unit_metadata <- list(replicate = list(1:dim(yields)[1]), simulation = list(1:dim(yields)[2]), value = "currency (arbitrary)")
```

```{r}
head(yields_dat)
```

```{r}
ggplot(yields_dat) + geom_density(aes(value)) 
````

Compare over `r length(seed)` simulation configurations with different parameters and different training data, the GP does consistently well.  Only one case has relatively poor outcomes.

```{r}
ggplot(yields_dat) + geom_density(aes(value, color=as.factor(simulation), fill=as.factor(simulation)), alpha=.5)
````

Within a replicate number, outcomes not that varied over parameters.

```{r}
ggplot(yields_dat) + geom_density(aes(value, group=as.factor(replicate)), alpha=.7)
````

```{r}
save(list=ls(), file="sensitivity.rda")
```