### 2013-10-14 inst/doc/nonparametric-bayes.Rmd
writing

% Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach

<!-- Run R code for analysis, to be called by figures -->
```{r plot-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library(nonparametricbayes) 
opts_chunk$set(external=TRUE)
read_chunk("../examples/BUGS/external-chunks.R")

opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=1, 
               comment=NA, verbose=TRUE, fig.width=4, fig.height=3)

# Name the cache path and fig.path based on filename...
opts_chunk$set(fig.path = paste("figure/",
                                gsub(".Rmd", "", knitr:::knit_concord$get('infile')),
                                "-", sep=""),
               cache.path = paste("cache/", 
                                  gsub(".Rmd", "", knitr:::knit_concord$get('infile') ), 
                                "/", sep=""))

toggle = "markup" #"hide"  # toggles `results` chunk option

library(ggplot2) 
theme_set(theme_bw(base_size=12))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r posterior-mode}
```
```{r stateeq}
```
```{r sdp-pars, dependson="stateeq"}
```
```{r obs, dependson="sdp-pars"}
```

```{r mle, dependson="obs"}
```
```{r mle-output, dependson="mle", results=toggle}
```

```{r gp-priors}
```
```{r gp, dependson=c("gp-priors", "obs")}
```
```{r gp_traces_densities, dependson="gp"}
```
```{r gp-output, dependson="gp", results=toggle}
```

```{r jags-setup}
```
```{r common-priors}
```
```{r allen-model}
```
```{r allen-priors, dependson="common-priors"}
```
```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results=toggle}
```
```{r allen-traces, dependson="allen-mcmc"}
```
```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
```
```{r allen-output, dependson=c("posterior-mode", "allen-traces"), results=toggle}
```

```{r ricker-model}
```
```{r ricker-priors, dependson="common-priors"}
```
```{r ricker-mcmc, dependson="ricker-model", results=toggle}
```
```{r ricker-traces, dependson="ricker-mcmc"}
```
```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors")}
```
```{r ricker-output, dependson=c("posterior-mode", "ricker-traces"), results=toggle}
```

```{r myers-model}
```
```{r myers-priors}
```
```{r myers-mcmc, dependson="myers-model", results=toggle}
```
```{r myers-traces, dependson="myers-mcmc"}
```
```{r myers-posteriors, dependson="myers-traces"}
```
```{r myers-output, dependson=c("posterior-mode", "myers-traces"), results=toggle}
```

```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
```
```{r gp-opt, dependson="gp-output"}
```
```{r mle-opt, dependson="mle-output"}
```
```{r allen-opt, dependson="allen-output"}
```
```{r ricker-opt, dependson="ricker-output"}
```
```{r myers-opt, dependson="myers-output"}
```
```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
```
```{r sims, dependson="assemble-opt"}
```

```{r profits, dependson="sims", results=toggle}
```

```{r dic_calc, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output"), include=FALSE, echo=FALSE}
```

Abstract
========

Model uncertainty and limited data coverage are fundamental challenges to
robust ecosystem management.  These challenges are acutely highlighted
by concerns that many ecological systems may contain tipping points.
Before a collapse, we do not know where the tipping points lie, if the
exist at all.  Hence, we know neither a complete model of the system
dynamics nor do we have access to data in some large region of state-space
where such a tipping point might exist.  These two sources of uncertainty
frustrate state-of-the-art parametric approaches to decision theory
and optimal control.  I will illustrate how a non-parametric approach
using a Gaussian Process prior provides a more flexible representation
of this inherent uncertainty.  Consequently, we can adapt the Gaussian
Process prior to a stochastic dynamic programming framework in order to
make robust management predictions under both model and uncertainty and
limited data.

Introduction
============

Decision making under uncertainty is a ubiquitous challenge of natural 
resource management and conservation.

The sudden collapse of fisheries and other ecosystems is an increasingly
widespread phenomenon and a pressing concern for ecological management
and conservation.  Ecological dynamics are frequently complex and difficult to 
measure, making uncertainty in our understanding a prediction a persistent
challenge to effective management. 
Decision-theoretic approaches provide a framework to determine the best 
sequence of actions in face of uncertainty, but only when that uncertainty
can be meaningfully quantified [@Fischer2009].

<!-- Uncertainty outside the data without the correct model has not been handled. -->
<!-- __We don't have the model__ -->

Uncertainty enters the decision-making process at many levels: intrinsic 
stochasticity in biological processes, measurements, and implementation of 
policy [_e.g._ @Reed1979; @Clark1986; @Roughgarden1997; @Sethi2005], parameteric 
uncertainty [_e.g._ @Ludwig1982; @Hilborn1997; @McAllister1998; @Schapaugh2013],
and model or structural uncertainty [_e.g._ @Williams2001; @Cressie2009;  @Athanassoglou2012].
Of these, structural uncertainty incorporates the least a priori knowledge or 
assumptions and is generally the hardest to quantify. Typical approaches assume 
a weak notion of uncertainty where a correct or reasonable approximation of the dynamics
must be identified from among a handful of alternative models.  Here we consider an approach
that addresses uncertainty at each of these levels without assuming the dynamics follow
a particular (i.e. parametric) structure.

<!-- __We don't have the data where we need it__ -->
<!-- What do we call this?  Extrapolation uncertainty?  Pathological Uncertainty? -->

An additional source of uncertainty that has recieved less attention[^1]
arises when applying a dynamical model outside the range of data on
which it has been estimated.  This extrapolation uncertainty is felt
most keenly in decision-theoretic applications, as (a) exploring the
potential action space typically involves considering actions that
may move the system outside the range of observed behavior, and (b)
decision-theoretic alogrithms rely not only on reasonable estimates of
the expected outcomes, but depend on the weights given to all possible
outcomes [_e.g._ @Weitzman2013].

<!-- 
## Why we don't know the model
- Complex dynamics  [@Glaser2013](http://doi.org/10.1111/faf.12037 "Complex dynamics may limit prediction in marine fisheries")
- model choice and model averaging approaches
-->

<!--
## Why we don't have data where we need it
- Concerns of tipping points 
- Danger of learning 
-->

<!-- Perhaps this is a better-worded version of the above??

Often the set of feasible states is larger than the range of the observed
states.  Though most systems subject to some degree of stochasticity
can enter a state outside previously observed values by chance alone
(such as a particularly low or high nutrient year), this concern is
most keenly felt in the decision-theoretic context. If we are observing
the fluctuations of a given fish stock over many years under a fixed
harvesting pressure, we might develop and test a model that could
reasonably predict the frequency of a deviation of a given size, even
when such a deviation has not been previously observed.
(The same analogy might be made to predicting hurricanes or other
extreme weather events from historical data, which may be more accurately
predicted in a constant climate than in one being perturbed).

-->

<!-- Integrate this into the paragraph, rather than as a footnote?? Make more consise? -->

[^1]: The concept of adaptive probing is one area that has explicitly
addressed this kind of uncertainty, reaching rather opposite conclusions
than what we observe here. Adaptive probing strategies follow from "Dual
Control" or "Active Adaptive Management" approaches (e.g. @Ludwig1982)
that can trade off short term utility by choosing actions that can reduce
uncertainty.  Adaptive probing strategies arise when it is valuable to
intentionally force a system far from the observed values even when the
expected value such actions is low, as it provides much faster learning
and consequent reduction of model uncertainty that can allow greater value
to be derived later on.  For instance, @Ludwig1982 show that it may be
advantageous to fish an unexploited population very heavily at first
to obtain a better estimate of the recruitment rate.  This intuitive
strategy when a population is governed by a Ricker or Beverton-Holt-like
dynamic would clearly be disastrous if instead the dynamics contained
an unforeseen tipping point.  The best way to learn where the edge lies
may be to walk up to it, but it is also the most dangerous.

<!-- Does this belong?  Improve transitions -->

Concerns over the potential for tipping points in ecological dynamics
[@Scheffer2001] highlight the dangers of uncertainty in ecological
management and pose a substantial challenge to existing decision-theoretic
approaches [@Brozovic2011].  Because intervention is often too late 
after a tipping point has been crossed (but see @Hughes2013), management
is most often concerned with avoiding potentially catastrophic tipping
points before any data is available at or following a transition that
would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].

_Map_

Here we illustrate how a stochastic dynamic programming (SDP) algorithm
[@Mangel1988; @Possingham1997, Marescot2013] can be driven by the
predictions from a Bayesian non-parametric (BNP) approach [@Munch2005a].
This provides two distinct advantages compared with contemporary
approaches.  First, using a BNP sidesteps the need for an accurate
model-based description of the system dynamics.  Second, the BNP can
better reflect uncertainty that arises when extrapolating a model outside
of the data on which it was fit.  We illustrate that when the correct
model is not known, this latter feature is crucial to providing a robust
decision-theoretic approach in face of substantial structural uncertainty.

<!-- More Consise versions are now above 
- Parametric uncertainty

As the parameter values for these models must be estimated from limited data,  
there will always be some uncertainty associated with these values.  This uncertainty
further compounds the intrinsic variability introduced by demographic or environmental 
noise.  The degree of uncertainty in the parameter values can be inferred from the data
and reflected in the estimates of the transition probabilities [@Walters1982; @Mangel1988; @Mangel1997; @Schapaugh2013].

- Structural uncertainty

Estimates of parameter uncertainty are only as good as the parametric
models themselves.  Often we do not understand the system dynamics well
enough to know if a model provides a good approximation over the relevant
range of states and timescales (criteria that we loosely refer to as
defining the "right" or "true" model.)  So called structural or model
uncertainty is a more difficult problem than parametric uncertainty.
Typical solutions involve either model choice, model averaging, or
introducing yet greater model complexity of which others may be special
cases (model averaging being one such way to construct such a model)
[@Williams2001; @Athanassoglou2012; @Cressie2009].  Even setting aside
other computational and statistical concerns (e.g. [@Cressie2009]), these
approaches do not address our second concern - representing uncertainty
outside the observed data range.

- Limits of state space

Resource management and conservation planning seek to determine the
optimal set of feasible actions to maximize the value of some objectives
(see [@Halpern2012])  Process error, measurement error, implementation
error [@Reed1979, @Clark1986, @Roughgarden1996, @Sethi2005].  These
sources of stochasticity in turn mean that model parameters can only
be estimated approximately, requiring parametric uncertainty also be
considered [@Ludwig1982, ].
-->

<!-- necessary? --> <!-- should reflect the dual problem of extrapolation and model uncertainty better -->

This paper represents the first time the SDP decision-making framework has
been used without an a priori model of the underlying dynamics through
the use of the BNP approach.  In contrast to parametric models which
can only reflect uncertainty in parameter estimates, the BNP approach
provides a more state-space dependent representation of uncertainty.
This permits a much greater uncertainty far from the observed data than
near the observed data.  These features allow the GP-SDP approach to
find robust management solutions in face of limited data and without
knowledge of the correct model structure.

-  _Note on "not magic"_: honest uncertainty + SDP

The idea that any approach can perform well without either having to
know the model or have particularly good data should immediately draw
suspicion.  The reader must bear in mind that the strength of our approach
comes not from black-box predictive power from such limited information,
but rather, by providing a more honest expression of uncertainty outside
the observed data without sacrificing the predictive capacity near the
observed data. By coupling this more accurate description of what is known
and unknown to the decision-making under uncertainty framework provided
by stochastic dynamic programming, we are able to obtain more robust
management policies than with common parametric modeling approaches.

- Note on: Why fisheries

The economic value and ecological concern have made marine
fisheries the crucible for much of the founding work [@Gordon1954;
@Reed1979; @May1979; @Ludwig1982] in managing ecosystems under
uncertainty.  Global trends [@Worm2006] and controversy [@Hilborn2007;
@Worm2009] have made understanding these challenges all the more pressing.

<!--move to  much later -->

- _Note on comparing models_ (via value function rather than by "fit").

_Do we need this?_

The nature of decision-making problems provides a convenient way to compare 
models.  Rather than compare models in terms of best fit to data or fret over
the appropriate penalty for model complexity, model performance is defined 
in the concrete terms of the decision-maker's objective function, which we
will take as given. (Much argument can be made over the 'correct' objective
function, e.g. how to account for the social value of fish left in the sea
vs. the commercial value of fish harvested; see @Halpern2013 for further 
discussion of this issue.  Alternatively, we can always compare model performance
across multiple potential objective functions.)  The decision-maker
does not necessarily need a model that provides the best mechanistic understanding
or the best long-term outcome, but rather the one that best estimates the 
probabilities of being in different states as a result of the possible actions.

## Background on the Gaussian Process

- Background on non-parametric modeling.

Addressing the difficulty posed by extrapolation without knowing the
true model requires a nonparametric approach to model fitting: one
that does not assume a fixed structure but rather depends on the size
of the data (e.g. non-parametric regression or a Dirichlet process).
This established terminology is nevertheless unfortunate, as (a)
this approach still involves the estimation of parameters, and (b),
Statisticians use non-parametric to mean both this property (structure
is not fixed by the parameters) and an entirely different (and probably
more familiar) case in which the model does not assume any distribution
(e.g. non-parametric bootstrap, order statistics).  Some literature
thus uses the term semi-parametric, which merely adds ambiguity to
the confusion.

This non-parametric property -- having a structure explicitly dependent on
the data -- is precisely the property that makes this approach attractive
in face of the limited data sampling challenges discussed above.
Having fit a parametric model to some data, the model is completely
described by the values (or posterior distributions) of it's parameters.
The non-parametric model is not captured by its parameter values or
distributions alone. Either the model scales with the complexity of the data
on which it is estimated (e.g. nonparametric heirarchical approaches such
as the Dirchlet process) or the data points become themselves part of the model
specification, as in the nonparametric regression used here.  we shall see here.

- Definition

- Previous application

The use of Gaussian process (GP) regression (or "kriging" in the geospatial
literature) to formulate a predictive model is relatively new in the
context of modeling dynamical systems [@Kocijan2005], and was first introduced
in the context ecological modeling and fisheries management in @Munch2005.
An accessible and thorough introduction to the formulation and use of
GPs can be found in @Rasmussen2006.

- Why it is particularly suited to these two problems
- (Why this is a novel application thereof)

The essence of the GP approach can be captured in the
following thought experiment: An exhaustive parametric approach to the
challenge of structural uncertainty might proceed by writing down all
possible functional forms for the underlying dynamical system with all
possible parameter values for each form, and then consider searching
over this huge space to select the most likely model and parameters;
or using a Bayesian approach, assign priors to each of these possible
models and infer the posterior distribution of possible models. The
GP approach can be thought of as a computationally
efficient approximation to this approach. GPs represent
a large class of models that can be though of as capturing or reasonably
approximating the set of models in this collection.  By modeling at the
level of the process, rather than the level of parametric equation,
we can more concisely capture the possible behavior of these curves.
In place of a parametric model of the dynamical system, the GP 
approach postulates a prior distribution of (n-dimensional)
curves that can be though of as approximations to a range of possible
(parametric) models that might describe the data. The GP allows us
to consider probabilities on a large set of possible curves simultaneously.

The posterior distribution for the hyper-parameters of the Gaussian 
process model are estimated by Metropolis-Hastings algorithm, again with
details and code provided in the Appendix.  @Rasmussen2006 provides
an excellent general introduction to Gaussian Processes and @Munch2005 
first discusses their application in the context of population dynamics
models such as fisheries stock-recruitment relationships.

Approach and Methods
====================

### Summary of approach

### Statement of the optimal control problem

- Underlying model 
- Available data 
- Value function

For simplicity we assume profit is simply linear in the realized harvest (only
enforcing the restriction that harvest can not exceed available stock)

### Parametric models

- Statement of the models

We consider three candidate parametric models of the stock-recruitment
dynamics: The Ricker model, the Allen model [Allen 2005](), the Myers
model. The familiar Ricker model involves two parameters, corresponding
to a growth rate and a carrying capacity, and cannot support alternative
stable state dynamics (though as growth rate increases it exhibits a
periodic attractor that proceeds through period-doubling into chaos. We
will generally focus on dynamics below the chaotic threshold for
the purposes of this analysis.) The Allen model resembles the Ricker
dynamics with an added Allee effect parameter [Courchamp](), below
which the population cannot persist.  The Myers model also has three
parameters and contains an Allee threshold, but has compensatory rather
than over-compensatory density dependence (resembling a Beverton-Holt
curve rather than a Ricker curve at high densities.)

We assume multiplicative log-normal noise perturbs the growth predicted 
by the each of the deterministic model skeletons described above. This 
introduces one additional parameter $\sigma$ that must be estimated by each
model.

<!-- equations just in appendix? -->

As we simulate training data from the Allen model (ref section), we will
refer to this as the structurally correct model.  The Ricker model is
thus a reasonable approximation of these dynamics far from the Allee
threshold (but lacks threshold dynamics), while the Myers model shares
the essential feature of a threshold but differs in the structure. Thus
we have three potential parametric models of the stock dynamics.

- Bayesian inference of parametric models

We infer posterior distributions for the parameters of each model
in a Bayesian context using Gibbs sampling (implemented in R [@RTeam]
using jags, [@R2jags]).  We choose uninformative uniform priors for all
parameters (See Appendix, Figures S1-S3, and Table S1, and the R code
provided). One-step-ahead predictions of these model fits are shown in
Figure 1.

- SDP via parametric models

An optimal policy function is then inferred through stochastic dynamic
programming for each model given the posterior distributions of the
parameter estimates.  This policy maximizes the expectation of the value
function integrated over the parameter uncertainty. (code implementing
this algorithm provided in the Appendix).

### The Gaussian Process model

- Statement of model

... more on GP ... [Munch 2005]()

We also estimate a simple Gaussian Process defined by
a radial basis function kernel of two parameters: $\ell$, which gives
the characteristic length-scale over which correlation between two 
points (e.g. any two points $X_t, X_{t+1}$, and $X_{t+\tau}, X_{t+1+\tau}$)
in state-space decays, and $\sigma$, which gives the scale of the 
process noise by which observations $Y_{t+1}$ may differ from their
predicted values $X_{t+1}$ given an observation of the previous state,
$X_t$.

- Inference of the model

Also unlike parametric models, this posterior distribution is still
conditional on the training data. As such, the uncertainty near the
observed data.

We use a Metropolis-Hastings Markov Chain Monte Carlo to infer posterior
distributions of the two parameters of the GP (Figure S4, code in
appendix), under weakly informative Gaussian priors (see parameters in
table S5). As the posterior distributions differ substantially from the
priors (Figure S4), we can be assured that most of the information in
the posterior comes from the data rather than the prior belief.

- SDP via the model

Though we are unaware of prior application of this type, it is reasonably
straight-forward to adapt the Gaussian Process for Stochastic Dynamic
Programming.  Recall that unlike the parametric models the Gaussian
process with fixed parameters already predicts a distribution of
curves rather than a single curve. We must first integrate over 
this distribution of curves given a sampling of parameter values drawn
from the posterior distribution of the two GP parameters, before
integrating over the posterior of those parameters themselves.

Results
=======

<!-- 
### Figure 1: Fitted Models

- All models fit the data quite well
- Information criteria would pick the simple, incorrect model.
-->

```{r Figureb_posteriors, dependson=c("assemble-models", "par-fns"), include=TRUE, echo=FALSE, fig.cap="Points show the training data of stock-size over time.  Curves show the posterior step-ahead predictions based on each of the estimated models.", fig.width=8, fig.height=6}
```

All models fit the observed data rather closely and with relatively small uncertainty, as illustrated in the posterior predictive curves in Figure 1.  Figure 1 shows the training data of stock sizes observed over time as points, overlaid with the step-ahead predictions of each estimated model using the parameters sampled from their posterior distributions.  Each model manages to fit the observed data rather closely. Compared to the expected value of the true model most estimates appear to overfit, predicting fluctuations that are actually due purely to stochasticity in growth rate.  Model-choice criteria shown in Table 1 penalize more complex models and show a slight preference for the simpler Ricker model over the more complicated alternate stable state models (Allen and Myers).  Details on MCMC estimates for each model, traces, and posterior distributions can be found in the appendix.

```{r deviances, dependson=c("dic_calc"), include=TRUE, results="asis", echo=FALSE}
```

<!-- 
### Figure 2 
- Data comes from limited region of state-space 
- (Should really show uncertainty of all models here.  Capture the forecast uncertainty several steps down the road?)   
-->

```{r statespace_posteriors, dependson=c("assemble-models"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.  Graph shows the expected value for the function $f$ under each model.  Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.  The training data is also shown as black points.  The GP is conditioned on (0,0), shown as a pseudo-data point.", include=TRUE, echo=FALSE, fig.width=8, fig.height=6}
```

The mean inferred state space dynamics of each model
relative to the true model used to generate the data is shown in Figure 2, 
predicting the relationship between observed stock size (x-axis) to the stock size
after recruitment the following year.  Note that in contrast to the  other models shown, the expected 
Gaussian process corresponds to a distribution of curves - as indicated
by the gray band - which itself has a mean shown in black. Parameter uncertainty
(not shown) spreads out the estimates further.  
The observed data from which each model is estimated is also shown.  The observations come
from only a limited region of state space corresponding to unharvested
or weakly harvested system.  No observations occur at the theoretical
optimum harvest rate or near the tipping point.

```{r out_of_sample_predictions, include=TRUE, echo=FALSE, fig.width=8, fig.height=6}
```

<!--
### Figure 3: Inferred Policies

- Inferred policies differ substantially among models
- The structurally correct model and the GP are close to the true model
- alternatives are not close
-->

```{r Figure2, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.  Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", include=TRUE, echo=FALSE}
```

Despite the similarities in model fits to the observed data, the policies
inferred under each model differ widely, as shown in Figure 3.  
Policies are shown in terms of target
escapement, $S_t$.  Under models such as this a constant escapement
policy is expected to be optimal [@Reed1979], whereby population levels
below a certain size $S$ are unharvested, while above that size the harvest
strategy aims to return the population to $S$, resulting in the hockey-stick
shaped policies shown.  Only the structurally correct model (Allen model) and the GP 
produce policies close to the true optimum policy (where both the underlying 
model structure and parameter values are known without error).

```{r Figure3, dependson=c("sim"), fig.cap="Gaussian process inference outperforms parametric estimates. Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.",  include=TRUE, echo=FALSE}
```

The consequences of managing 100 replicate realizations of the 
simulated fishery under each of the policies estimated is shown in Figure 4.  As expected
from the policy curves, the structurally correct model under-harvests,
leaving the stock to vary around it's un-fished optimum.  The structurally
incorrect Ricker model over-harvests the population passed
the tipping point consistently, resulting in the immediate crash of the stock and 
thus derives minimal profits.

These results are robust across a range of stochastic realizations, models, and parameter values.  
The results across this range can most easily be compared 
by using the relative differences in net present value realized by each of the model,
as shown in Figure 5.  The BNP-SDP approach most consistently realizes a value 
close to the optimal solution, and importantly avoids ever driving the system across
the tipping point, which results in the near-zero value cases in the parametric models.

```{r Figure4, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates. For each data set, the three models are estimated as described above. Values plotted are the averages of a given policy over 100 replicate simulations. Details and code provided in the supplement.",  include=TRUE, echo=FALSE}

```

Discussion 
==========

- All models are "good fits" to the originally observed data. 
- (Simple model choice immediately leads us astray)

<!--

Though simple mechanistically motivated models offer the greatest potential 
to increase our basic understanding of ecological processes [@Cuddington2013; @Geritz2012], 
such models can be not only inaccurate but misleading when relied upon in a
quantitative decision making framework.  
-->

1. We do not know what the correct models are for ecological systems.
1. We have limited data from which to estimate the model -- in particular,
   such models may be misleading in predicting the probability of outcomes
   outside the training data.

These aspects are common to many conservation decision making problems, which thus merit
greater use of non-parametric approaches that can best take advantage of them.

### Traditional model-choice approaches can be positively misleading.

These results illustrate that model-choice approaches would be positively
misleading -- supporting simpler models that cannot express tipping point
dynamics merely on account of them being similar.  As the data shown
comes only from the basin of attraction near the unfished equilibrium,
near which all of the models are approximately linear and approximately
identical.

Model choice approaches trade off model complexity and fit to the data. 
When the data come from a limited region of state-space -- as is necessarily 
the case whenever there is a potential concern about tipping point dynamics --
simpler models can fit just as well and will tend to outperform more complex 
ones.  This approach would be appropriate when the dynamics can be expected 
to remain in the region of the training data; for instance, if we only 
considered the forecasting accuracy of the unfished population dynamics under
each model.

In contrast, the decision-maker's problem of setting appropriate harvest levels
cannot exclude regions of state-space outside the observed range when integrating
over all possible decisions to find the optimal choice.  Such problems are not 
constrained to fisheries management but ubiquitous across ecological decision-making
and conservation where the greatest concerns involve entering previously unobserved
regions of state-space -- whether that is the collapse of a fishery, the spread
of an invasive, or the loss of habitat.

### BNP-SDP expresses larger uncertainty in regions where the data are poor

The parametric models perform worst when they propose a management strategy
outside the range of the observed data. The non-parametric Bayesian approach, 
in contrast, allows a predictive model that expresses a great deal of uncertainty
about the probable dynamics outside the observed range, while retaining very
good predictive accuracy in the range observed.  The management policy 
dictated by the GP balance this uncertainty against
the immediate value of the harvest, and act to stabilize the population 
dynamics in a region of state space in which the predictions can be 
reliably reflected by the data.

### BNP-SDP has good predictive accuracy where data are good

While expressing larger uncertainty outside the observed data, the GP
can also provide a better fit with smaller uncertainty inside the range
of the observed data. This arises from the greater flexibility of the 
Gaussian process, which describes a large family of possible curves.
Despite this flexibility, the GP can be described in relatively few 
parameters and is thus far less likely to overfit.

Future directions
-----------------

### Higher dimensions

In this simulated example, the underlying
dynamics are truly governed by a simple parametric model, allowing
the parametric approaches to be more accurate.  Similarly, because the
dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
than other attractors such as limit-cycles resulting in oscillations),
the training data provides relatively limited information about the
dynamics.  For these reasons, we anticipate that in higher-dimensional
examples characteristic of ecosystem management problems that the machine
learning approach will prove even more valuable.

### Real-time learning

In our treatment here we have ignored the possibility of learning during the 
management phase, in which the additional observations of the stock size could
potentially improve parameter estimates.  While we intend to address this 
possibility in future work in the context of these non-parametric models,
we have not addressed it here for pedagogical reasons. In the context presented
here, it is clear that the differences in performance arise from differences
in the uncertainty inherent in the model formulations, rather than from 
differing abilities to learn.  Because we consider a threshold system, 
online learning would not change this generic feature of a lack of data in a
certain range of the state space which is better captured by the Gaussian process.

Acknowledgments
===============

This work was partially supported by the Center for Stock Assessment
Research, a partnership between the University of California Santa Cruz
and the Fisheries Ecology Division, Southwest Fisheries Science Center,
Santa Cruz, CA and by NSF grant EF-0924195 to MM and NSF grant DBI-1306697
to CB.

Appendix
========

## Model definitions and estimation

Equation S1: Ricker model.

$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } $$

Figure S1: Ricker model: prior and posterior distributions for parameter estimates.

```{r, echo=FALSE, include=TRUE, fig.height=6}
plot_ricker_traces
```

```{r, echo=FALSE, include=TRUE}
plot_ricker_posteriors
```
Table S1: Parameterization of the priors

```{r echo=FALSE, results = "asis", include = TRUE}
xtable::xtable(data.frame(
  parameter = c("r0", "K", "sigma"),
  lower_bound = c(r0_prior_p[1], K_prior_p[1], stdQ_prior_p[1]),
  upper_bound = c(r0_prior_p[2], K_prior_p[2], stdQ_prior_p[2])))
```

$$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} $$

Eq S2: Myers model 
Figure S2: Myers model: Traces, prior and posterior distributions for parameter estimates.

```{r, echo=FALSE, include=TRUE, fig.height=6}
plot_myers_traces
```

```{r, echo=FALSE, include=TRUE}
plot_myers_posteriors
```

Table S2: Parameterization of the priors
```{r echo=FALSE, results="asis", include = TRUE}
xtable::xtable(data.frame(parameter = c("r0", "K", "theta", "sigma"),
           lower_bound = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
           upper_bound = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2])))
```

Eq S3: Allen model

$$f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} $$

Figure S3: Allen model: prior and posterior distributions for parameter estimates.

```{r, echo=FALSE, include=TRUE, fig.height=6}
plot_allen_traces
```

```{r, echo=FALSE, include=TRUE}
plot_allen_posteriors
```

Table S3: Parameterization of the priors

```{r echo=FALSE, results = "asis", include=TRUE}
xtable::xtable(data.frame(parameter = c("r0", "K", "theta", "sigma"),
           lower_bound = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
           upper_bound = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2])))
```

Eq S4: GP model 
Figure S4: GP model: prior and posterior distributions for parameter estimates.

```{r echo=FALSE, include=TRUE}
gp_assessment_plots
```

Table S4: Parameterization of the priors

## Optimal Control Problem

We seek the harvest policy $h(x)$ that maximizes:

$$ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  $$

subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and the state
equation

$$X_{t+1} = Z_t f(S_t)  $$
$$S_t = X_t - h_t $$

Where $Z_t$ is multiplicative noise function with mean 1, representing
stochastic growth. We will consider log-normal noise with shape parameter
$\sigma_g$.

Form this we can write down the Bellman recursion as:

$$V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)$$

For simplicity we assume profit is simply linear in the realized harvest (only
enforcing the restriction that harvest can not exceed available stock), $\Pi(h,x) = min(h,x)$.

### Pseudocode for the Bellman iteration

```r
 V1 <- sapply(1:length(h_grid), function(h){
      delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
    })
    # find havest, h that gives the maximum value
    out <- sapply(1:gridsize, function(j){
      value <- max(V1[j,], na.rm = T) # each col is a diff h, max over these
      index <- which.max(V1[j,])  # store index so we can recover h's 
      c(value, index) # returns both profit value & index of optimal h.  
    })
    # Sets V[t+1] = max_h V[t] at each possible state value, x
    V <- out[1,]                        # The new value-to-go
    D[,OptTime-time+1] <- out[2,]       # The index positions
```

### Training data

Eacho of our models $f(S_t)$ must be estimated from training data, which
we simulate from the Allen model with parameters $r = $ ` r p[1]`, 
$K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g` 
for $T=$ `r Tobs` timesteps, starting at initial condition $X_0 = $ `r Xo`. 
The training data can be seen in Figure 1.

### Sensitivity Analysis