### 2013-11-18 inst/doc/nonparametric-bayes.tex
writing, fixing references, figure tweaks

\documentclass[author-year, review]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}
\usepackage{lineno} % add 
%\linenumbers % turns line numbering on 
\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%% Redefines the elsarticle footer
\makeatletter
\def\ps@pprintTitle{%
 \let\@oddhead\@empty
 \let\@evenhead\@empty
 \def\@oddfoot{\it \hfill\today}%
 \let\@evenfoot\@oddfoot}
\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

\begin{document}
\begin{frontmatter}
  \title{Avoiding tipping points in the management of ecological systems: a
non-parametric Bayesian approach}
  \author[cstar]{Carl Boettiger\corref{cor1}}
  \cortext[cor1]{Corresponding author}
  \ead{cboettig@ucsc.edu}
  \author[cstar]{Marc Mangel}
  \author[noaa]{Stephan B. Munch}
  \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA 95064, USA}
  \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
 \end{frontmatter}

\section{Abstract}\label{abstract}

Model uncertainty and limited data coverage are fundamental challenges
to robust ecosystem management. These challenges are acutely highlighted
by concerns that many ecological systems may contain tipping points.
Before a collapse, we do not know where the tipping points lie, if the
exist at all. Hence, we know neither a complete model of the system
dynamics nor do we have access to data in some large region of
state-space where such a tipping point might exist. These two sources of
uncertainty frustrate state-of-the-art parametric approaches to decision
theory and optimal control. I will illustrate how a non-parametric
approach using a Gaussian Process prior provides a more flexible
representation of this inherent uncertainty. Consequently, we can adapt
the Gaussian Process prior to a stochastic dynamic programming framework
in order to make robust management predictions under both model and
uncertainty and limited data.

\section{Introduction}\label{introduction}

Decision making under uncertainty is a ubiquitous challenge of natural
resource management and conservation. Ecological dynamics are frequently
complex and difficult to measure, making uncertainty in our
understanding a prediction a persistent challenge to effective
management. Decision-theoretic approaches provide a framework to
determine the best sequence of actions in face of uncertainty, but only
when that uncertainty can be meaningfully quantified (Fischer et al.
2009). The sudden collapse of fisheries and other ecosystems has
increasingly emphasized the difficulties of formulating even
qualitatively correct models of the underlying processes.

Uncertainty enters the decision-making process at many levels: intrinsic
stochasticity in biological processes, measurements, and implementation
of policy (\emph{e.g.} Reed 1979; C. W. Clark and Kirkwood 1986;
Roughgarden and Smith 1996; Sethi et al. 2005), parameteric uncertainty
(\emph{e.g.} Ludwig and Walters 1982; Hilborn and Mangel 1997;
McAllister 1998; Schapaugh and Tyre 2013), and model or structural
uncertainty (\emph{e.g.} B. K. Williams 2001; Cressie et al. 2009;
Athanassoglou and Xepapadeas 2012). Of these, structural uncertainty
incorporates the least a priori knowledge or assumptions and is
generally the hardest to quantify. Typical approaches assume a weak
notion of model uncertainty in which the correct model (or reasonable
approximation) of the dynamics must be identified from among a handful
of alternative models. Here we consider an approach that addresses
uncertainty at each of these levels without assuming the dynamics follow
a particular (i.e.~parametric) structure.

Model uncertainty is particularly insidious when model predictions must
be made outside of the range of data on which the model was estimated.

An additional source of uncertainty that has recieved less
attention{[}\^{}1{]} arises when applying a dynamical model outside the
range of data on which it has been estimated. This extrapolation
uncertainty is felt most keenly in decision-theoretic applications, as
(a) exploring the potential action space typically involves considering
actions that may move the system outside the range of observed behavior,
and (b) decision-theoretic alogrithms rely not only on reasonable
estimates of the expected outcomes, but depend on the weights given to
all possible outcomes (\emph{e.g.} Weitzman 2013).

Concerns over the potential for tipping points in ecological dynamics
(Scheffer et al. 2001) highlight the dangers of uncertainty in
ecological management and pose a substantial challenge to existing
decision-theoretic approaches (Brozović and Schlenker 2011). Because
intervention is often too late after a tipping point has been crossed
(but see Hughes et al. (2013)), management is most often concerned with
avoiding potentially catastrophic tipping points before any data is
available at or following a transition that would more clearly reveal
these regime shift dynamics (e.g. Bestelmeyer et al. 2012).

Here we illustrate how a stochastic dynamic programming (SDP) algorithm
(Mangel and Clark 1988; Marescot et al. 2013) can be driven by the
predictions from a Bayesian non-parametric (BNP) approach (Munch,
Kottas, and Mangel 2005). This provides two distinct advantages compared
with contemporary approaches. First, using a BNP sidesteps the need for
an accurate model-based description of the system dynamics. Second, the
BNP can better reflect uncertainty that arises when extrapolating a
model outside of the data on which it was fit. We illustrate that when
the correct model is not known, this latter feature is crucial to
providing a robust decision-theoretic approach in face of substantial
structural uncertainty.

This paper represents the first time the SDP decision-making framework
has been used without an a priori model of the underlying dynamics
through the use of the BNP approach. In contrast to parametric models
which can only reflect uncertainty in parameter estimates, the BNP
approach provides a more state-space dependent representation of
uncertainty. This permits a much greater uncertainty far from the
observed data than near the observed data. These features allow the
GP-SDP approach to find robust management solutions in face of limited
data and without knowledge of the correct model structure.

The idea that any approach can perform well without either having to
know the model or have particularly good data should immediately draw
suspicion. The reader must bear in mind that the strength of our
approach comes not from black-box predictive power from such limited
information, but rather, by providing a more honest expression of
uncertainty outside the observed data without sacrificing the predictive
capacity near the observed data. By coupling this more accurate
description of what is known and unknown to the decision-making under
uncertainty framework provided by stochastic dynamic programming, we are
able to obtain more robust management policies than with common
parametric modeling approaches.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Note on: Why fisheries
\end{itemize}

The economic value and ecological concern have made marine fisheries the
crucible for much of the founding work (H. S. Gordon 1954; Reed 1979;
May et al. 1979; Ludwig and Walters 1982) in managing ecosystems under
uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
2007; Worm et al. 2009) have made understanding these challenges all the
more pressing.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \emph{Note on comparing models} (via value function rather than by
  ``fit'').
\end{itemize}

The nature of decision-making problems provides a convenient way to
compare models. Rather than compare models in terms of best fit to data
or fret over the appropriate penalty for model complexity, model
performance is defined in the concrete terms of the decision-maker's
objective function, which we will take as given. (Much argument can be
made over the `correct' objective function, e.g.~how to account for the
social value of fish left in the sea vs.~the commercial value of fish
harvested; see Halpern et al. (2013) for further discussion of this
issue. Alternatively, we can always compare model performance across
multiple potential objective functions.) The decision-maker does not
necessarily need a model that provides the best mechanistic
understanding or the best long-term outcome, but rather the one that
best estimates the probabilities of being in different states as a
result of the possible actions.

\subsection{Background on the Gaussian
Process}\label{background-on-the-gaussian-process}

Addressing the difficulty posed by extrapolation without knowing the
true model requires a nonparametric approach to model fitting: one that
does not assume a fixed structure but rather depends on the size of the
data (e.g.~non-parametric regression or a Dirichlet process). This
established terminology is nevertheless unfortunate, as (a) this
approach still involves the estimation of parameters, and (b),
Statisticians use non-parametric to mean both this property (structure
is not fixed by the parameters) and an entirely different (and probably
more familiar) case in which the model does not assume any distribution
(e.g.~non-parametric bootstrap, order statistics). Some literature thus
uses the term semi-parametric, which merely adds ambiguity to the
confusion.

This non-parametric property -- having a structure explicitly dependent
on the data -- is precisely the property that makes this approach
attractive in face of the limited data sampling challenges discussed
above. Having fit a parametric model to some data, the model is
completely described by the values (or posterior distributions) of it's
parameters. The non-parametric model is not captured by its parameter
values or distributions alone. Either the model scales with the
complexity of the data on which it is estimated (e.g.~nonparametric
heirarchical approaches such as the Dirchlet process) or the data points
become themselves part of the model specification, as in the
nonparametric regression used here. we shall see here.

The use of Gaussian process (GP) regression (or ``kriging'' in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems (Kocijan et al. 2005), and
was first introduced in the context ecological modeling and fisheries
management in Munch et al. (2005). An accessible and thorough
introduction to the formulation and use of GPs can be found in Rasmussen
and Williams (2006).

The essence of the GP approach can be captured in the following thought
experiment: An exhaustive parametric approach to the challenge of
structural uncertainty might proceed by writing down all possible
functional forms for the underlying dynamical system with all possible
parameter values for each form, and then consider searching over this
huge space to select the most likely model and parameters; or using a
Bayesian approach, assign priors to each of these possible models and
infer the posterior distribution of possible models. The GP approach can
be thought of as a computationally efficient approximation to this
approach. GPs represent a large class of models that can be though of as
capturing or reasonably approximating the set of models in this
collection. By modeling at the level of the process, rather than the
level of parametric equation, we can more concisely capture the possible
behavior of these curves. In place of a parametric model of the
dynamical system, the GP approach postulates a prior distribution of
(n-dimensional) curves that can be though of as approximations to a
range of possible (parametric) models that might describe the data. The
GP allows us to consider probabilities on a large set of possible curves
simultaneously.

The posterior distribution for the hyper-parameters of the Gaussian
process model are estimated by Metropolis-Hastings algorithm, again with
details and code provided in the Appendix. Rasmussen and Williams (2006)
provides an excellent general introduction to Gaussian Processes and
Munch et al. (2005) first discusses their application in the context of
population dynamics models such as fisheries stock-recruitment
relationships.

\section{Approach and Methods}\label{approach-and-methods}

\subsubsection{Statement of the optimal control
problem}\label{statement-of-the-optimal-control-problem}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Underlying model
\item
  Available data
\end{itemize}

For simplicity we assume profit is simply linear in the realized harvest
(only enforcing the restriction that harvest can not exceed available
stock)

\subsubsection{Parametric models}\label{parametric-models}

We consider three candidate parametric models of the stock-recruitment
dynamics: The Ricker model, the Allen model \href{}{Allen 2005}, the
Myers model. The familiar Ricker model involves two parameters,
corresponding to a growth rate and a carrying capacity, and cannot
support alternative stable state dynamics (though as growth rate
increases it exhibits a periodic attractor that proceeds through
period-doubling into chaos. We will generally focus on dynamics below
the chaotic threshold for the purposes of this analysis.) The Allen
model resembles the Ricker dynamics with an added Allee effect parameter
\href{}{Courchamp}, below which the population cannot persist. The Myers
model also has three parameters and contains an Allee threshold, but has
compensatory rather than over-compensatory density dependence
(resembling a Beverton-Holt curve rather than a Ricker curve at high
densities.)

We assume multiplicative log-normal noise perturbs the growth predicted
by the each of the deterministic model skeletons described above. This
introduces one additional parameter $\sigma$ that must be estimated by
each model.

As we simulate training data from the Allen model (ref section), we will
refer to this as the structurally correct model. The Ricker model is
thus a reasonable approximation of these dynamics far from the Allee
threshold (but lacks threshold dynamics), while the Myers model shares
the essential feature of a threshold but differs in the structure. Thus
we have three potential parametric models of the stock dynamics.

We infer posterior distributions for the parameters of each model in a
Bayesian context using Gibbs sampling (implemented in R (Team 2013)
using jags, (Su and Yajima 2013)). We choose uninformative uniform
priors for all parameters (See Appendix, Figures S1-S3, and Table S1,
and the R code provided). One-step-ahead predictions of these model fits
are shown in Figure 1.

An optimal policy function is then inferred through stochastic dynamic
programming for each model given the posterior distributions of the
parameter estimates. This policy maximizes the expectation of the value
function integrated over the parameter uncertainty. (code implementing
this algorithm provided in the Appendix).

\subsubsection{The Gaussian Process
model}\label{the-gaussian-process-model}

We also estimate a simple Gaussian Process defined by a radial basis
function kernel of two parameters: $\ell$, which gives the
characteristic length-scale over which correlation between two points in
state-space decays, and $\sigma$, which gives the scale of the process
noise by which observations $Y_{t+1}$ may differ from their predicted
values $X_{t+1}$ given an observation of the previous state, $X_t$.
Munch, Kottas, and Mangel (2005) gives an accessible introduction to the
use of Gaussian Processes in providing a Bayesian nonparametric
description of the stock-recruitment relationship.

In contrast to the parametric models, this posterior distribution is
still conditional on the training data. As such, the uncertainty near
the observed data.

We use a Metropolis-Hastings Markov Chain Monte Carlo to infer posterior
distributions of the two parameters of the GP (Figure S4, code in
appendix), under weakly informative Gaussian priors (see parameters in
table S5). As the posterior distributions differ substantially from the
priors (Figure S4), we can be assured that most of the information in
the posterior comes from the data rather than the prior belief.

Though we are unaware of prior application of this type, it is
reasonably straight-forward to adapt the Gaussian Process for Stochastic
Dynamic Programming. Recall that unlike the parametric models the
Gaussian process with fixed parameters already predicts a distribution
of curves rather than a single curve. We must first integrate over this
distribution of curves given a sampling of parameter values drawn from
the posterior distribution of the two GP parameters, before integrating
over the posterior of those parameters themselves.

\section{Results}\label{results}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-Figureb_posteriors.pdf}
\caption{Points show the training data of stock-size over time. Curves
show the posterior step-ahead predictions based on each of the estimated
models.}
\end{figure}

All models fit the observed data rather closely and with relatively
small uncertainty, as illustrated in the posterior predictive curves in
Figure 1. Figure 1 shows the training data of stock sizes observed over
time as points, overlaid with the step-ahead predictions of each
estimated model using the parameters sampled from their posterior
distributions. Each model manages to fit the observed data rather
closely. Compared to the expected value of the true model most estimates
appear to overfit, predicting fluctuations that are actually due purely
to stochasticity in growth rate. Model-choice criteria shown in Table 1
penalize more complex models and show a slight preference for the
simpler Ricker model over the more complicated alternate stable state
models (Allen and Myers). Details on MCMC estimates for each model,
traces, and posterior distributions can be found in the appendix.

\% latex table generated in R 3.0.2 by xtable 1.7-1 package \% Thu Nov
14 14:10:28 2013

\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & Allen & Ricker & Myers \\ 
  \hline
DIC & 50.14 & 49.45 & 50.61 \\ 
  AIC & -24.60 & -30.07 & -27.19 \\ 
  BIC & -17.85 & -25.00 & -20.44 \\ 
   \hline
\end{tabular}
\end{table}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-statespace_posteriors.pdf}
\caption{Graph of the inferred Gaussian process compared to the true
process and maximum-likelihood estimated process. Graph shows the
expected value for the function $f$ under each model. Two standard
deviations from the estimated Gaussian process covariance with (light
grey) and without (darker grey) measurement error are also shown. The
training data is also shown as black points. The GP is conditioned on
(0,0), shown as a pseudo-data point.}
\end{figure}

The mean inferred state space dynamics of each model relative to the
true model used to generate the data is shown in Figure 2, predicting
the relationship between observed stock size (x-axis) to the stock size
after recruitment the following year. Note that in contrast to the other
models shown, the expected Gaussian process corresponds to a
distribution of curves - as indicated by the gray band - which itself
has a mean shown in black. Parameter uncertainty (not shown) spreads out
the estimates further. The observed data from which each model is
estimated is also shown. The observations come from only a limited
region of state space corresponding to unharvested or weakly harvested
system. No observations occur at the theoretical optimum harvest rate or
near the tipping point.

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-out_of_sample_predictions.pdf}
\caption{plot of chunk out\_of\_sample\_predictions}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-Figure2.pdf}
\caption{The steady-state optimal policy (infinite boundary) calculated
under each model. Policies are shown in terms of target escapement,
$S_t$, as under models such as this a constant escapement policy is
expected to be optimal (Reed 1979).}
\end{figure}

Despite the similarities in model fits to the observed data, the
policies inferred under each model differ widely, as shown in Figure 3.
Policies are shown in terms of target escapement, $S_t$. Under models
such as this a constant escapement policy is expected to be optimal
(Reed 1979), whereby population levels below a certain size $S$ are
unharvested, while above that size the harvest strategy aims to return
the population to $S$, resulting in the hockey-stick shaped policies
shown. Only the structurally correct model (Allen model) and the GP
produce policies close to the true optimum policy (where both the
underlying model structure and parameter values are known without
error).

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-Figure3.pdf}
\caption{Gaussian process inference outperforms parametric estimates.
Shown are 100 replicate simulations of the stock dynamics (eq 1) under
the policies derived from each of the estimated models, as well as the
policy based on the exact underlying model.}
\end{figure}

The consequences of managing 100 replicate realizations of the simulated
fishery under each of the policies estimated is shown in Figure 4. As
expected from the policy curves, the structurally correct model
under-harvests, leaving the stock to vary around it's un-fished optimum.
The structurally incorrect Ricker model over-harvests the population
passed the tipping point consistently, resulting in the immediate crash
of the stock and thus derives minimal profits.

These results are robust across a range of stochastic realizations,
models, and parameter values. The results across this range can most
easily be compared by using the relative differences in net present
value realized by each of the model, as shown in Figure 5. The BNP-SDP
approach most consistently realizes a value close to the optimal
solution, and importantly avoids ever driving the system across the
tipping point, which results in the near-zero value cases in the
parametric models.

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-Figure4.pdf}
\caption{Histograms of the realized net present value of the fishery
over a range of simulated data and resulting parameter estimates. For
each data set, the three models are estimated as described above. Values
plotted are the averages of a given policy over 100 replicate
simulations. Details and code provided in the supplement.}
\end{figure}

\section{Discussion}\label{discussion}

In any modeling effort, models must be chosen for the task at
hand.\\Though simple mechanistically motivated models offer the greatest
potential to increase our basic understanding of ecological processes
(Cuddington, Fortin, and Gerber 2013; Geritz and Kisdi 2012), such
models can be not only inaccurate but misleading when relied upon in a
quantitative decision making framework.\\In this paper we have tackled
two aspects of uncertainty that are both common to many ecological
desicion-making problems and fundamentally challenging to existing
approaches which largely rely on parametric models:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  We do not know what the correct models are for ecological systems.
\item
  We have limited data from which to estimate the model -- in
  particular, such models may be misleading in predicting the
  probability of outcomes outside the training data.
\end{enumerate}

We have illustrated how the use of non-parametric approaches can provide
more reliable solutions in the sequential decision-making problem.

\subsubsection{Traditional model-choice approaches can be positively
misleading.}\label{traditional-model-choice-approaches-can-be-positively-misleading.}

These results illustrate that model-choice approaches would be
positively misleading -- supporting simpler models that cannot express
tipping point dynamics merely on account of them being similar. As the
data shown comes only from the basin of attraction near the unfished
equilibrium, near which all of the models are approximately linear and
approximately identical.

Model choice approaches trade off model complexity and fit to the data.
When the data come from a limited region of state-space -- as is
necessarily the case whenever there is a potential concern about tipping
point dynamics -- simpler models can fit just as well and will tend to
outperform more complex ones. This approach would be appropriate when
the dynamics can be expected to remain in the region of the training
data; for instance, if we only considered the forecasting accuracy of
the unfished population dynamics under each model.

In contrast, the decision-maker's problem of setting appropriate harvest
levels cannot exclude regions of state-space outside the observed range
when integrating over all possible decisions to find the optimal choice.
Such problems are not constrained to fisheries management but ubiquitous
across ecological decision-making and conservation where the greatest
concerns involve entering previously unobserved regions of state-space
-- whether that is the collapse of a fishery, the spread of an invasive,
or the loss of habitat.

\subsubsection{BNP-SDP expresses larger uncertainty in regions where the
data are
poor}\label{bnp-sdp-expresses-larger-uncertainty-in-regions-where-the-data-are-poor}

The parametric models perform worst when they propose a management
strategy outside the range of the observed data. The non-parametric
Bayesian approach, in contrast, allows a predictive model that expresses
a great deal of uncertainty about the probable dynamics outside the
observed range, while retaining very good predictive accuracy in the
range observed. The management policy dictated by the GP balance this
uncertainty against the immediate value of the harvest, and act to
stabilize the population dynamics in a region of state space in which
the predictions can be reliably reflected by the data.

\subsubsection{BNP-SDP has good predictive accuracy where data are
good}\label{bnp-sdp-has-good-predictive-accuracy-where-data-are-good}

While expressing larger uncertainty outside the observed data, the GP
can also provide a better fit with smaller uncertainty inside the range
of the observed data. This arises from the greater flexibility of the
Gaussian process, which describes a large family of possible curves.
Despite this flexibility, the GP can be described in relatively few
parameters and is thus far less likely to overfit.

\subsubsection{Risk-prone and risk-adverse value
functions}\label{risk-prone-and-risk-adverse-value-functions}

The degree to which the decision-making part of the algorithm (the SDP)
chooses to explore or avoid the resulting region of uncertainty can also
be influenced by the curvature of the value (profit) function $\Pi$.
Both to simplify the intution and avoid biasing this result, we have
chosen a profits that are linear in the catch and thus neither
risk-prone nor risk adverse. Making this function concave, representing
the typical assumption of diminishing returns, would make the SDP more
risk-adverse (as larger-than-expected stock sizes offer dimished returns
relative to the cost of smaller-than-expected stock sizes), and
strengthen the result shown here in which the BNP solution tends to
avoid the region of uncertainty. Sufficiently convex or risk-prone
functions could lead the SDP to attempt higher exploitation rates
despite the uncertainty. Understanding the relative roles of such
functions would be a promising direction for future investigation.

\subsubsection{The role of the prior}\label{the-role-of-the-prior}

Lastly, it should be noted that outside the data, the NBP reverts to the
prior, and consequently the choice of the prior can also play a
significant role in determining the optimal policy inferred by the SDP.
In the examples shown here we have selected a prior that is both
relatively uninformative (due to the broad priors placed on its
parameters $\ell$ and $\sigma$ and simple (mean zero, radial basis
function kernel). In practice, both the choice of mean and the choice of
the covariance function may be chosen to confer particular biological
properties, as well as more biologically informed priors for $\ell$ and
$\sigma$. In principle, this may allow a manager to improve the
performance of the BNP-SDP approach by adding only enough additional
detail as is justified. For instance, it would be possible to use a
linear or a Ricker-shaped mean in the prior without making the much
stronger assumption that the Ricker is the structurally correct model.
However, this influence raises challenges as well. For instance, in
choosing a trivial mean-zero prior, we bias the dynamics to transitions
to 0 in step $X_{t+1}$ from any stock size $X_t$ in the prior year, in
the absence of any other data. Future research must make sure that both
the prior and the value function are chosen appropriately for the
problem at hand.

\subsection{Future directions}\label{future-directions}

\subsubsection{Higher dimensions}\label{higher-dimensions}

In this simulated example, the underlying dynamics are truly governed by
a simple parametric model, allowing the parametric approaches to be more
accurate. Similarly, because the dynamics are one-dimensional dynamics
and lead to stable nodes (rather than other attractors such as
limit-cycles resulting in oscillations), the training data provides
relatively limited information about the dynamics. For these reasons, we
anticipate that in higher-dimensional examples characteristic of
ecosystem management problems that the machine learning approach will
prove even more valuable.

\subsubsection{Real-time learning}\label{real-time-learning}

In our treatment here we have ignored the possibility of learning during
the management phase, in which the additional observations of the stock
size could potentially improve parameter estimates. Of particular
interest in the context of the extreme uncertainty considered here is
the notion of ``active adaptive management'' or ``adaptive probing'',
which may actively seek to reduce uncertainty. The concept of adaptive
probing is one area that has explicitly addressed the extrapolation
uncertainty addressed here, though typically without the additional
issue of model uncertainty. As a result, adaptive probing strategies
suggest rather opposite conclusions than what we observe here. Such
adaptive probing or Dual Control (e.g. Ludwig and Walters (1982))
approaches trade off short term utility by choosing actions that can
reduce uncertainty. Adaptive probing strategies arise when it is
valuable to intentionally force a system far from the observed values
even when the expected value such actions is low, as it provides much
faster learning and consequent reduction of model uncertainty that can
allow greater value to be derived later on. For instance, Ludwig and
Walters (1982) show that it may be advantageous to fish an unexploited
population very heavily at first to obtain a better estimate of the
recruitment rate. This intuitive strategy when a population is governed
by a Ricker or Beverton-Holt-like dynamic would clearly be disastrous if
instead the dynamics contained an unforeseen tipping point. The best way
to learn where the edge lies may be to walk up to it, but it is also the
most dangerous. Future work should attempt to understand when such
active adaptive learning is valuable, and when it will increase the risk
of an irreversible transition.

\section{Acknowledgments}\label{acknowledgments}

This work was partially supported by the Center for Stock Assessment
Research, a partnership between the University of California Santa Cruz
and the Fisheries Ecology Division, Southwest Fisheries Science Center,
Santa Cruz, CA and by NSF~grant EF-0924195 to MM and NSF grant
DBI-1306697 to CB.

\section{Appendix}\label{appendix}

\subsection{Model definitions and
estimation}\label{model-definitions-and-estimation}

Equation S1: Ricker model.

\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]

Figure S1: Ricker model: prior and posterior distributions for parameter
estimates.

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-2.pdf}
\caption{Posteriors from the MCMC estimate of the Ricker model}
\end{figure}

Table S1: Parameterization of the priors

\% latex table generated in R 3.0.2 by xtable 1.7-1 package \% Thu Nov
14 00:55:33 2013

\begin{table}[ht]
\centering
\begin{tabular}{rlrr}
  \hline
 & parameter & lower\_bound & upper\_bound \\ 
  \hline
1 & r0 & 0.00 & 10.00 \\ 
  2 & K & 0.00 & 40.00 \\ 
  3 & sigma & 0.00 & 100.00 \\ 
   \hline
\end{tabular}
\end{table}

\[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \]

Eq S2: Myers model Figure S2: Myers model: Traces, prior and posterior
distributions for parameter estimates.

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-4.pdf}
\caption{Traces from the MCMC estimate of the Myers model}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-5.pdf}
\caption{Posterior distributions from the MCMC estimates of the Myers
model}
\end{figure}

Table S2: Parameterization of the priors \% latex table generated in R
3.0.2 by xtable 1.7-1 package \% Thu Nov 14 00:55:36 2013

\begin{table}[ht]
\centering
\begin{tabular}{rlrr}
  \hline
 & parameter & lower\_bound & upper\_bound \\ 
  \hline
1 & r0 & 0.00 & 10.00 \\ 
  2 & K & 0.00 & 40.00 \\ 
  3 & theta & 0.00 & 10.00 \\ 
  4 & sigma & 0.00 & 100.00 \\ 
   \hline
\end{tabular}
\end{table}

Eq S3: Allen model

\[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]

Figure S3: Allen model: prior and posterior distributions for parameter
estimates.

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-7.pdf}
\caption{Traces from the MCMC estimate of the Allen model}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-8.pdf}
\caption{Posteriors from the MCMC estimate of the Allen model}
\end{figure}

Table S3: Parameterization of the priors

\% latex table generated in R 3.0.2 by xtable 1.7-1 package \% Thu Nov
14 00:55:40 2013

\begin{table}[ht]
\centering
\begin{tabular}{rlrr}
  \hline
 & parameter & lower\_bound & upper\_bound \\ 
  \hline
1 & r0 & 0.00 & 10.00 \\ 
  2 & K & 0.00 & 40.00 \\ 
  3 & theta & 0.00 & 10.00 \\ 
  4 & sigma & 0.00 & 100.00 \\ 
   \hline
\end{tabular}
\end{table}

Eq S4: GP model Figure S4: GP model: prior and posterior distributions
for parameter estimates.

\begin{verbatim}
$traces_plot
\end{verbatim}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-101.pdf}
\caption{plot of chunk unnamed-chunk-10}
\end{figure}

\begin{verbatim}

$posteriors_plot
\end{verbatim}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-unnamed-chunk-102.pdf}
\caption{plot of chunk unnamed-chunk-10}
\end{figure}

Table S4: Parameterization of the priors

\subsection{Optimal Control Problem}\label{optimal-control-problem}

We seek the harvest policy $h(x)$ that maximizes:

\[ \max_{h_t} \sum_{t \in 0}^{\infty}  \Pi_t(X_t, h_t) \delta^t  \]

subject to the profit function $\Pi(X_t,h)$, discount rate $\delta$, and
the state equation

\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]

Where $Z_t$ is multiplicative noise function with mean 1, representing
stochastic growth. We will consider log-normal noise with shape
parameter $\sigma_g$.

Form this we can write down the Bellman recursion as:

\[V_t(x_t) = \max_h \mathbf{E} \left(\Pi(h_t, x_t) + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)\]

For simplicity we assume profit is simply linear in the realized harvest
(only enforcing the restriction that harvest can not exceed available
stock), $\Pi(h,x) = min(h,x)$.

\subsubsection{Pseudocode for the Bellman
iteration}\label{pseudocode-for-the-bellman-iteration}

\begin{Shaded}
\begin{Highlighting}[]
 \NormalTok{V1 <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(h_grid), function(h)\{}
      \NormalTok{delta *}\StringTok{ }\NormalTok{F[[h]] %*%}\StringTok{ }\NormalTok{V +}\StringTok{  }\KeywordTok{profit}\NormalTok{(x_grid, h_grid[h]) }
    \NormalTok{\})}
    \CommentTok{# find havest, h that gives the maximum value}
    \NormalTok{out <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(}\DecValTok{1}\NormalTok{:gridsize, function(j)\{}
      \NormalTok{value <-}\StringTok{ }\KeywordTok{max}\NormalTok{(V1[j,], }\DataTypeTok{na.rm =} \NormalTok{T) }\CommentTok{# each col is a diff h, max over these}
      \NormalTok{index <-}\StringTok{ }\KeywordTok{which.max}\NormalTok{(V1[j,])  }\CommentTok{# store index so we can recover h's }
      \KeywordTok{c}\NormalTok{(value, index) }\CommentTok{# returns both profit value & index of optimal h.  }
    \NormalTok{\})}
    \CommentTok{# Sets V[t+1] = max_h V[t] at each possible state value, x}
    \NormalTok{V <-}\StringTok{ }\NormalTok{out[}\DecValTok{1}\NormalTok{,]                        }\CommentTok{# The new value-to-go}
    \NormalTok{D[,OptTime-time}\DecValTok{+1}\NormalTok{] <-}\StringTok{ }\NormalTok{out[}\DecValTok{2}\NormalTok{,]       }\CommentTok{# The index positions}
\end{Highlighting}
\end{Shaded}

\subsubsection{Training data}\label{training-data}

Eacho of our models $f(S_t)$ must be estimated from training data, which
we simulate from the Allen model with parameters $r = $
\texttt{r p{[}1{]}}, $K =$ \texttt{r p{[}2{]}}, $C =$
\texttt{r p{[}3{]}}, and $\sigma_g =$ \texttt{r sigma\_g} for $T=$ 40
timesteps, starting at initial condition $X_0 = $ 5.5. The training data
can be seen in Figure 1.

\subsubsection{Sensitivity Analysis}\label{sensitivity-analysis}

Athanassoglou, Stergios, and Anastasios Xepapadeas. 2012. ``Pollution
Control with Uncertain Stock Dynamics: When, and How, to Be
Precautious.'' \emph{Journal of Environmental Economics and Management}
63 (3) (may): 304--320. doi:10.1016/j.jeem.2011.11.001.
\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069611001409}.

Bestelmeyer, Brandon T., Michael C. Duniway, Darren K. James, Laura M.
Burkett, and Kris M. Havstad. 2012. ``A Test of Critical Thresholds and
Their Indicators in a Desertification-prone Ecosystem: More Resilience
Than We Thought.'' Edited by Katharine Suding. \emph{Ecology Letters}
(dec). doi:10.1111/ele.12045.
\url{http://doi.wiley.com/10.1111/ele.12045}.

Brozović, Nicholas, and Wolfram Schlenker. 2011. ``Optimal Management of
an Ecosystem with an Unknown Threshold.'' \emph{Ecological Economics}
(jan): 1--14. doi:10.1016/j.ecolecon.2010.10.001.
\url{http://linkinghub.elsevier.com/retrieve/pii/S0921800910004167}.

Clark, Colin W., and Geoffrey P. Kirkwood. 1986. ``On Uncertain
Renewable Resource Stocks: Optimal Harvest Policies and the Value of
Stock Surveys.'' \emph{Journal of Environmental Economics and
Management} 13 (3) (sep): 235--244. doi:10.1016/0095-0696(86)90024-0.
\url{http://linkinghub.elsevier.com/retrieve/pii/0095069686900240}.

Cressie, Noel, Catherine a Calder, James S. Clark, Jay M. Ver Hoef, and
Christopher K. Wikle. 2009. ``Accounting for Uncertainty in Ecological
Analysis: the Strengths and Limitations of Hierarchical Statistical
Modeling.'' \emph{Ecological Applications} 19 (3) (apr): 553--70.
doi:10.1890/07-0744.1.
\url{http://www.ncbi.nlm.nih.gov/pubmed/19425416}.

Cuddington, Kim M., M. J. Fortin, and L. R. Gerber. 2013.
``Process-based Models Are Required to Manage Ecological Systems in a
Changing World.'' \emph{Ecosphere} 4 (February): 1--12.
doi:10.1890/ES12-00178.1.
\url{http://www.esajournals.org/doi/abs/10.1890/ES12-00178.1}.

Fischer, Joern, Garry D. Peterson, Toby a Gardner, Line J. Gordon, Ioan
Fazey, Thomas Elmqvist, Adam Felton, Carl Folke, and Stephen Dovers.
2009. ``Integrating Resilience Thinking and Optimisation for
Conservation.'' \emph{Trends in Ecology \& Evolution} 24 (10) (oct):
549--54. doi:10.1016/j.tree.2009.03.020.
\url{http://www.ncbi.nlm.nih.gov/pubmed/19665820}.

Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical Ecology: Why
Mechanistic Models?'' \emph{Journal of Mathematical Biology} 65 (6-7)
(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.

Gordon, H. S. 1954. ``The Economic Theory of a Common-property Resource:
the Fishery.'' \emph{The Journal of Political Economy} 62 (2): 124--142.
\url{http://www.jstor.org/stable/10.2307/1825571}.

Halpern, Benjamin S., Carissa J. Klein, Christopher J. Brown, Maria
Beger, Hedley S. Grantham, Sangeeta Mangubhai, Mary Ruckelshaus, et al.
2013. ``Achieving the Triple Bottom Line in the Face of Inherent
Trade-offs Among Social Equity, Economic Return, and Conservation.''
\emph{Proceedings of the National Academy of Sciences of the United
States of America} 110 (15) (apr): 6229--34.
doi:10.1073/pnas.1217689110.
\href{http://www.ncbi.nlm.nih.gov/pubmed/23530207 http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3625307/\&tool=pmcentrez/\&rendertype=abstract}{http://www.ncbi.nlm.nih.gov/pubmed/23530207
http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3625307\textbackslash{}\&tool=pmcentrez\textbackslash{}\&rendertype=abstract}.

Hilborn, Ray. 2007. ``Reinterpreting the State of Fisheries and Their
Management.'' \emph{Ecosystems} 10 (8) (oct): 1362--1369.
doi:10.1007/s10021-007-9100-5.
\url{http://www.springerlink.com/index/10.1007/s10021-007-9100-5}.

Hilborn, Ray, and Marc Mangel. 1997. \emph{The Ecological Detective:
Confronting Models with Data}. Princeton University Press.
\url{http://www.amazon.com/The-Ecological-Detective-Confronting-Models/dp/0691034974}.

Hughes, Terry P., Cristina Linares, Vasilis Dakos, Ingrid a van de
Leemput, and Egbert H. van Nes. 2013. ``Living Dangerously on Borrowed
Time During Slow, Unrecognized Regime Shifts.'' \emph{Trends in Ecology
\& Evolution} 28 (3) (mar): 149--55. doi:10.1016/j.tree.2012.08.022.
\url{http://www.ncbi.nlm.nih.gov/pubmed/22995893}.

Kocijan, Juš, Agathe Girard, Bla\textbackslash{}vz Banko, and Roderick
Murray-Smith. 2005. ``Dynamic Systems Identification with Gaussian
Processes.'' \emph{Mathematical and Computer Modelling of Dynamical
Systems} 11 (4): 411--424. doi:10.1080/13873950500068567.
\url{http://www.tandfonline.com/doi/abs/10.1080/13873950500068567}.

Ludwig, Donald, and Carl J. Walters. 1982. ``Optimal Harvesting with
Imprecise Parameter Estimates.'' \emph{Ecological Modelling} 14 (3-4)
(jan): 273--292. doi:10.1016/0304-3800(82)90023-0.
\url{http://linkinghub.elsevier.com/retrieve/pii/0304380082900230}.

Mangel, Marc, and Colin W. Clark. 1988. \emph{Dynamic Modeling in
Behavioral Ecology}. Edited by John Krebs and Tim Clutton-Brock.
Princeton: Princeton University Press.

Marescot, Lucile, Guillaume Chapron, Iadine Chadès, Paul L. Fackler,
Christophe Duchamp, Eric Marboutin, and Olivier Gimenez. 2013. ``Complex
Decisions Made Simple: a Primer on Stochastic Dynamic Programming.''
\emph{Methods in Ecology and Evolution} (jun).
doi:10.1111/2041-210X.12082.
\url{http://doi.wiley.com/10.1111/2041-210X.12082}.

May, Robert M., John R. Beddington, Colin W. Clark, Sidney J. Holt, and
R. M. Laws. 1979. ``Management of Multispecies Fisheries.''
\emph{Science (New York, N.Y.)} 205 (4403) (jul): 267--77.
doi:10.1126/science.205.4403.267.
\url{http://www.ncbi.nlm.nih.gov/pubmed/17747032}.

McAllister, M. K. 1998. ``Bayesian Stock Assessment: a Review and
Example Application Using the Logistic Model.'' \emph{ICES Journal of
Marine Science} 55 (6) (dec): 1031--1060. doi:10.1006/jmsc.1998.0425.
\href{http://icesjms.oxfordjournals.org/content/55/6/1031.short http://icesjms.oxfordjournals.org/cgi/doi/10.1006/jmsc.1998.0425}{http://icesjms.oxfordjournals.org/content/55/6/1031.short
http://icesjms.oxfordjournals.org/cgi/doi/10.1006/jmsc.1998.0425}.

Munch, Stephan B., Athanasios Kottas, and Marc Mangel. 2005. ``Bayesian
Nonparametric Analysis of Stockrecruitment Relationships.''
\emph{Canadian Journal of Fisheries and Aquatic Sciences} 62 (8) (aug):
1808--1821. doi:10.1139/f05-073.
\url{http://www.nrcresearchpress.com/doi/abs/10.1139/f05-073}.

Munch, Stephan B., Melissa L. Snover, George M. Watters, and Marc
Mangel. 2005. ``A Unified Treatment of Top-down and Bottom-up Control of
Reproduction in Populations.'' \emph{Ecology Letters} 8 (7) (may):
691--695. doi:10.1111/j.1461-0248.2005.00766.x.
\url{http://doi.wiley.com/10.1111/j.1461-0248.2005.00766.x}.

Rasmussen, Carl Edward, and C. K. I. Williams. 2006. \emph{Gaussian
Processes for Machine Learning}. Edited by Thomas Dietterich. Boston:
MIT Press,. \url{www.GaussianProcess.org/gpml}.

Reed, William J. 1979. ``Optimal Escapement Levels in Stochastic and
Deterministic Harvesting Models.'' \emph{Journal of Environmental
Economics and Management} 6 (4) (dec): 350--363.
doi:10.1016/0095-0696(79)90014-7.
\href{http://www.sciencedirect.com/science/article/pii/0095069679900147 http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}{http://www.sciencedirect.com/science/article/pii/0095069679900147
http://linkinghub.elsevier.com/retrieve/pii/0095069679900147}.

Roughgarden, Joan E., and F. Smith. 1996. ``Why Fisheries Collapse and
What to Do About It.'' \emph{Proceedings of the National Academy of
Sciences of the United States of America} 93 (10): 5078.
\url{http://www.pnas.org/content/93/10/5078.short}.

Schapaugh, Adam W., and Andrew J. Tyre. 2013. ``Accounting for
Parametric Uncertainty in Markov Decision Processes.'' \emph{Ecological
Modelling} 254 (apr): 15--21. doi:10.1016/j.ecolmodel.2013.01.003.
\url{http://linkinghub.elsevier.com/retrieve/pii/S0304380013000306}.

Scheffer, Marten, Stephen R. Carpenter, J. A. Foley, C. Folke, and B.
Walker. 2001. ``Catastrophic Shifts in Ecosystems.'' \emph{Nature} 413
(6856) (oct): 591--6. doi:10.1038/35098000.
\url{http://www.ncbi.nlm.nih.gov/pubmed/11595939}.

Sethi, Gautam, Christopher Costello, Anthony Fisher, Michael Hanemann,
and Larry Karp. 2005. ``Fishery Management Under Multiple Uncertainty.''
\emph{Journal of Environmental Economics and Management} 50 (2) (sep):
300--318. doi:10.1016/j.jeem.2004.11.005.
\url{http://linkinghub.elsevier.com/retrieve/pii/S0095069605000057}.

Su, Yu-Sung, and Masanao Yajima. 2013. \emph{R2jags: A Package for
Running Jags from R}. \url{http://cran.r-project.org/package=R2jags}.

Team, R. Core. 2013. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{http://www.r-project.org/}.

Weitzman, Martin L. 2013. ``A Precautionary Tale of Uncertain Tail
Fattening.'' \emph{Environmental and Resource Economics} 55 (2) (mar):
159--173. doi:10.1007/s10640-013-9646-y.
\url{http://link.springer.com/10.1007/s10640-013-9646-y}.

Williams, Byron K. 2001. ``Uncertainty , Learning , and the Optimal
Management of Wildlife.'' \emph{Environmental and Ecological Statistics}
8: 269--288. doi:10.1023/A:1011395725123.

Worm, Boris, Edward B. Barbier, Nicola Beaumont, J. Emmett Duffy, Carl
Folke, Benjamin S. Halpern, Jeremy B. C. Jackson, et al. 2006. ``Impacts
of Biodiversity Loss on Ocean Ecosystem Services.'' \emph{Science (New
York, N.Y.)} 314 (5800) (nov): 787--90. doi:10.1126/science.1132294.
\url{http://www.ncbi.nlm.nih.gov/pubmed/17082450}.

Worm, Boris, Ray Hilborn, Julia K. Baum, Trevor A. Branch, Jeremy S.
Collie, Christopher Costello, Michael J. Fogarty, et al. 2009.
``Rebuilding Global Fisheries.'' \emph{Science (New York, N.Y.)} 325
(5940) (jul): 578--85. doi:10.1126/science.1173146.
\url{http://www.ncbi.nlm.nih.gov/pubmed/19644114}.

\end{document}