### 2014-05-06 inst/examples/aic_replicates.Rmd

```{r}
## Set the paths for cache and figure
library(methods)
library(knitr)
basename <- gsub(".Rmd", "", file) # knitr:::knit_concord$get('infile')) 
opts_chunk$set(fig.path = paste("figure/", basename, "-", sep=""),
               cache.path = paste("cache/", basename, "/", sep=""))
opts_chunk$set(cache = 1)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, comment = NA, verbose = TRUE)
```

```{r}
library(nonparametricbayes)

models <- c("Myers","Allen")

parameters <- list(Myers = list(
  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2))),
                   Allen = list(
  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)))
                   )
nuisance_pars <- c("sigma_g")
nuisance_values <- list(sigma_g = c(0.01, 0.05, 0.1))
replicates <- c(1111, 2222, 3333, 4444, 5555, 6666, 7777, 8888) # seeds

## @knitr posterior-mode
require(modeest)
posterior.mode <- function(x) {
  mlv(x, method="shorth")$M
}

sensitivity <- function(model, parameters, nuisance, seed){

  if(model == "Myers")
    f <- Myers
  else if(model == "Allen")
    f <- RickerAllee

  sigma_g <- nuisance[["sigma_g"]]
  z_g <- function() rlnorm(1, 0, sigma_g)
  p <- parameters

  config = list(
    f = f,
    p = p,
    sigma_g = sigma.g,
    z_g = z_g,
    x_grid = seq(0, 15, length=50),
    h_grid = x_grid,
    profit = function(x,h) pmin(x, h),
    delta = 0.01,
    OptTime = 50,  # stationarity with unstable models is tricky thing
    reward = 0,
    xT = 0,
    Xo = 5.5,# observations start from
    x0 = 8, # simulation under policy starts from
    Tobs = 40,
    MaxT = 1000) # timeout for value iteration convergence

  attach(config)
  # replicate over random seed
  yields <- sapply(seed, 
                   function(seed_i){
                     set.seed(seed_i)

## Simulate data
                     x <- numeric(Tobs)
                     x[1] <- Xo
                     nz <- 1
                     for(t in 1:(Tobs-1))
                       x[t+1] = z_g() * f(x[t], h=0, p=p)

                     X = c(rep(0,nz), pmax(rep(0,Tobs-1), x[1:(Tobs-1)]))
                     Y = c(rep(0,nz), x[2:Tobs])

## GP Stuff

                     ## @knitr gp-priors
                     s2.p <- c(5,5)  
                     d.p = c(10, 1/0.1)

                     ## @knitr gp
                     gp <- gp_mcmc(X, y=Y, n=1e5, s2.p = s2.p, d.p = d.p)
                     gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)

                     matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid) 
                     opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)

## @knitr jags-setup
y <- x 
N <- length(x);
jags.data <- list("N","y")
n.chains <- 6
n.iter <- 1e6
n.burnin <- floor(10000)
n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
n.update <- 10

## @knitr common-priors
stdQ_prior_p <- c(1e-6, 100)
stdR_prior_p <- c(1e-6, .1)
stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])

## @knitr allen-model
K_prior_p <- c(0.01, 20.0)
r0_prior_p <- c(0.01, 6.0)
theta_prior_p <- c(0.01, 20.0)

bugs.```
```{r}
model <- 
paste(sprintf(
"model{
  K     ~ dunif(%s, %s)
  r0    ~ dunif(%s, %s)
  theta ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)", 
  K_prior_p[1], K_prior_p[2],
  r0_prior_p[1], r0_prior_p[2],
  theta_prior_p[1], theta_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2]),

  "
  iQ <- 1 / (stdQ * stdQ);
  y[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K)* (y[t] - theta) / K 
    y[t+1] ~ dlnorm(mu[t], iQ) 
  }
}")
writeLines(bugs.model, "allen_process.bugs")

## @knitr allen-priors
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
r0_prior <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
theta_prior <- function(x) dunif(x, theta_prior_p[1], theta_prior_p[2])
par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
                    r0 = r0_prior, theta = theta_prior,
                    stdQ = stdQ_prior)

## @knitr allen-mcmc
jags.params=c("K","r0","theta","stdQ") # be sensible about the order here
jags.inits <- function(){
  list("K"= 10 * rlnorm(1,0, 0.1),
       "r0"= 1 * rlnorm(1,0, 0.1) ,
       "theta"=   5 * rlnorm(1,0, 0.1) , 
       "stdQ"= abs( 0.1 * rlnorm(1,0, 0.1)),
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

set.seed(1234)
# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
allen_jags <- do.call(jags, list(data=jags.data, inits=jags.inits, 
                                      jags.params, n.chains=n.chains, 
                                      n.iter=n.iter, n.thin=n.thin, 
                                      n.burnin=n.burnin, 
                                      model.file="allen_process.bugs"))

# Run again iteratively if we haven't met the Gelman-Rubin convergence criterion
recompile(allen_jags) # required for parallel
allen_jags <- do.call(autojags, 
											list(object=allen_jags, n.update=n.update, 
                           n.iter=n.iter, n.thin = n.thin))

## @knitr allen-traces
tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes the hard way...
allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
## @knitr allen-output
A <- allen_posteriors
A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
pardist <- acast(A, index ~ variable)

## @knitr ricker-model
K_prior_p <- c(0.01, 40.0)
r0_prior_p <- c(0.01, 20.0)
bugs.```
```{r}
model <- 
paste(sprintf(
"model{
  K    ~ dunif(%s, %s)
  r0    ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)", 
  K_prior_p[1], K_prior_p[2],
  r0_prior_p[1], r0_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2]),
  "
  iQ <- 1 / (stdQ * stdQ);
  y[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K) 
    y[t+1] ~ dlnorm(mu[t], iQ) 
  }
}")
writeLines(bugs.model, "ricker_process.bugs")

## @knitr ricker-priors
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
r0_prior <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
                   r0 = r0_prior, stdQ = stdQ_prior)

## @knitr ricker-mcmc
jags.params=c("K","r0", "stdQ")
jags.inits <- function(){
  list("K"= 10 * rlnorm(1,0,.5),
       "r0"= rlnorm(1,0,.5),
       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345) 
ricker_jags <- do.call(jags, 
                       list(data=jags.data, inits=jags.inits, 
                            jags.params, n.chains=n.chains, 
                            n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
                            model.file="ricker_process.bugs"))
recompile(ricker_jags)
ricker_jags <- do.call(autojags, 
                       list(object=ricker_jags, n.update=n.update, 
														n.iter=n.iter, n.thin = n.thin, 
														progress.bar="none"))

## @knitr ricker-traces
tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
names(ricker_posteriors) = c("index", "variable", "value", "chain")

## @knitr ricker-output
A <- ricker_posteriors
A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
ricker_pardist <- acast(A, index ~ variable)

## @knitr myers-model
r0_prior_p <- c(.0001, 10.0)
theta_prior_p <- c(.0001, 10.0)
K_prior_p <- c(.0001, 40.0)
bugs.```
```{r}
model <- 
paste(sprintf(
"model{
  r0    ~ dunif(%s, %s)
  theta    ~ dunif(%s, %s)
  K    ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)", 
  r0_prior_p[1], r0_prior_p[2],
  theta_prior_p[1], theta_prior_p[2],
  K_prior_p[1], K_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2]),

  "
  iQ <- 1 / (stdQ * stdQ);

  y[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- log(r0)  + theta * log(y[t]) - log(1 + pow(abs(y[t]), theta) / K)
    y[t+1] ~ dlnorm(mu[t], iQ) 
  }
}")
writeLines(bugs.model, "myers_process.bugs")

## @knitr myers-priors
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
r_prior     <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
theta_prior <- function(x) dunif(x, theta_prior_p[1], theta_prior_p[2])
par_priors <- list( deviance = function(x) 0 * x, K = K_prior,
                    r0 = r_prior, theta = theta_prior, 
                    stdQ = stdQ_prior)

## @knitr myers-mcmc
jags.params=c("r0", "theta", "K", "stdQ")
jags.inits <- function(){
  list("r0"= 1 * rlnorm(1,0,.1), 
       "K"=    10 * rlnorm(1,0,.1),
       "theta" = 1 * rlnorm(1,0,.1),  
       "stdQ"= sqrt(0.2) * rlnorm(1,0,.1),
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345)
myers_jags <- do.call(jags, 
                      list(data=jags.data, inits=jags.inits, 
													 jags.params, n.chains=n.chains, 
													 n.iter=n.iter, n.thin=n.thin,
                           n.burnin=n.burnin, 
                           model.file="myers_process.bugs"))
recompile(myers_jags)
myers_jags <- do.call(autojags, 
                      list(myers_jags, n.update=n.update, 
                           n.iter=n.iter, n.thin = n.thin, 
                           progress.bar="none"))

## @knitr myers-traces
tmp <- lapply(as.mcmc(myers_jags), as.matrix) # strip classes
myers_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
names(myers_posteriors) = c("index", "variable", "value", "chain")
## @knitr myers-output
A <- myers_posteriors
A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
myers_pardist <- acast(A, index ~ variable)

allen_deviance  <- - posterior.mode(pardist[,'deviance'])
ricker_deviance <- - posterior.mode(ricker_pardist[,'deviance'])
myers_deviance  <- - posterior.mode(myers_pardist[,'deviance'])
true_deviance   <- 2*estf(c(p, sigma_g))
mle_deviance    <- 2*estf(c(est$p, est$sigma_g))
aictable <- data.frame(Allen = allen_deviance + 2*(1+length(bayes_pars)),  # +1 for noise parameter
                       Ricker = ricker_deviance + 2*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + 2*(1+length(myers_bayes_pars)), 
                       row.names = c("AIC"))
bictable <- data.frame(Allen = allen_deviance + log(length(x))*(1+length(bayes_pars)), 
                       Ricker = ricker_deviance + log(length(x))*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + log(length(x))*(1+length(myers_bayes_pars)), 
                       row.names = c("BIC"))

df <- rbind(dictable, aictable, bictable)

                   })

  dat <- melt(yields, id=names(yields[[1]]))
}
```

```{r}
model <- "Allen"
allen1.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```
```{r}
model <- "Allen"
allen2.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```
```{r}
model <- "Allen"
allen1.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))
```
```{r}
model <- "Allen"
allen2.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))

```
```{r}
model <- "Myers"
Myers1.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```
```{r}
model <- "Myers"
Myers2.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```
```{r}
model <- "Myers"
Myers1.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))

```
```{r}
model <- "Myers"
Myers2.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
## Assemble into data.frame
allen_dat <- rbind(allen1.01, allen1.05, 
             allen2.01, allen2.05) 
myers_dat <- rbind(Myers1.01, Myers1.05, 
             Myers2.01, Myers2.05)
```