### 2014-05-08 sensitivity.Rmd

```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls())
source("~/.knitr_defaults.R")
opts_knit$set(upload.fun = socialR::flickr.url)
opts_chunk$set(error=TRUE)
library(knitcitations)
library(nonparametricbayes) 
library(reml)
```

```{r}
models <- c("Myers","Allen")

parameters <- list(Myers = list(
  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2)),
  c(r=1.5 + rnorm(1, 0, .1), theta=2.5 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .2))),
                   Allen = list(
  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)),
  c(r=2 + rnorm(1, 0, .1), K=8 + rnorm(1, 0, .1), C=5 + rnorm(1, 0, .2)))
                   )
nuisance_pars <- c("sigma_g")
nuisance_values <- list(sigma_g = c(0.01, 0.05, 0.1))
replicates <- c(1111, 2222, 3333, 4444, 5555, 6666, 7777, 8888) # seeds

```

Rather than attempt a large list of nested loops, we do better to launch seperate scripts and write to a common data format.  We define a function that takes as arguments all things we wish to vary in the senstivity analysis, and returns a data.frame with the values of interest

```{r }             
sensitivity <- function(model, parameters, nuisance, seed){

  if(model == "Myers")
    f <- Myers
  else if(model == "Allen")
    f <- RickerAllee

sigma_g <- nuisance[["sigma_g"]]
p <- parameters
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1
x_grid <- seq(0, 15, length=50)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 50  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
Xo <- 5.5# observations start from
x0 <- 8 # simulation under policy starts from
Tobs <- 40
MaxT <- 1000 # timeout for value iteration convergence

  # replicate over random seed
  yields <- sapply(seed, 
                   function(seed_i){
                     set.seed(seed_i)

                     x <- numeric(Tobs)
                     x[1] <- Xo
                     nz <- 1
                     for(t in 1:(Tobs-1))
                       x[t+1] = z_g() * f(x[t], h=0, p=p)

                     X = c(rep(0,nz), pmax(rep(0,Tobs-1), x[1:(Tobs-1)]))
                     Y = c(rep(0,nz), x[2:Tobs])

                     ## @knitr gp-priors
                     s2.p <- c(5,5)  
                     d.p = c(10, 1/0.1)

                     ## @knitr gp
                     gp <- gp_mcmc(X, y=Y, n=1e5, s2.p = s2.p, d.p = d.p)
                     gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)

                     matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid) 
                     opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)

                     ## @knitr mle-opt
                     matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
                     opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

                     OPT = data.frame(GP = opt_gp$D, True = opt_true$D)
                     policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
                     names(policies) <- c("stock", "method", "value")

                     sims <- lapply(OPT, function(D){
                       set.seed(1)
                       lapply(1:100, function(i) 
                         ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
                              )
                       })

                     dat <- melt(sims, id=names(sims[[1]][[1]]))
                     sims_data <- data.table(dat)
                     setnames(sims_data, c("L1", "L2"), c("method", "reps")) 
                     # Legend in original ordering please, not alphabetical: 
                     sims_data$method = factor(sims_data$method, ordered=TRUE, levels=names(OPT))

                     Profit <- sims_data[, sum(profit), by=c("reps", "method")]
                     tmp <- dcast(Profit, reps ~ method)
                     tmp <- tmp / tmp[,"True"]
                     tmp <- melt(tmp[2:dim(tmp)[2]])
                     tmp$value[tmp$variable == "GP"]  
                     })

  yields_dat <- melt(yields)
  names(yields_dat) <- c("replicate", "simulation", "value")

  # Make ids factors, not integers
  yields_dat$replicate <- as.factor(yields_dat$replicate)
  yields_dat$simulation <- as.factor(yields_dat$simulation)

  # definitions of id codes.  In this case the id number is it's own definition.  
  rep_ids <-  levels(yields_dat$replicate)
  names(rep_ids) <- rep_ids
  sim_ids <-  levels(yields_dat$simulation)
  names(sim_ids) <- sim_ids

  dat <- data.frame(model = model, 
                    pars = as.list(parameters), 
                    replicate = yields_dat$replicate, 
                    sim = yields_dat$simulation, 
                    value = yields_dat$value, 
                    noise = sigma_g)
}
```

```{r}
model <- "Allen"
allen1.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Allen"
allen2.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Allen"
allen1.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Allen"
allen2.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Myers"
Myers1.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Myers"
Myers2.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Myers"
Myers1.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
model <- "Myers"
Myers2.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1234, 2222, 3333))
```

```{r}
save(list=ls(), file="sensitivity.rda")
```

```{r data}
allen_dat <- rbind(allen1.01, allen1.05, 
             allen2.01, allen2.05) 
m <- rbind(Myers1.01, Myers1.05, 
             Myers2.01, Myers2.05)
myers_dat <- m[c(1:2,4,3,5:8)]
names(myers_dat) <- names(allen_dat)
model_dat <- rbind(allen_dat, myers_dat)
```

```{r}
dat <- model_dat
dat$pars.r <- factor(dat$pars.r, labels=c("A", "B", "C", "D"))
dat <- dat[c(1:2,5:6, 8, 7)]
dat$noise <- factor(dat$noise)
names(dat) <- c("model", "parameters", "replicate", "simulation", "noise", "value")
```

```{r}
ggplot(dat) + geom_histogram(aes(value)) + xlim(0,1.0) + theme_bw() + xlab("value as fraction of the optimal")

ggplot(dat) + geom_histogram(aes(value, fill=noise)) + xlim(0,1.0) + theme_bw() + xlab("value as fraction of the optimal") 
```

```{r sensitivity}
ggplot(dat) + geom_histogram(aes(value, fill=noise)) + xlim(0,1.0) + theme_bw() + xlab("value as fraction of the optimal") + facet_wrap(~model, ncol=1)
```

```{r}
## Extract acutal parameter values corresponding to each parameter set
  p1 = levels(factor(model_dat$pars.r))
  p2 = levels(factor(model_dat$pars.K))
  p3 = levels(factor(model_dat$pars.C))

  A = c(r = p1[1], K = p2[1], theta = p3[1])
  B = c(r = p1[2], K = p2[2], theta = p3[2])
  C = c(r = p1[3], K = p2[3], C = p3[3])
  D = c(r = p1[4], K = p2[4], C = p3[4])
```

```{r eml, dependson='data'}
  rep_ids <-  levels(dat$replicate)
  sim_ids <-  levels(dat$sim)

  require(reml)
  col.defs = c(model = "name of the model used to simulate the data",
               parameters = "which set of model parameters was used",
               replicate = "replicate id numbers.  Each unique simulation of the model results in a policy function which is then replicated under each replicate id number to determine the realized value relative to the optimal value",
               simulation = "simulation id number. For each set of parameters for the model, multiple simulations of training data are generated and assigned unique simulation numbers",            
               noise = "sigma_g value, multiplicative growth noise scale, as a log-normal log-standard deviation"
               value = "the value derived from this realization of the policy function estimated with the Gaussian process relative to the theoretical optimum value if the true model and parameter values were known",)

  unit.defs = list(model = c("Myers" = "a 3 parameter model of Beverton-Holt-like recruitment with the addition of an Allee effect (and fourth parameter for scaling the log-normal growth noise); based on a Myers et al 1995 (doi:10.1126/science.269.5227.1106)", 
                               "Allen" = "a 3 parameter model of Ricker-like recruitment with the addition of an Allee effect (and fourth parameter for scaling the log-normal growth noise); based on Allen et al 2005 (doi:10.1080/10236190412331335373)"),
                  parameters = 
  c("A" = paste("First set of parameter values drawn at random for Myers model. r, K, theta respectively are:", A), 
    "B" = paste("Second set of parameter values drawn at random for Myers model. r, K, theta respectively are:", B), 
    "C" = paste("First set of parameter values drawn at random for Myers model. r, K, C respectively are:", C), 
    "D" = paste("Second set of parameter values drawn at random for Myers model. r, K, C respectively are:", D)), 
                  replicate = rep_ids,
                  simulation = sim_ids, 
                  noise = "density/density (pure)",
                  value = "Realized Dollars/Potential Dollars")

emldat =  data.set(dat, col.defs=col.defs, unit.defs=unit.defs)  
eml_write(emldat, creator="Carl Boettiger <cboettig@gmail.com>")
```

```{r, eval=FALSE}
require(reshape2)
require(snowfall)
sfInit(parallel=TRUE, cpu=8)
dat <- sfLapply(nuisance_values[["sigma_g"]], function(sigma){
  dat <- lapply(parameters[["Allen"]], function(p){
         sensitivity("Allen", 
                   parameters = p, 
                   nuisance = c(sigma_g = sigma), 
                   seed=c(1234, 2222, 3333))
              })
  dat <- melt(dat, id=names(dat[[1]]))
})
dat <- melt(dat, id=names(dat[[1]]))
```
```{r eval=FALSE}
require(reshape2)
myers_dat <- sfLapply(nuisance_values[["sigma_g"]], function(sigma){
  dat <- lapply(parameters[["Myers"]], function(p){
         sensitivity("Myers", 
                   parameters = p, 
                   nuisance = c(sigma_g = sigma), 
                   seed=c(1234, 2222, 3333))
              })
  dat <- melt(dat, id=names(dat[[1]]))
})
myers_dat <- melt(myers_dat, id=names(dat[[1]]))

```