### 2014-05-23 manuscripts/manuscript.tex
updates as sent to steve

\documentclass[author-year, 12pt,review]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}
\usepackage{lineno} % add 
  \linenumbers % turns line numbering on 
\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%% Redefines the elsarticle footer
\makeatletter
\def\ps@pprintTitle{%
 \let\@oddhead\@empty
 \let\@evenhead\@empty
 \def\@oddfoot{\it \hfill\today}%
 \let\@evenfoot\@oddfoot}
\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{longtable}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Avoiding tipping points in fisheries management: A Bayesian Non-Parametric approach},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{Avoiding tipping points in fisheries management: A Bayesian
Non-Parametric approach}
    \author[cstar]{Carl Boettiger\corref{c1}}
   \ead{cboettig@gmail.com} 
   \cortext[c1]{Corresponding author}
    \author[cstar]{Marc Mangel}

    \author[noaa]{Stephan Munch}

      \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and
Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA
95064, USA}    
    \address[marc]{Center for Stock Assessment Research, Department of Applied Math and
Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA
95064, USA and Department of Biology, University of Bergen, Bergen,
Norway 9020}    
    \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric
Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}

  \begin{abstract}
  Model uncertainty and limited data coverage are fundamental challenges
  to robust management of human intervention in a natural system. These
  challenges are acutely highlighted by concerns that many ecological
  systems may contain tipping points, such as Allee population
  sizes.\\Before a collapse, we do not know where the tipping points lie,
  if they exist at all. Hence, we know neither a complete model of the
  system dynamics nor do we have access to data in some large region of
  state-space where such a tipping point might exist. We illustrate how a
  Bayesian Non-Parametric (BNP) approach using a Gaussian Process (GP)
  prior provides a flexible representation of this inherent uncertainty.
  We embed GPs in a stochastic dynamic programming framework in order to
  make robust management predictions with both model uncertainty and
  limited data. We conduct a set of simulations to evaluate this approach
  as compared with the standard approach of using model selection to
  choose from a set of candidate models. We find that model selection
  erroneously favors models without tipping points -- leading to harvest
  policies that guarantee extinction. The GP approach performs nearly as
  well as the true model and significantly outperforms standard
  appraoches. We illustrate this using examples of simulated
  single-species dynamics, where the standard model selection approach
  should be most effective, and find that it still fails to account for
  uncertainty appropriately and leads to population crashes, while
  management based on the GP approach does not, since it does not
  underestimate the uncertainty outside of the observed data.
  \end{abstract}
   \begin{keyword} Bayesian \sep Structural Uncertainty \sep Nonparametric \sep Optimal Control \sep Decision Theory \sep Gaussian Processes \sep Fisheries Management \sep \end{keyword}
 \end{frontmatter}

\section{Introduction}\label{introduction}

Decision making under uncertainty is a ubiquitous challenge in the
management of human intervention in natural resources and conservation.
Ecological dynamics are frequently complex and difficult to measure,
making uncertainty in our understanding a persistent challenge.
Decision-theoretic approaches provide a framework to determine the best
sequence of actions in face of uncertainty, but only when that
uncertainty can be meaningfully quantified (Fischer et al. 2009). Over
the last four decades (but stretching even further back in time), marked
by the seminal contributions of Clark (1976); Clark (2009) and Walters
and Hilborn (1978), dynamic optimization methods, particularly
Stochastic Dynamic Programming (SDP, Clark (2009)), have become
increasingly important as a means of understanding how to management
human intervention into natural systems. Concomittantly, there has been
increasing recognition of the importance of Allee population sizes or
`tipping points' (Scheffer et al. 2001, Polasky et al. 2011) that
management actions may inadvertently cause populations to cross.
Scheffer et al. (2009) emphasized the difficulties of formulating even
qualitatively correct models of the underlying processes.

We develop a novel approach to address these concerns in the context of
fisheries; though the underlying challenges and methods are germane to
many other problems of conservation or natural resource exploitation,
such as forestry. Economic value and ecological concern have made marine
fisheries the crucible for much of the founding work (Gordon 1954, Clark
1976, 2009, May et al. 1979, Reed 1979, Ludwig and Walters 1982) on
management under uncertainty. Global declines (Worm et al. 2006) and the
controversy surrounding their interpretation (Hilborn 2007, Worm et al.
2009) have made understanding these challenges all the more pressing.

Uncertainty enters the decision-making process at many levels: intrinsic
stochasticity in biological processes, measurements, and implementation
of policy (\emph{e.g.} Reed 1979, Clark and Kirkwood 1986, Roughgarden
and Smith 1996, Sethi et al. 2005), parametric uncertainty (\emph{e.g.}
Ludwig and Walters 1982, Hilborn and Mangel 1997, McAllister 1998,
Schapaugh and Tyre 2013), and model or structural uncertainty
(\emph{e.g.} Williams 2001, Cressie et al. 2009, Athanassoglou and
Xepapadeas 2012). Of these, structural uncertainty is generally the
hardest to quantify. Typical approaches assume a weak notion of model
uncertainty in which the correct model (or reasonable approximation) of
the dynamics must be identified from among a handful of alternative
models, using either model choice, model averaging, or introducing yet
greater model complexity of which others may be special cases (model
averaging being one such way to construct such a model) (Williams 2001,
Cressie et al. 2009, Athanassoglou and Xepapadeas 2012). Even setting
aside other computational and statistical concerns (e.g. Cressie et al.
(2009)), these approaches do not address our second concern -
representing uncertainty outside the observed data range.

Model uncertainty is particularly insidious when model predictions must
be made outside of the range of data on which the model was estimated,
and models may be abused (Mangel et al. 2001). Extrapolation uncertainty
is felt most keenly in decision-theoretic (or optimal control)
applications, since (a) exploring the potential action space typically
involves considering actions that may move the system outside the range
of observed behavior, and (b) decision-theoretic algorithms rely not
only on reasonable estimates of the expected outcomes, but depend on the
weights given to all possible outcomes (\emph{e.g.} Weitzman 2013).

The dual concerns of model uncertainty and incomplete data coverage pose
a substantial challenge to existing decision-theoretic approaches
(Brozović and Schlenker 2011). Because intervention is often (but not
always, see Hughes et al. (2013)) too late after a tipping point has
been crossed, management is most often concerned with avoiding
potentially catastrophic tipping points before any data is available at
or following a transition that would more clearly reveal these regime
shift dynamics (e.g. Bestelmeyer et al. 2012).

We illustrate how a stochastic dynamic programming (SDP) algorithm
(Mangel and Clark 1988, Marescot et al. 2013) can be driven by the
predictions from a Bayesian Non-Parametric (BNP) model of population
dynamics (Munch et al. 2005a). This provides two distinct advantages
compared with contemporary approaches. First, using a BNP sidesteps the
need for an accurate model-based description of the system dynamics.
Second, the BNP can better reflect uncertainty that arises when
extrapolating a model outside of the data on which it was fit. We
illustrate that when the correct model is not known, this latter feature
is crucial to providing a robust decision-theoretic approach in face of
substantial structural uncertainty.

This paper represents the first ecological application of the SDP
decision-making framework without an a priori model of the underlying
dynamics. In contrast to parametric models which can only reflect
uncertainty in parameter estimates, the BNP approach provides a more
state-space dependent representation of uncertainty. This permits a much
greater uncertainty far from the observed data than near the observed
data. These features allow the Gaussian Process Dynamic Programming
(GPDP) approach to find robust management solutions in face of limited
data and without knowledge of the correct model structure.

To illustrate GPDP, we consider the performance of the GPDP policy
against the policies derived under several alternative parametric models
(Reed 1979, Ludwig and Walters 1982, Mangel and Clark 1988). The nature
of decision-making problems provides a compelling and pragmatic way to
compare models: Rather than compare models in terms of best fit to data,
we define model performance in the concrete terms of the
decision-maker's objectives.

\section{Approach and Methods}\label{approach-and-methods}

We first describe the requirements of dynamic optimization for the
management of human intervention in natural resource systems. After that
we describe three parametric models for population dynamics and the GP
description of population dynamics.

\subsubsection{Requirements of dynamic
optimization}\label{requirements-of-dynamic-optimization}

Dynamic optimization requires characterizing the dynamics of a state
variable (or variables), a control action, and a value function. In this
paper, we consider only a single state variable, so that the ideas can
be described as simply as possible. That is, we focus on a class of
one-dimensional models of population dynamics as a `best-case' scenario
for the parametric model performance relative to the GP. This is
abest-case scenario for the parametric models because the underlying
dynamics are always simuated from one the parametric models in the set,
whereas in reality we never have the ``true'' model among our
candidates. By choosing one-dimensional models with few parameters, we
limit the chance that poor performance will be due to our inability to
estimate parameters well, something that becomes a more severe problem
for higher-dimensional parametric models. Further, the parametric models
we consider are those most commonly used in modeling stock-recruitment
dynamics or to model sudden transitions between alternative stable
states.

\subsubsection{Parametric models}\label{parametric-models}

We let $X_t$ denote the size (numbers or biomass) of the focal
population at time $t$ and assume that in the absence of take its
dynamics are

\[ X_{t+1}= Z_t f(X_t, h_t, p) \]

Where Z(t) is log-normally distributed process stochasticity (Reed 1979)
and $p$ is a vector of parameters to be estimated from the data. We
describe three choices for $f(X_t,p)$ in the next section.In this simple
model, the control action is a harvest or take, $h_t$, measured in the
same units as $X$, at time $t$. Thus, in the presence of take, the
population size on the right hand side of Eqn 1 is replaced by
$S_t=X_t-h_t$.

To construct the value function, we consider a return when $X_t=x$ and
harvest $h_t=h$ denoted as the reward, $R(x,h)$. For example, if the
return is simply the harvest at time $t$, then $R(x,h)=min(x,h)$. We
assume that future harvests are discounted relative to current ones at a
constant rate of discount $\delta$ and ask for the harvest policy that
maximizes total discounted harvest between the current time $t$ and a
final time $T$, that is we seek to maximize over choices of harvest
$\mathbf{E}_{X_t+1} [ \sum_{t = 0}^{T}  R_t(X_t, h_t) \delta^t]$, where
the state dynamics are given by Eqn 1 and $\mathbf{E}$ denotes the
expectation over the process stochasticity in the future population
state.

In order to find that policy, we introduce the value function$V_t(x_t)$
representing the total discounted catch from time $t$ onwards given that
$X_t=x$. This value function satisfies an equation of Stochastic Dynamic
Programming (SDP, Mangel and Clark 1988, Clark and Mangel 2000, Clark
2009), sometimes called the Bellman-Hamilton-Jacobi equation (Mangel
2014).

\begin{equation}
V_t(x) = \max_{h_t} \left[R(h_t, x_t) + \delta \cdot \mathbf{E}_{X_t+1} V_{t+1}( Z_{t+1} f(x - h_t)) \right]
\end{equation}

That is, at time $t$, when population size is $x$ and harvest $h_t$ is
applied, the immediate return is $R(h_t, x_t)$ and the population size
after the take is $x-h_t$, which is then translated into $X_{t+1}$ by
Eqn 1. We consider the finite time problem with \$T=\$1000, which we
solve using value iteration algorithm (Mangel and Clark 1988, Clark and
Mangel 2000).

\subsubsection{Parametric models}\label{parametric-models}

We consider three candidate parametric models of the stock-recruitment
dynamics: The Ricker model, the Allen model (Allen and Tanner 2005), the
Myers model (Myers et al. 1995), equations (1)-(3), which we
parameterize as follows for $f(S_t,p)$ in Eqn 1. In all three, we let
$K$ denote the carrying capacity and $r$ the maximum per capita growth
rate. In the Ricker model,

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \[ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } \]
\end{enumerate}

In the Allen model,

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{1}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \[ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_c\right)} \]
\end{enumerate}

Where $r$ and $K$ are as before, and $X_c$ denotes the location of the
unstable steady state (i.e., the tipping point).

In the Myers model,

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{2}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \[ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 + \frac{S_t^\theta}{K}}\]
\end{enumerate}

Where $\theta = 1$ corresponding to Beverton-Holt dynamics and $\theta$
\textgreater{} 2 leads to Allee effects and multiple stable states.

The Ricker model involves two parameters, corresponding to a growth rate
and a carrying capacity, and cannot support alternative stable growth
rate dynamics. The Allen model resembles the Ricker dynamics with an
added Allee effect parameter (Courchamp et al. 2008), below which the
population cannot persist. The Myers model also has three parameters and
contains an Allee threshold, but has compensatory rather than
over-compensatory density dependence (resembling a Beverton-Holt curve
rather than a Ricker curve at high densities.)

The multiplicative log-normal stochasticity perturbs the growth
predicted by each of these deterministic model skeletons. This
introduces one additional parameter $\sigma$ that must be estimated.

Because of our interest in management performance in the presence of
tipping points, all of our simulations are based on the Allen model. The
Allen model is thus ``structurally correct'' and is expected to provide
a best-case scenario when the `true' dynamics are known. The Ricker
model is a reasonable approximation of these dynamics far from the Allee
threshold (but lacks threshold dynamics), while the Myers model shares
the essential feature of a threshold but differs in the structure.

We consider a period of 40 years of training data: long enough that the
estimates do not depend on the particular realization, while longer
times are not likely to provide substantial improvement (i.e.~the
results are not sensitive to this interval). Each of the models
(described below) is fit to the same training data (Figure 1).

We inferred posterior distributions for the parameters of each model by
Gibbs sampling (Gelman et al. (2003) implemented in R (R Core Team 2013)
using jags, (Su and Masanao Yajima 2013)). We choose uniform priors for
all parameters (See appendix Tables S1-S3; R code provided). We show
one-step-ahead predictions of these model fits in Figure 1. We tested
each chain for Gelman-Rubin convergence and results were robust to
longer runs. For each simulation we also applied several commonly used
model selection criteria (AIC, BIC, DIC, see Burnham and Anderson
(2002)) to identify the best fitting model.

\subsubsection{The Gaussian Process
model}\label{the-gaussian-process-model}

The core difference for our purpose between the GP representation and
the parametric models we have already introduced is the
``non-parametric'' property, which means that an estimated GP model is
defined explicitly in reference to some observed data. Unlike the models
above, it cannot be specified by the value of some parameters alone. The
nonparametric moniker for the GP approach is unfortunate as (a) it still
involves the estimation of parameters, and (b), statisticians also use
non-parametric to mean models that make only weak distributional
assumptions (e.g. Lehmann 1975). However, we retain the terminology for
consistency with previous studies.

The use GP methods to formulate a predictive model is relatively new in
the context of modeling dynamical systems (Kocijan et al. 2005), and was
first introduced in the context ecological modeling and fisheries
management in Munch et al. (2005b). GP models have subsequently been
used to test for the presence of Allee effects (Sugeno and Munch 2013a),
estimate the maximum reproductive rate (Sugeno and Munch 2013b),
determine temporal variation in food availability (Sigourney et al.
2012), and provide a basis for identifying model-misspecification
(Thorson et al. 2014). An accessible and thorough introduction to the
formulation and use of GPs can be found in Rasmussen and Williams
(2006).

A GP is an infinite dimensional (because it considers a continuous
random variable) stochastic process for which any realization is a
multivariate normal distribution. Thus, to characterize the GP we need a
mean function and a covariance matrix. We proceed as follows.

Assuming the data $X_o$ are observed with some process noise,

\[y = f(x) + \varepsilon\]

$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP, the
predicted data $X_p$ given the observations obeys

\[f(X_p|X_o) \sim \mathcal{N}(E,C)\]
\[E = K(X_p, X_o) \left(K(X_o,X_o) - \sigma \mathbb{I} \right)  ^{-1} y\]
\[C = K(X_p, X_p) - K(X_p, X_o) K(X_o,X_o)^{-1} K(X_o, X_p)\]

Where $K$ is the Gaussian process kernel. (Compare to Eqn 2.22 of
Rasmussen and Williams (2006) for a more detailed derivation.)
Throughout, we use the radial basis function kernel:

\[ K(x,y) = \exp\left(\frac{-(x-y)^2}{2 \ell^2} \right)\]

We use vague inverse Gamma priors on both the lengthscale $\ell$ and
process noise $\sigma$ parameters, of the form

\[f(x; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{-\alpha - 1}\exp\left(-\frac{\beta}{x}\right)\]

For the $\sigma$ prior, $\alpha = $ 5 and $\beta = $ 5. For $\ell$
prior, $\alpha = $ 10 and $\beta = $ 10.

Using the simulated training data we also estimate a Gaussian process
defined by a radial basis function kernel of two parameters: $\ell$,
which gives the characteristic length-scale over which correlation
between two points in state-space decays, and $\sigma$, which gives the
scale of the process noise by which observations $Y_{t+1}$ may differ
from their predicted values $X_{t+1}$ given an observation of the
previous state, $X_t$. Munch et al. (2005a) gives an accessible
introduction to the use of Gaussian processes in providing a Bayesian
nonparametric description of the stock-recruitment relationship.

We use a Metropolis-Hastings Markov Chain Monte Carlo (Gelman et al.
(2003)) to infer posterior distributions of the two parameters of the GP
(Figure S13, code in appendix), under weakly informative Gaussian priors
(see parameters in table S5). As the posterior distributions differ
substantially from the priors (Figure S13), we can be assured that most
of the information in the posterior comes from the data rather than the
prior belief.

\subsection{The method of Gaussian Process Dynamic
Programming}\label{the-method-of-gaussian-process-dynamic-programming}

It is reasonably straight-forward to derive the harvest policy from the
estimated BP by inserting it into a SDP algorithm. The only difference
is that the uncertainty in the future state under the GP, $X_{t+1}$,
includes both process uncertainty (based on the estimation of $\sigma$)
and ``structural uncertainty'' of the posterior collection of
curves.\\(Recall that given the parameters, a parametric model is
represented by a single curve, while the GP remains a distribution of
curves).\\From the GP posteriors, we can write down the (discritized)
transition matrix representing the probability of going to each state
$X_{t+1}$ given any current state $X_t$ and any harvest $h_t$ (See the
function \texttt{gp\_transition\_matrix()} in the provided R package).
Given this transition matrix, we use the same value iteration algorithm
as in the parametric case to determine the optimal policy.

\section{Results}\label{results}

\subsection{Parametric and GP models for population
dynamics}\label{parametric-and-gp-models-for-population-dynamics}

To ensure our results are robust to our choice of parameters, we perform
our analysis over 96 different scenarios. To better understand the
process, we will first describe in detail the results of a single
scenario.

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_1.pdf}
\caption{Points show the training data of stock-size over time. Curves
show the posterior step-ahead predictions based on each of the estimated
models. Observe that all models are fitting the data reasonably well.}
\end{figure}

All models fit the observed data rather closely and with relatively
small uncertainty, as illustrated in the posterior predictive curves in
Figure 1. Figure 1 shows the training data of stock sizes observed over
time as points, overlaid with the step-ahead predictions of each
estimated model using the parameters sampled from their posterior
distributions. Each model manages to fit the observed data rather
closely. Compared to the true model most estimates appear to over-fit,
predicting fluctuations that are actually due purely to stochasticity.
All model selection criteria shown in Table 1 penalize more complex
models and show a preference for the simpler Ricker model over the more
complicated alternate stable state models (Allen and Myers). Details on
MCMC estimates for each model, traces, and posterior distributions can
be found in the appendix.

\begin{table}[ht]
\begin{center}
\begin{tabular}{rrrr}
  \hline
 & Allen & Ricker & Myers \\ 
  \hline
DIC & 50.14 & 49.45 & 50.61 \\ 
  AIC & -24.60 & -30.07 & -27.19 \\ 
  BIC & -17.85 & -25.00 & -20.44 \\ 
   \hline
\end{tabular}
\caption{Model selection scores for several common criteria all (wrongly) select the simplest model. As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion. This highlights the danger of model choice when the selected model will be used outside of the observed range of the data.}
\end{center}
\end{table}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2.pdf}
\caption{Graph of the inferred Gaussian process compared to the true
process and maximum-likelihood estimated process. Graph shows the
expected value for the function $f$ under each model. Two standard
deviations from the estimated Gaussian process covariance with (light
grey) and without (darker grey) measurement error are also shown. The
training data is also shown as black points. The GP is conditioned on
(0,0), shown as a pseudo-data point.}
\end{figure}

The mean inferred state space dynamics of each model relative to the
true model used to generate the data is shown in Figure 2, predicting
the relationship between observed stock size (x-axis) to the stock size
after recruitment the following year. Note that in contrast to the other
models shown, the expected Gaussian process corresponds to a
distribution of curves - as indicated by the gray band - which itself
has a mean shown in black. Parameter uncertainty (not shown) spreads out
the estimates further. The observed data from which each model is
estimated is also shown. The observations come from only a limited
region of state space corresponding to unharvested or weakly harvested
system. No observations occur at the theoretical optimum harvest rate or
near the tipping point.

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_3.pdf}
\caption{Outside the range of the training data (Figure 1), the true
dynamics (black dots) fall outside the uncertainty (two standard
deviations, colored bands) of the structurally incorrect parametric
models (Ricker, Myers), but inside the uncertainty predicted by the GP.
Points show the stock size simulated by the true model. Overlay shows
the range of states predicted by each model, based on the state observed
in the previous time step. The Ricker model always (wrongly) predicts
positive population growth, while the actual population shrinks in each
step as the initial condition falls below the Allee threshold of the
underlying model (Allen). Note that because it does not assume a
parametric form but instead relies more directly on the data, the GP is
both more pessimistic and more uncertain about the future state than the
parametric models.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_4.pdf}
\caption{The steady-state optimal policy (infinite boundary) calculated
under each model. Policies are shown in terms of target escapement,
$S_t$, as under models such as this a constant escapement policy is
expected to be optimal (Reed 1979).}
\end{figure}

Despite the similarities in model fits to the observed data, the
policies inferred under each model differ widely, as shown in Figure 4.
Policies are shown in terms of target escapement, $S_t$. Under models
such as this a constant escapement policy is expected to be optimal
(Reed 1979), whereby population levels below a certain size $S$ are
unharvested, while above that size the harvest strategy aims to return
the population to $S$, resulting in the hockey-stick shaped policies
shown. Only the structurally correct model (Allen model) and the GP
produce policies close to the true optimum policy (where both the
underlying model structure and parameter values are known without
error).

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_5.pdf}
\caption{In the management context, GPDP outperforms approaches based on
parametric models. Shown are 100 replicate simulations of the stock
dynamics (eq 1) under the policies derived from each of the estimated
models, as well as the policy based on the exact underlying model.}
\end{figure}

The consequences of managing 100 replicate realizations of the simulated
fishery under each of the policies estimated is shown in Figure 5. As
expected from the policy curves, the structurally correct model
under-harvests, leaving the stock to vary around it's un-fished optimum.
The structurally incorrect Ricker model over-harvests the population
passed the tipping point consistently, resulting in the immediate crash
of the stock and thus derives minimal long-term catch.

The results across replicate stochastic simulations can most easily be
compared by using the relative differences in net present value realized
by each of the model. Figure 6 shows the distribution of values realized
by each simulation shown in Figure 5. The GPDP most consistently
realizes a value close to the optimal solution, and importantly avoids
ever driving the system across the tipping point, which results in the
near-zero value cases in the parametric models.

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_6.pdf}
\caption{Histograms of the realized net present value of the fishery
over a range of simulated data and resulting parameter estimates. For
each data set, the three models are estimated as described above. Values
plotted are the averages of a given policy over 100 replicate
simulations. Details and code provided in the supplement.}
\end{figure}

\subsection{Sensitivity Analysis}\label{sensitivity-analysis}

These results are not sensitive to the modeling details of the
simulation. The GPDP estimate remains very close to the optimal solution
obtained by knowing the true model across changes to the training
simulation, noise scale, parameters or structure of the underlying
model, as seen in Figure 7. The sensitivity analysis uses a factorial
design of eight replicates of three different randomly generated
parameter sets for each of two different generating models (Allen and
Myers) over two different noise levels (0.01 and 0.05), for a total of 8
x 3 x 2 x 2 = 96 scenarios. In Figure 7, we pool results across
different random seeds (111, 222, 333), noise values of 0.01 and 0.05
(not large enough to violate the self-sustaining criterion of Reed
(1979)), and different randomly generated parameters sets for each
model, and using either the Myers or Allen as the underlying structure.
While the GPDP performs slightly closer to optimal on the smaller noise
levels, and closer on the Allen model then the Myers model, the
differences are small compared to the values observed in using the wrong
parametric model. The GPDP solution consistently avoids either failing
to harvest at all or crashing the stock, which results in the very low
values seen in parametric models in the previous figure. The sensitivity
analysis is compared against the optimal solution alone and not shown
against the estimates of the various parametric models (but see Figure
6), where convergence of the Gibbs sampling could contribute to their
poorer performance if not checked by hand for each configuration.

Changing the noise or the distance between stable and unstable points
does not impact the performance of the GP relative to the optimal
solution given the true model and true parameters. The parametric models
are more senstive to this difference. Large noise relative to the
distance between the stable and unstable point increases the chance of a
stochastic transition. More precisely, if we let $L = K - x_c$, then the
probability of a transition in a given window of time $T$ scales as

\[P(x_t < x_c | t \in T) \propto  \exp\left(\frac{L^2}{\sigma^2}\right),\]

(following the Ahrenius relationship, see Gardiner (2009) for the
derivation). As a result, the impact of using a model which
underestimates the risk of harvesting past the critical point is much
more severe, since this such a situation occurs more often. Conversely,
with large enough distance between the optimal escapement and unstable
points relative to the noise, the chance of a transition becomes
vanishingly small and all models can be modeled near-optimally. Models
that underestimate the cost incurred by stock sizes fluctuating
significantly below the optimal escapement level will not perform poorly
as long as those fluctuations are sufficently rare.

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_7.pdf}
\caption{Sensitivity Analysis. Histogram shows the ratio of the realized
net present value derived when managing under the GPDP over the optimal
value given the true model and true parameters. Values near 1 indicate
nearly optimal performance. Simulations cover a range of underlying
parameter, model, replicate, and noise levels.}
\end{figure}

\section{Discussion}\label{discussion}

Though simple mechanistically motivated models offer the greatest
potential to increase our basic understanding of ecological processes
(Geritz and Kisdi 2012, Cuddington et al. 2013), such models can be both
inaccurate and misleading when relied upon in a quantitative decision
making framework. In this paper we have tackled two aspects of
uncertainty that are both common to many ecological decision-making
problems and fundamentally challenging to existing approaches which
largely rely on parametric models:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  We do not know what the correct models are for ecological systems.
\item
  We have limited data from which to estimate the model -- in
  particular, such models may be misleading in predicting the
  probability of outcomes outside the training data.
\end{enumerate}

We have illustrated how the use of non-parametric approaches can provide
more reliable solutions in the sequential decision-making problem.

\subsubsection{Traditional model-choice approaches can be positively
misleading.}\label{traditional-model-choice-approaches-can-be-positively-misleading.}

These results illustrate that model-choice approaches would be
positively misleading -- supporting simpler models that cannot express
tipping point dynamics merely on account of them being similar. As the
data shown comes only from the basin of attraction near the unfished
equilibrium, near which all of the models are approximately linear and
approximately identical.

Model selection criteria trade off model complexity and fit to the data.
When the data come from a limited region of state-space -- as is
necessarily the case whenever there is a potential concern about tipping
point dynamics -- simpler models therefore fit just as well and tend to
be selected more foten than complex ones. This approach would be
appropriate when the dynamics can be expected to remain in the region of
the training data; for instance, if we only considered the forecasting
accuracy of the unfished population dynamics under each model.

In contrast, the decision-maker's problem of setting appropriate harvest
levels cannot exclude regions of state-space outside the observed range
when integrating over all possible decisions to find the optimal choice.
Such problems are not constrained to fisheries management but ubiquitous
across ecological decision-making and conservation where the greatest
concerns involve entering previously unobserved regions of state-space
-- whether that is the collapse of a fishery, the spread of an invasive,
or the loss of habitat.

\subsubsection{GPDP population dynamcis capture larger uncertainty in
regions where the data are
poor}\label{gpdp-population-dynamcis-capture-larger-uncertainty-in-regions-where-the-data-are-poor}

The parametric models perform worst when they propose a management
strategy outside the range of the observed data. The non-parametric
Bayesian approach, in contrast, allows a predictive model that expresses
a great deal of uncertainty about the probable dynamics outside the
observed range, while retaining very good predictive accuracy in the
range observed. The management policy dictated by the GP balance this
uncertainty against the immediate value of the harvest, and act to
stabilize the population dynamics in a region of state space in which
the predictions can be reliably reflected by the data.

\subsubsection{The role of the prior}\label{the-role-of-the-prior}

Lastly, it should be noted that outside the data, the GP reverts to the
prior, and consequently the choice of the prior can also play a
significant role in determining the optimal policy inferred by the SDP.
In the examples shown here we have selected a prior that is both
relatively uninformative (due to the broad priors placed on its
parameters $\ell$ and $\sigma$ and simple (mean zero, radial basis
function kernel). In practice, both the choice of mean and the choice of
the covariance function may be chosen to confer particular biological
properties, as well as more biologically informed priors for $\ell$ and
$\sigma$. In principle, this may allow a manager to improve the
performance of the GPDP by adding only enough additional detail as is
justified. For instance, it would be possible to use a linear or a
Ricker-shaped mean in the prior without making the much stronger
assumption that the Ricker is the structurally correct model (Sugeno and
Munch 2013a). Future research should focus on identifying criteria to
ensure the prior and the value function are chosen appropriately for the
problem at hand.

\section{Acknowledgments}\label{acknowledgments}

This work was partially supported by NOAA-IAM grant to SM and Alec
McCall and administered through the Center for Stock Assessment
Research, a partnership between the University of California Santa Cruz
and the Fisheries Ecology Division, Southwest Fisheries Science Center,
Santa Cruz, CA and by NSF grant EF-0924195 to MM and NSF grant
DBI-1306697 to CB.

\appendix
\renewcommand*{\thefigure}{S\arabic{figure}}
\renewcommand*{\thetable}{S\arabic{table}} \setcounter{figure}{0}\}
\setcounter{table}{0}\}

\section{Appendix: Model definitions and
estimation}\label{appendix-model-definitions-and-estimation}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_S1.pdf}
\caption{Traces from the MCMC estimates of the GP model show reasonable
mixing (no trend) and sampling rejection rate (no piecewise jumps)}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_S2.pdf}
\caption{Posterior distributions from the MCMC estimate of the GP model.
Prior curves shown in red.}
\end{figure}

\newpage

\subsection{Training data (Figures
1-6)}\label{training-data-figures-1-6}

Each of our models $f(S_t)$ must be estimated from training data, which
we simulate from the Allen model with parameters $r = $ 2, $K =$ 8,
$C =$ 5, and $\sigma_g =$ 0.05 for $T=$ 40 timesteps, starting at
initial condition $X_0 = $ 5.5. The training data can be seen in Figure
1.

\subsection{Training data for sensitivity
analyses}\label{training-data-for-sensitivity-analyses}

A further 96 unique training data sets are generated for the sensitivity
analysis, as described in the main text.

\subsection{Code}\label{code}

All code used in producing this analysis can be found in the R package
accompanying the paper, available from
{[}http://github.com/cboettig/nonparametric-bayes{]}. The code for all
results shown here is dynamically embedded into the manuscript using
\texttt{knitr} (Xie 2012).

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_S3.pdf}
\caption{Traces from the MCMC estimates of the Ricker model show
reasonable mixing (no trend) and sampling rejection rate (no piecewise
jumps)}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_S4.pdf}
\caption{Posteriors from the MCMC estimate of the Ricker model}
\end{figure}

\begin{longtable}[c]{@{}ccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.15\columnwidth}\centering
parameter
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
lower.bound
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
upper.bound
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering
$r$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
0.01
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
20
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$K$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
0.01
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
40
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$\sigma$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
1e-06
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
100
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Parameterization range for the uniform priors in the Ricker
model}
\end{longtable}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_S5.pdf}
\caption{Traces from the MCMC estimates of the Myers model show
reasonable mixing (no trend) and sampling rejection rate (no piecewise
jumps)}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_S6.pdf}
\caption{Posterior distributions from the MCMC estimates of the Myers
model}
\end{figure}

\begin{longtable}[c]{@{}ccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.15\columnwidth}\centering
parameter
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
lower.bound
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
upper.bound
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering
$r$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
1e-04
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
10
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$K$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
1e-04
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
40
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$\theta$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
1e-04
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
10
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$\sigma$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
1e-06
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
100
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Parameterization range for the uniform priors in the Myers
model}
\end{longtable}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_S7.pdf}
\caption{Traces from the MCMC estimates of the Allen model show
reasonable mixing (no trend) and sampling rejection rate (no piecewise
jumps)}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_S8.pdf}
\caption{Posteriors from the MCMC estimate of the Allen model}
\end{figure}

\begin{longtable}[c]{@{}ccc@{}}
\toprule\addlinespace
\begin{minipage}[b]{0.15\columnwidth}\centering
parameter
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
lower.bound
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\centering
upper.bound
\end{minipage}
\\\addlinespace
\midrule\endhead
\begin{minipage}[t]{0.15\columnwidth}\centering
$r$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
0.01
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
6
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$K$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
0.01
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
20
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$X_C$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
0.01
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
20
\end{minipage}
\\\addlinespace
\begin{minipage}[t]{0.15\columnwidth}\centering
$\sigma$
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
1e-06
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\centering
100
\end{minipage}
\\\addlinespace
\bottomrule
\addlinespace
\caption{Parameterization range for the uniform priors in the Allen
model}
\end{longtable}

Allen, D., and K. Tanner. 2005. Infusing active learning into the
large-enrollment biology class: seven strategies, from the simple to
complex. Cell biology education 4:262--8.

Athanassoglou, S., and A. Xepapadeas. 2012. Pollution control with
uncertain stock dynamics: When, and how, to be precautious. Journal of
Environmental Economics and Management 63:304--320.

Bestelmeyer, B. T., M. C. Duniway, D. K. James, L. M. Burkett, and K. M.
Havstad. 2012. A test of critical thresholds and their indicators in a
desertification-prone ecosystem: more resilience than we thought.
Ecology Letters:n/a--n/a.

Brozović, N., and W. Schlenker. 2011. Optimal management of an ecosystem
with an unknown threshold. Ecological Economics:1--14.

Burnham, K. P., and D. R. Anderson. 2002. Model Selection and
Multi-Model Inference. Page 496. Springer.

Clark, C. W. 1976. Mathematical Bioeconomics. WileyNew York.

Clark, C. W. 2009. Mathematical Bioeconomics. WileyNew York.

Clark, C. W., and G. P. Kirkwood. 1986. On uncertain renewable resource
stocks: Optimal harvest policies and the value of stock surveys. Journal
of Environmental Economics and Management 13:235--244.

Clark, C. W., and M. Mangel. 2000. Dynamic state variable models in
ecology. Oxford University PressOxford.

Courchamp, F., L. Berec, and J. Gascoigne. 2008. Allee Effects in
Ecology and Conservation. Page 256. Oxford University Press, USA.

Cressie, N., C. a Calder, J. S. Clark, J. M. {Ver Hoef}, and C. K.
Wikle. 2009. Accounting for uncertainty in ecological analysis: the
strengths and limitations of hierarchical statistical modeling.
Ecological Applications 19:553--70.

Cuddington, K. M., M. Fortin, and L. Gerber. 2013. Process-based models
are required to manage ecological systems in a changing world. Ecosphere
4:1--12.

Fischer, J., G. D. Peterson, T. a Gardner, L. J. Gordon, I. Fazey, T.
Elmqvist, A. Felton, C. Folke, and S. Dovers. 2009. Integrating
resilience thinking and optimisation for conservation. Trends in ecology
\& evolution 24:549--54.

Gardiner, C. 2009. Stochastic Methods: A Handbook for the Natural and
Social Sciences (Springer Series in Synergetics). Page 447. Springer.

Gelman, A., J. B. Carlin, H. S. Stern, and D. B. Rubin. 2003. Bayesian
Data Analysis. 2nd editions. Chapman; Hall/CRC.

Geritz, S. A. H., and E. Kisdi. 2012. Mathematical ecology: why
mechanistic models? Journal of mathematical biology 65:1411--5.

Gordon, H. 1954. The economic theory of a common-property resource: the
fishery. The Journal of Political Economy 62:124--142.

Hilborn, R. 2007. Reinterpreting the State of Fisheries and their
Management. Ecosystems 10:1362--1369.

Hilborn, R., and M. Mangel. 1997. The Ecological Detective: Confronting
Models with data. Page 330. Princeton University Press.

Hughes, T. P., C. Linares, V. Dakos, I. a van de Leemput, and E. H. van
Nes. 2013. Living dangerously on borrowed time during slow, unrecognized
regime shifts. Trends in ecology \& evolution 28:149--55.

Kocijan, J., A. Girard, B. Banko, and R. Murray-Smith. 2005. Dynamic
systems identification with Gaussian processes. Mathematical and
Computer Modelling of Dynamical Systems 11:411--424.

Lehmann, E. L. 1975. Nonparametrics: Statistical Methods Based on Ranks.
Holden-Day, IncSan Francisco.

Ludwig, D., and C. J. Walters. 1982. Optimal harvesting with imprecise
parameter estimates. Ecological Modelling 14:273--292.

Mangel, M. 2014. Stochastic Dynamic Programming Illuminates the Link
Between Environment. Bulletin of Mathematical Biology in press.

Mangel, M., and C. W. Clark. 1988. Dynamic Modeling in Behavioral
Ecology. (J. Krebs and T. Clutton-Brock, Eds.). Princeton University
PressPrinceton.

Mangel, M., 0. Fiksen, and J. Giske. 2001. Theoretical and statistical
models in natural resource management and research. Pages 57--71
\emph{in} T. M. Shenk and A. B. Franklin, editors. Modeling in natural
resource management, development, interpretation and application. Island
PressWashington DC.

Marescot, L., G. Chapron, I. Chadès, P. L. Fackler, C. Duchamp, E.
Marboutin, and O. Gimenez. 2013. Complex decisions made simple: a primer
on stochastic dynamic programming. Methods in Ecology and
Evolution:n/a--n/a.

May, R. M., J. R. Beddington, C. W. Clark, S. J. Holt, and R. M. Laws.
1979. Management of multispecies fisheries. Science (New York, N.Y.)
205:267--77.

McAllister, M. 1998. Bayesian stock assessment: a review and example
application using the logistic model. ICES Journal of Marine Science
55:1031--1060.

Munch, S. B., A. Kottas, and M. Mangel. 2005a. Bayesian nonparametric
analysis of stock-recruitment relationships. Canadian Journal of
Fisheries and Aquatic Sciences 62:1808--1821.

Munch, S. B., M. L. Snover, G. M. Watters, and M. Mangel. 2005b. A
unified treatment of top-down and bottom-up control of reproduction in
populations. Ecology Letters 8:691--695.

Myers, R. A., N. J. Barrowman, J. A. Hutchings, and A. a Rosenberg.
1995. Population dynamics of exploited fish stocks at low population
levels. Science (New York, N.Y.) 269:1106--8.

Polasky, S., S. R. Carpenter, C. Folke, and B. Keeler. 2011.
Decision-making under great uncertainty: environmental management in an
era of global change. Trends in ecology \& evolution:1--7.

R Core Team. 2013. R: A Language and Environment for Statistical
Computing. R Foundation for Statistical ComputingVienna, Austria.

Rasmussen, C. E., and C. K. I. Williams. 2006. Gaussian Processes for
Machine Learning. (Thomas Dietterich, Ed.). MIT Press,Boston.

Reed, W. J. 1979. Optimal escapement levels in stochastic and
deterministic harvesting models. Journal of Environmental Economics and
Management 6:350--363.

Roughgarden, J. E., and F. Smith. 1996. Why fisheries collapse and what
to do about it. Proceedings of the National Academy of Sciences of the
United States of America 93:5078.

Schapaugh, A. W., and A. J. Tyre. 2013. Accounting for parametric
uncertainty in Markov decision processes. Ecological Modelling
254:15--21.

Scheffer, M., J. Bascompte, W. A. Brock, V. Brovkin, S. R. Carpenter, V.
Dakos, H. Held, E. H. van Nes, M. Rietkerk, and G. Sugihara. 2009.
Early-warning signals for critical transitions. Nature 461:53--9.

Scheffer, M., S. R. Carpenter, J. A. Foley, C. Folke, and B. Walker.
2001. Catastrophic shifts in ecosystems. Nature 413:591--6.

Sethi, G., C. Costello, A. Fisher, M. Hanemann, and L. Karp. 2005.
Fishery management under multiple uncertainty. Journal of Environmental
Economics and Management 50:300--318.

Sigourney, D. B., S. B. Munch, and B. H. Letcher. 2012. Combining a
Bayesian nonparametric method with a hierarchical framework to estimate
individual and temporal variation in growth. Ecological Modelling
247:125--134.

Su, Y.-S., and Masanao Yajima. 2013. R2jags: A Package for Running jags
from R.

Sugeno, M., and S. B. Munch. 2013a. A semiparametric Bayesian method for
detecting Allee effects. Ecology 94:1196--1204.

Sugeno, M., and S. B. Munch. 2013b. A semiparametric Bayesian approach
to estimating maximum reproductive rates at low population sizes.
Ecological applications : a publication of the Ecological Society of
America 23:699--709.

Thorson, J. T., K. Ono, and S. B. Munch. 2014. A Bayesian approach to
identifying and compensating for model misspecification in population
models. Ecology 95:329--41.

Walters, C. J., and R. Hilborn. 1978. Ecological Optimization and
Adaptive Management. Annual Review of Ecology and Systematics
9:157--188.

Weitzman, M. L. 2013. A Precautionary Tale of Uncertain Tail Fattening.
Environmental and Resource Economics 55:159--173.

Williams, B. K. 2001. Uncertainty , learning , and the optimal
management of wildlife. Environmental and Ecological Statistics
8:269--288.

Worm, B., E. B. Barbier, N. Beaumont, J. E. Duffy, C. Folke, B. S.
Halpern, J. B. C. Jackson, H. K. Lotze, F. Micheli, S. R. Palumbi, E.
Sala, K. a Selkoe, J. J. Stachowicz, and R. Watson. 2006. Impacts of
biodiversity loss on ocean ecosystem services. Science (New York, N.Y.)
314:787--90.

Worm, B., R. Hilborn, J. K. Baum, T. A. Branch, J. S. Collie, C.
Costello, M. J. Fogarty, E. a Fulton, J. a Hutchings, S. Jennings, O. P.
Jensen, H. K. Lotze, P. M. Mace, T. R. McClanahan, C. Minto, S. R.
Palumbi, A. M. Parma, D. Ricard, A. a Rosenberg, R. Watson, and D.
Zeller. 2009. Rebuilding global fisheries. Science (New York, N.Y.)
325:578--85.

Xie, Y. 2012. knitr: A general-purpose package for dynamic report
generation in R.

\end{document}