### 2014-07-09 manuscripts/supplement.Rmd
fix supplement title, drop abstract

---
layout: 12pt,review
linenumbers: true
title: "Supplement for: Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
  - name: Marc Mangel
    affiliation: cstar
  - name: Stephan Munch
    affiliation: noaa
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research, 
      Department of Applied Math and Statistics, 
      University of California, Mail Stop SOE-2,
      Santa Cruz, CA 95064, USA
  - code: marc
    address: |
      Center for Stock Assessment Research, Department of Applied Math
      and Statistics, University of California,
      Mail Stop SOE-2, Santa Cruz, CA 95064, USA and Department of
      Biology, University of Bergen, Bergen, Norway 9020
  - code: noaa
    address: | 
      Southwest Fisheries Science Center, 
      National Oceanic and Atmospheric Administration, 
      110 Shaffer Road, Santa Cruz, CA 95060, USA

bibliography: components/references.bib
csl: components/ecology.csl
documentclass: components/elsarticle

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    template: components/elsarticle.latex
    keep_tex: true

---

```{r knit, include=FALSE}
library("methods")
library("rmarkdown")
render("manuscript.Rmd")
```

\appendix
\renewcommand*{\thefigure}{S\arabic{figure}}
\renewcommand*{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}

Appendix: Model definitions and estimation
===========================================

```{r figure_S1, fig.cap = "Traces from the MCMC estimates of the GP model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", dependson="plot-options"}
gp_assessment_plots$traces_plot
```

```{r figure_S2, fig.cap="Posterior distributions from the MCMC estimate of the GP model. Prior curves shown in red."}
gp_assessment_plots$posteriors_plot
```

\newpage

Training data (Figures 1-6)
----------------------------

Each of our models $f(S_t)$ must be estimated from training data, which
we simulate from the Allen model with parameters $r = $ `r p[1]`, 
$K =$ `r p[2]`, $C =$ `r p[3]`, and  $\sigma_g =$ `r sigma_g` 
for $T=$ `r Tobs` timesteps, starting at initial condition $X_0 = $ `r Xo`. 
The training data can be seen in Figure 1.

Training data for sensitivity analyses
-------------------------------------

A further 96 unique training data sets are generated for the sensitivity analysis, as described in the main text.

Code
----

All code used in producing this analysis can be found in the R package accompanying the paper, available from [http://github.com/cboettig/nonparametric-bayes]. The code for all results shown here is dynamically embedded into the manuscript using `knitr` [@knitr].

<!--
Data
----

While the data can be regenerated using the code provided, for convenience CSV files of the data shown in each graph are also provided, along with appropriate metadata written in the Ecological Metadata Language (EML).

-->

```{r figure_S3, fig.height=6, fig.cap="Traces from the MCMC estimates of the Ricker model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
plot_ricker_traces
```

```{r figure_S4, fig.cap="Posteriors from the MCMC estimate of the Ricker model", fig.width=6, fig.height=4}
ggplot(ricker_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(ricker_posteriors, "components/data/ricker_posteriors.csv")
```

```{r Table S1, results = "asis"}
pander::pandoc.table(ricker_priors_xtable,
  caption = "Parameterization range for the uniform priors in the Ricker model")
```

```{r figure_S5, fig.height=6, fig.cap="Traces from the MCMC estimates of the Myers model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
plot_myers_traces
```

```{r figure_S6, fig.cap="Posterior distributions from the MCMC estimates of the Myers model", fig.width=6, fig.height=6}
ggplot(myers_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(myers_posteriors, "components/data/myers_posteriors.csv")
```

```{r TableS2, results="asis"}
pander::pandoc.table(myers_priors_xtable,
           caption = "Parameterization range for the uniform priors in the Myers model")
```

```{r figure_S7, fig.height=6, fig.cap="Traces from the MCMC estimates of the Allen model show reasonable mixing (no trend) and sampling rejection rate (no piecewise jumps)", fig.width=6}
plot_allen_traces
```

```{r figure_S8, fig.cap="Posteriors from the MCMC estimate of the Allen model", fig.width=6, fig.height=6}
ggplot(allen_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
  facet_wrap(~ variable, scale="free", ncol=2)
write.csv(allen_posteriors, "components/data/allen_posteriors.csv")
```

```{r TableS3, results = "asis"}
pander::pandoc.table(allen_priors_xtable,
  caption = "Parameterization range for the uniform priors in the Allen model")
```

```{r sensitivity-calc}
source("components/sensitivity.R")

models <- c("Myers","Allen")

parameters <- list(
  Myers = list(
    c(r=1.5 + rnorm(1, 0, 1), theta=2 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1)),
    c(r=1.5 + rnorm(1, 0, 1), theta=2.5 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1))),
  Allen = list(
    c(r=2 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1), C=4 + rnorm(1, 0, 1)),
    c(r=2 + rnorm(1, 0, 1), K=10 + rnorm(1, 0, 1), C=4 + rnorm(1, 0, 1)))
)
nuisance_pars <- c("sigma_g")
nuisance_values <- list(sigma_g = c(0.01, 0.05))
```

```{r allensets}
model <- "Allen"
allen1.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1111, 2222, 3333))
allen2.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1111, 2222, 3333))
allen1.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1111, 2222, 3333))
allen2.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1111, 2222, 3333))
```

```{r myerssets}
model <- "Myers"
Myers1.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1111, 2222, 3333))
Myers2.01 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[1]), 
                   seed=c(1111, 2222, 3333))
Myers1.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[1]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1111, 2222, 3333))
Myers2.05 <- sensitivity(model, 
                   parameters = parameters[[model]][[2]], 
                   nuisance = c(sigma_g = nuisance_values$sigma_g[2]), 
                   seed=c(1111, 2222, 3333))

## Assemble into data.frame
allen_dat <- rbind(allen1.01, allen1.05, allen2.01, allen2.05) 
m <- rbind(Myers1.01, Myers1.05, Myers2.01, Myers2.05)
myers_dat <- m[c(1:2,4,3,5:8)]
names(myers_dat) <- names(allen_dat)
model_dat <- rbind(allen_dat, myers_dat)
dat <- model_dat
dat$pars.r <- factor(dat$pars.r, labels=c("A", "B", "C", "D"))
dat <- dat[c(1:2,5:6, 8, 7)]
dat$noise <- factor(dat$noise)
names(dat) <- c("model", "parameters", "replicate", "simulation", "noise", "value")

## Extract acutal parameter values corresponding to each parameter set
p1 = as.numeric(levels(factor(model_dat$pars.r)))
p2 = as.numeric(levels(factor(model_dat$pars.K)))
p3 = as.numeric(levels(factor(model_dat$pars.C)))

set.A = c(r = p1[1], K = p2[1], theta = p3[1])
set.B = c(r = p1[2], K = p2[2], theta = p3[2])
set.C = c(r = p1[3], K = p2[3], C = p3[3])
set.D = c(r = p1[4], K = p2[4], C = p3[4])
AllenParameterSets <- rbind(set.A, set.B)
MyersParameterSets <- rbind(set.C, set.D)

sensitivity_dat <- dat
```

```{r figure_S9, fig.height=6, fig.width=10, dependson=c("sensitivity-calc", "export-data"), fig.cap="Sensitivity Analysis.  Histograms shows the ratio of the realized net present value derived when managing under the GPDP over the optimal value given the true model and true parameters. Values of 1 indicate optimal performance. Columns indicate different models, rows different noise levels, and colors indicate the parameter set used. Grouped over stochastic replicates applying the contol policy and stochastic replicates of training data generated from the model indicated, see raw data for details. Randomly chosen parameter values for the models shown in tables below."}
ggplot(sensitivity_dat) + 
  geom_histogram(aes(value, fill=parameters)) + 
  xlim(0,1.0) + 
  theme_bw() + 
  xlab("value as fraction of the optimal") + 
  facet_grid(noise~model)
write.csv(sensitivity_dat, "components/data/sensitivity_dat.csv") 
```

```{r AllenSetsTable, results="asis"}
pandoc.table(AllenParameterSets, caption="Randomly chosen parameter sets for the Allen models in Figure S9." )
```
```{r MyersSetsTable, results="asis"}
pandoc.table(MyersParameterSets, caption="Randomly chosen parameter sets for the Myers models in Figure S9." )
```

```{r unlink, include=FALSE}
unlink("ricker_process.bugs")
unlink("allen_process.bugs")
unlink("myers_process.bugs")
```