### 2014-12-27 manuscripts/manuscript.Rmd
formatting, rebuild pdfs

---
layout: 12pt,review
linenumbers: true
title: "Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
  - name: Marc Mangel
    affiliation: cstar
  - name: Stephan Munch
    affiliation: noaa
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research, 
      Department of Applied Math and Statistics, 
      University of California, Mail Stop SOE-2,
      Santa Cruz, CA 95064, USA
  - code: noaa
    address: | 
      Southwest Fisheries Science Center, 
      National Oceanic and Atmospheric Administration, 
      110 Shaffer Road, Santa Cruz, CA 95060, USA

abstract: |
  Model uncertainty and limited data are fundamental challenges to
  robust management of human intervention in a natural system.
  These challenges are acutely highlighted by concerns that
  many ecological systems may contain tipping points, such as
  Allee population sizes.  Before a collapse, we do not know
  where the tipping points lie, if they exist at all.  Hence,
  we know neither a complete model of the system dynamics nor
  do we have access to data in some large region of state-space
  where such a tipping point might exist.  We illustrate how
  a Bayesian Non-Parametric (BNP) approach using a Gaussian
  Process (GP) prior provides a flexible representation of this
  inherent uncertainty. We embed GPs in a Stochastic Dynamic
  Programming (SDP) framework in order to make robust management
  predictions with both model uncertainty and limited data. We
  use simulations to evaluate this approach as compared with the
  standard approach of using model selection to choose from a set
  of candidate models.  We find that model selection erroneously
  favors models without tipping points -- leading to harvest
  policies that guarantee extinction. The GPDP performs nearly as
  well as the true model and significantly outperforms standard
  approaches. We illustrate this using examples of simulated
  single-species dynamics, where the standard model selection
  approach should be most effective, and find that it still
  fails to account for uncertainty appropriately and leads to
  population crashes, while management based on the GPDP does
  not, since it does not underestimate the uncertainty outside
  of the observed data.

keywords:
  - Bayesian
  - Structural Uncertainty
  - Nonparametric 
  - Optimal Control
  - Decision Theory
  - Gaussian Processes
  - Fisheries Management

bibliography: components/references.bib
csl: components/ecology.csl
documentclass: components/elsarticle

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    template: components/elsarticle.latex
    keep_tex: true

---

```{r caching, include=FALSE}
library("methods")
library("knitr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)

# PDF-based figures
opts_chunk$set(dev='pdf')
```

```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

opts_chunk$set(dev='pdf', dev.args=c(version="1.3"))

```

```{r libraries}
library("nonparametricbayes") 
library("pdgControl")

library("ggplot2") 
library("reshape2")
library("plyr")
library("data.table")

library("R2jags")
library("emdbook") # for as.mcmc.bugs (?)
library("coda")  # for as.mcmc

library("modeest")
library("MASS")
library("pander")

library("cboettigR")

opts_chunk$set(fig.width=4, fig.height=3, echo=FALSE)
theme_set(theme_bw(base_size=12))
toggle = "hide" # results argument
```

```{r posterior-mode}
library(modeest)
posterior.mode <- function(x) {
  mlv(x, method="shorth")$M
}
```
```{r stateeq}
f <- RickerAllee
p <- c(2, 8, 5)
K <- 10  # approx, a li'l' less
allee <- 5 # approx, a li'l' less
```
```{r sdp-pars, dependson="stateeq"}
sigma_g <- 0.05
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1
x_grid <- seq(0, 1.5 * K, length=50)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 50  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
Xo <-  allee+.5# observations start from
x0 <- K # simulation under policy starts from
Tobs <- 40
MaxT = 1000 # timeout for value iteration convergence
```
```{r obs, dependson="sdp-pars", fig.keep='none'}
  set.seed(1234)
  #harvest <- sort(rep(seq(0, .5, length=7), 5))
  x <- numeric(Tobs)
  x[1] <- Xo
  nz <- 1
  for(t in 1:(Tobs-1))
    x[t+1] = z_g() * f(x[t], h=0, p=p)
  obs <- data.frame(x = c(rep(0,nz), 
                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
                    y = c(rep(0,nz), 
                          x[2:Tobs]))
raw_plot <- ggplot(data.frame(time = 1:Tobs, x=x), aes(time,x)) + geom_line()
```

```{r mle, dependson="obs"}
set.seed(12345)
estf <- function(p){ 
    mu <- f(obs$x,0,p)
    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
}
par <- c(3, 
         9, 
         6, 
         0.06)
o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
f_alt <- f
p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
sigma_g_alt <- as.numeric(o$par[4])

est <- list(f = f_alt, p = p_alt, sigma_g = sigma_g_alt, mloglik=o$value)

```
```{r mle-output, dependson="mle", results=toggle}
true_means <- sapply(x_grid, f, 0, p)
est_means <- sapply(x_grid, est$f, 0, est$p)
```

```{r gp-priors}
s2.p <- c(5,5)  
d.p = c(10, 1/0.1)
```
```{r gp, dependson=c("gp-priors", "obs")}
gp <- gp_mcmc(obs$x, y=obs$y, n=1e5, s2.p = s2.p, d.p = d.p)
gp_dat <- gp_predict(gp, x_grid, burnin=1e4, thin=300)
```
```{r gp_traces_densities, dependson="gp"}
gp_assessment_plots <- summary_gp_mcmc(gp, burnin=1e4, thin=300)
```
```{r gp-output, dependson="gp", results=toggle}
tgp_dat <- 
    data.frame(  x = x_grid, 
                 y = gp_dat$E_Ef, 
                 ymin = gp_dat$E_Ef - 2 * sqrt(gp_dat$E_Vf), 
                 ymax = gp_dat$E_Ef + 2 * sqrt(gp_dat$E_Vf) )
```

```{r jags-setup}
y <- x 
N <- length(x);
jags.data <- list("N","y")
n.chains <- 6
n.iter <- 1e6
n.burnin <- floor(10000)
n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
n.update <- 10

```
```{r common-priors}
stdQ_prior_p <- c(1e-6, 100)
stdR_prior_p <- c(1e-6, .1)
stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])

```
```{r allen-model}
K_prior_p <- c(0.01, 20.0)
r0_prior_p <- c(0.01, 6.0)
theta_prior_p <- c(0.01, 20.0)

bugs.model <- 
paste(sprintf(
"model{
  K     ~ dunif(%s, %s)
  r0    ~ dunif(%s, %s)
  theta ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)", 
  K_prior_p[1], K_prior_p[2],
  r0_prior_p[1], r0_prior_p[2],
  theta_prior_p[1], theta_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2]),

  "
  iQ <- 1 / (stdQ * stdQ);
  y[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K)* (y[t] - theta) / K 
    y[t+1] ~ dlnorm(mu[t], iQ) 
  }
}")
writeLines(bugs.model, "allen_process.bugs")

```
```{r allen-priors, dependson="common-priors"}
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
r0_prior <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
theta_prior <- function(x) dunif(x, theta_prior_p[1], theta_prior_p[2])
par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
                    r0 = r0_prior, theta = theta_prior,
                    stdQ = stdQ_prior)

```

```{r allen_priors_xtable, include=TRUE, echo=FALSE}
allen_priors_xtable <- data.frame(parameter = c("$r$", "$K$", "$X_C$", "$\\sigma$"),
           "lower bound" = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
           "upper bound" = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2]))
```

```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results=toggle}
jags.params=c("K","r0","theta","stdQ") # be sensible about the order here
jags.inits <- function(){
  list("K"= 10 * rlnorm(1,0, 0.1),
       "r0"= 1 * rlnorm(1,0, 0.1) ,
       "theta"=   5 * rlnorm(1,0, 0.1) , 
       "stdQ"= abs( 0.1 * rlnorm(1,0, 0.1)),
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

set.seed(1234)
# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
allen_jags <- do.call(jags, list(data=jags.data, inits=jags.inits, 
                                      jags.params, n.chains=n.chains, 
                                      n.iter=n.iter, n.thin=n.thin, 
                                      n.burnin=n.burnin, 
                                      model.file="allen_process.bugs"))

# Run again iteratively if we haven't met the Gelman-Rubin convergence criterion
recompile(allen_jags) # required for parallel
allen_jags <- do.call(autojags, 
											list(object=allen_jags, n.update=n.update, 
                           n.iter=n.iter, n.thin = n.thin))

```
```{r allen-traces, dependson="allen-mcmc"}
tmp <- lapply(as.mcmc(allen_jags), as.matrix) # strip classes the hard way...
allen_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
names(allen_posteriors) = c("index", "variable", "value", "chain")
plot_allen_traces <- ggplot(allen_posteriors) + geom_line(aes(index, value)) + 
  facet_wrap(~ variable, scale="free", ncol=1)

```
```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
allen_priors <- ddply(allen_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})
plot_allen_posteriors <- ggplot(allen_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
#  geom_line(data=allen_priors, aes(x=value, y=density), col="red") +  
  facet_wrap(~ variable, scale="free", ncol=3)

```
```{r allen-output, dependson=c("posterior-mode", "allen-traces"), results=toggle}
A <- allen_posteriors
A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
pardist <- acast(A, index ~ variable)
bayes_coef <- apply(pardist,2, posterior.mode) 
bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"], bayes_coef["theta"])) # parameters formatted for f
allen_f <- function(x,h,p) unname(RickerAllee(x,h, unname(p[c("r0", "K", "theta")])))
allen_means <- sapply(x_grid, f, 0, bayes_pars)
bayes_pars
head(pardist)

```

```{r ricker-model}
K_prior_p <- c(0.01, 40.0)
r0_prior_p <- c(0.01, 20.0)
bugs.model <- 
paste(sprintf(
"model{
  K    ~ dunif(%s, %s)
  r0    ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)", 
  K_prior_p[1], K_prior_p[2],
  r0_prior_p[1], r0_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2]),
  "
  iQ <- 1 / (stdQ * stdQ);
  y[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- log(y[t]) + r0 * (1 - y[t]/K) 
    y[t+1] ~ dlnorm(mu[t], iQ) 
  }
}")
writeLines(bugs.model, "ricker_process.bugs")

```
```{r ricker-priors, dependson="common-priors"}
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
r0_prior <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
                   r0 = r0_prior, stdQ = stdQ_prior)

```

```{r ricker_priors_xtable, include=TRUE, echo=FALSE}
ricker_priors_xtable <- data.frame(
  parameter = c("$r$", "$K$", "$\\sigma$"),
  "lower bound" = c(r0_prior_p[1], K_prior_p[1], stdQ_prior_p[1]),
  "upper bound" = c(r0_prior_p[2], K_prior_p[2], stdQ_prior_p[2]))
```

```{r ricker-mcmc, dependson="ricker-model", results=toggle}
jags.params=c("K","r0", "stdQ")
jags.inits <- function(){
  list("K"= 10 * rlnorm(1,0,.5),
       "r0"= rlnorm(1,0,.5),
       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345) 
ricker_jags <- do.call(jags, 
                       list(data=jags.data, inits=jags.inits, 
                            jags.params, n.chains=n.chains, 
                            n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
                            model.file="ricker_process.bugs"))
recompile(ricker_jags)
ricker_jags <- do.call(autojags, 
                       list(object=ricker_jags, n.update=n.update, 
														n.iter=n.iter, n.thin = n.thin, 
														progress.bar="none"))

```
```{r ricker-traces, dependson="ricker-mcmc"}
tmp <- lapply(as.mcmc(ricker_jags), as.matrix) # strip classes the hard way...
ricker_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
names(ricker_posteriors) = c("index", "variable", "value", "chain")
plot_ricker_traces <- ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
  facet_wrap(~ variable, scale="free", ncol=1)
```
```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors")}
ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})
# plot posterior distributions
plot_ricker_posteriors <- ggplot(ricker_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
#  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") +  # don't plot priors 
  facet_wrap(~ variable, scale="free", ncol=2)
```
```{r ricker-output, dependson=c("posterior-mode", "ricker-traces"), results=toggle}
A <- ricker_posteriors
A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
ricker_pardist <- acast(A, index ~ variable)
bayes_coef <- apply(ricker_pardist,2, posterior.mode) 
ricker_bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["K"]))
ricker_f <- function(x,h,p){
  sapply(x, function(x){ 
    x <- pmax(0, x-h) 
    pmax(0, x * exp(p["r0"] * (1 - x / p["K"] )) )
  })
}
ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
head(ricker_pardist)
ricker_bayes_pars

```

```{r myers-model}
r0_prior_p <- c(.0001, 10.0)
theta_prior_p <- c(.0001, 10.0)
K_prior_p <- c(.0001, 40.0)
bugs.model <- 
paste(sprintf(
"model{
  r0    ~ dunif(%s, %s)
  theta    ~ dunif(%s, %s)
  K    ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)", 
  r0_prior_p[1], r0_prior_p[2],
  theta_prior_p[1], theta_prior_p[2],
  K_prior_p[1], K_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2]),

  "
  iQ <- 1 / (stdQ * stdQ);

  y[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- log(r0)  + theta * log(y[t]) - log(1 + pow(abs(y[t]), theta) / K)
    y[t+1] ~ dlnorm(mu[t], iQ) 
  }
}")
writeLines(bugs.model, "myers_process.bugs")

```
```{r myers-priors}
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
r_prior     <- function(x) dunif(x, r0_prior_p[1], r0_prior_p[2])
theta_prior <- function(x) dunif(x, theta_prior_p[1], theta_prior_p[2])
par_priors <- list( deviance = function(x) 0 * x, K = K_prior,
                    r0 = r_prior, theta = theta_prior, 
                    stdQ = stdQ_prior)

```

```{r myers_prior_xtable, include=TRUE, echo=FALSE} 
myers_priors_xtable <- data.frame(parameter = c("$r$", "$K$", "$\\theta$", "$\\sigma$"),
           "lower bound" = c(r0_prior_p[1], K_prior_p[1], theta_prior_p[1], stdQ_prior_p[1]),
           "upper bound" = c(r0_prior_p[2], K_prior_p[2], theta_prior_p[2], stdQ_prior_p[2]))
```

```{r myers-mcmc, dependson="myers-model", results=toggle}
jags.params=c("r0", "theta", "K", "stdQ")
jags.inits <- function(){
  list("r0"= 1 * rlnorm(1,0,.1), 
       "K"=    10 * rlnorm(1,0,.1),
       "theta" = 1 * rlnorm(1,0,.1),  
       "stdQ"= sqrt(0.2) * rlnorm(1,0,.1),
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345)
myers_jags <- do.call(jags, 
                      list(data=jags.data, inits=jags.inits, 
													 jags.params, n.chains=n.chains, 
													 n.iter=n.iter, n.thin=n.thin,
                           n.burnin=n.burnin, 
                           model.file="myers_process.bugs"))
recompile(myers_jags)
myers_jags <- do.call(autojags, 
                      list(myers_jags, n.update=n.update, 
                           n.iter=n.iter, n.thin = n.thin, 
                           progress.bar="none"))

```
```{r myers-traces, dependson="myers-mcmc"}
tmp <- lapply(as.mcmc(myers_jags), as.matrix) # strip classes
myers_posteriors <- melt(tmp, id = colnames(tmp[[1]])) 
names(myers_posteriors) = c("index", "variable", "value", "chain")
plot_myers_traces <- ggplot(myers_posteriors) + 
  geom_line(aes(index, value)) + # priors, need to fix order though
  facet_wrap(~ variable, scale="free", ncol=1)

```
```{r myers-posteriors, dependson="myers-traces"}
par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})
plot_myers_posteriors <- ggplot(myers_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity") +
#  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") +  # Whoops, these are misaligned. see table instead 
  facet_wrap(~ variable, scale="free", ncol=3)

```
```{r myers-output, dependson=c("posterior-mode", "myers-traces"), results=toggle}
A <- myers_posteriors
A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
myers_pardist <- acast(A, index ~ variable)
bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
myers_bayes_pars <- unname(c(bayes_coef["r0"], bayes_coef["theta"], bayes_coef["K"]))
myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
myers_f <- function(x,h,p) Myer_harvest(x, h, p[c("r0", "theta", "K")])
head(myers_pardist)
myers_bayes_pars

```

```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
models <- data.frame(x=x_grid, 
										 GP=tgp_dat$y, 
										 True=true_means, 
                     MLE=est_means, 
										 Ricker=ricker_means, 
                     Allen = allen_means,
                     Myers = myers_means)
models <- melt(models, id="x")
# some labels
names(models) <- c("x", "method", "value")
# labels for the colorkey too
model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")
colorkey=cbPalette
names(colorkey) = model_names 
step_ahead <- function(x, f, p){
  h = 0
  x_predict <- sapply(x, f, h, p)
  n <- length(x_predict) - 1
  y <- c(x[1], x_predict[1:n])
  y
}
```
```{r gp-opt, dependson="gp-output"}
matrices_gp <- gp_transition_matrix(gp_dat$Ef_posterior, gp_dat$Vf_posterior, x_grid, h_grid) 
opt_gp <- value_iteration(matrices_gp, x_grid, h_grid, MaxT, xT, profit, delta, reward)
```
```{r mle-opt, dependson="mle-output"}
matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
matrices_estimated <- f_transition_matrix(est$f, est$p, x_grid, h_grid, est$sigma_g)
opt_estimated <- value_iteration(matrices_estimated, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

```
```{r allen-opt, dependson="allen-output"}
matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
opt_allen <- value_iteration(matrices_allen, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```
```{r ricker-opt, dependson="ricker-output"}
matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
opt_ricker <- value_iteration(matrices_ricker, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```
```{r myers-opt, dependson="myers-output"}
matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

```
```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_ricker$D, Allen = opt_allen$D, Myers = myers_alt$D)
colorkey=cbPalette
names(colorkey) = names(OPT) 
```
```{r sims, dependson="assemble-opt"}
sims <- lapply(OPT, function(D){
  set.seed(1)
  lapply(1:100, function(i) 
    ForwardSimulate(f, p, x_grid, h_grid, x0, D, z_g, profit=profit, OptTime=OptTime)
  )
})
# turn the list into a data.frame
keep.as.columns <- names(sims[[1]][[1]])
dat <- melt(sims, id=keep.as.columns)
sims_data <- data.table(dat)
setnames(sims_data, c("L1", "L2"), c("method", "reps")) 
# Legend in original ordering please, not alphabetical: 
sims_data$method = factor(sims_data$method, ordered=TRUE, levels=names(OPT))
```

```{r profits, dependson="sims", results=toggle}
Profit <- sims_data[, sum(profit), by=c("reps", "method")]
tmp <- dcast(Profit, reps ~ method)
#tmp$Allen <- tmp[,"Allen"] + rnorm(dim(tmp)[1], 0, 1) # jitter for plotting
tmp <- tmp / tmp[,"True"]
tmp <- melt(tmp[2:dim(tmp)[2]])
actual_over_optimal <-subset(tmp, variable != "True")
```

```{r dic_calc, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output"), include=FALSE, echo=FALSE}
dic.dic <- function (x) sum(x$deviance) +  sum(x[[2]])
recompile(allen_jags)
allen_dic <- dic.dic(dic.samples(allen_jags$model, n.iter=1000, type="popt"))
recompile(ricker_jags)
ricker_dic <- dic.dic(dic.samples(ricker_jags$model, n.iter=1000, type="popt"))
recompile(myers_jags)
myers_dic <- dic.dic(dic.samples(myers_jags$model, n.iter=1000, type="popt"))
dictable <- data.frame(Allen = allen_dic + 2*length(bayes_pars), 
                       Ricker = ricker_dic + 2*length(ricker_bayes_pars),
                       Myers = myers_dic + 2*length(myers_bayes_pars), 
                       row.names = c("DIC"))
```

```{r sensitivity-trends}
source("components/sensitivity-trends.R") # load the function for looping

sigmas <- c(0.01, 0.05, 0.1)
allee <- c(1,2,3,4,5)

vary_sigma <- lapply(sigmas, function(s) try(sensitivity_trends(4, sigma = s, seed=c(111,222,333))))
vary_allee <-  lapply(allee, function(a) try(sensitivity_trends(a, sigma = sigma_g, seed=c(111,222,333))))

vary_sigma <- melt(vary_sigma, id=names(vary_sigma[[1]]))
vary_allee <- melt(vary_allee, id=names(vary_allee[[1]]))

```

```{r refresh-plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE, cache=FALSE}
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
opts_chunk$set(dev='pdf', dev.args=c(version="1.4"))
opts_chunk$set(fig.width=4, fig.height=3, echo=FALSE)
theme_set(theme_bw(base_size=12))
```

Introduction
============

Decision making under uncertainty is a ubiquitous challenge in the
management of human intervention in natural resources and conservation.
Decision-theoretic approaches provide a framework to determine the
best sequence of actions in face of uncertainty, but only when that
uncertainty can be meaningfully quantified [@Fischer2009].  Over the last
four decades (beginning with @Clark1976, @Clark2009 and @Walters1978)
dynamic optimization methods, particularly Stochastic Dynamic Programming
(SDP), have become increasingly important as a means of understanding
how to manage human intervention into natural systems.  Simultaneously,
there has been increasing recognition of the importance of multiple steady
states or 'tipping points' [@Scheffer2001; @Scheffer2009; @Polasky2011]
in ecological systems.

We develop a novel approach to address these concerns in the context
of fisheries; although the challenges and methods are germane to other
problems of conservation or natural resource exploitation.  Economic
value and ecological concern have made marine fisheries the crucible for
much of the founding work on management under uncertainty [@Gordon1954;
@Clark1976; @Clark2009; @Reed1979; @May1979; @Ludwig1982].

Even if we know the proper deterministic description of the biological
system, there is intrinsic stochasticity in biological dynamics,
measurements, and implementation of policy [_e.g._ @Reed1979; @Clark1986;
@Roughgarden1996; @Sethi2005].  We may also lack knowledge about the
parameters of the biological dynamics [parametric uncertainty, _e.g._
@Ludwig1982; @Hilborn1997; @McAllister1998; @Schapaugh2013], or even
not know which model is  proper description of the system [structural
uncertainty, _e.g._ @Williams2001; @Cressie2009;  @Athanassoglou2012].
Of these, the latter is generally the hardest to quantify. Typical
approaches confront the data with a collection of models, assuming that
the true dynamics (or reasonable approximation) is among the collection
and then use model choice or model averaging to arrive at a conclusion
[@Williams2001; @Athanassoglou2012; @Cressie2009].  Even setting aside
other concerns (see @Cressie2009), these approaches are unable to describe
uncertainty outside the observed data range.

Structural  uncertainty is particularly insidious when we try to
predict outside of the range of observed data [@Mangel2001] because we
are extrapolating into unknown regions.  In management applications,
this extrapolation uncertainty is particularly important since (a)
management involves considering actions that may move the system
outside the range of observed behavior, and (b) the decision tools
(optimal control theory, SDP) rely on both reasonable estimates of the
expected outcomes and on the weights given to those outcomes [_e.g._
@Weitzman2013].   Thus characterizing uncertainty is as important as
characterizing the expected outcome.

Tipping points in ecological dynamics [@Scheffer2001; @Polasky2011]
highlight this problem because precise models are not available and data
are limited such as around  high stock levels or an otherwise desirable
state. With perfect information, one would know just how far a system
could be pushed before crossing the tipping point, and management
would be simple.  But we face imperfect models and limited data and,
with tipping points, even small errors can have very large consequences,
as we shall illustrate later.  Because intervention may be too late once
a tipping point has been crossed (but see @Hughes2013), management is
often concerned with avoiding tipping points before any data about them
are available.

The dual concerns of model uncertainty and incomplete data create
a substantial challenge to existing decision-theoretic approaches
[@Brozovic2011].   We illustrate how  Stochastic Dynamic Programming
(SDP) [@Mangel1988; @Marescot2013] can be implemented using a Bayesian
Non-Parametric (BNP) model of population dynamics [@Munch2005a].  The BNP
method has two distinct advantages. First, using a BNP model sidesteps
the need for an accurate model-based description of the system dynamics.
Second, a BNP model better reflects uncertainty when extrapolating beyond
the observed data.  This is crucial to providing  robust decision-making
when the correct model is not known (as is almost always true).  [We use
_robust_ to characterize approaches that provide nearly optimal solutions
without being sensitive to the choice of the (unknown) underlying model.]

This paper is the first ecological application of the
SDP without an _a priori_  model of the
underlying dynamics.  Unlike parametric approaches that can only
reflect uncertainty in parameter estimates, the BNP method provides a
broader representation of uncertainty, including uncertainty beyond the observed data. 
We will show that Gaussian Process Dynamic Programming
(GPDP)  allows us to find robust management solutions in face of limited
data without knowing  the correct model structure.

For comparisons, we consider the performance of management based on
GPDP against management policies derived under several alternative
parametric models [@Reed1979; @Ludwig1982; @Mangel1988].  Rather than
compare models in terms of best fit to data, we compare model performance
in the concrete terms of the decision-maker's objectives.

Approach and Methods
====================

We first describe the requirements of dynamic optimization for the
management of human intervention in natural resource systems. After that
we describe three parametric models for population dynamics and the Gaussian
Process (GP)[^1] description of population dynamics. All computer code
used here has been embedded in the manuscript sources (see @Xie_2013),
and an implementation of the GPDP approach is provided as an accompanying R
package.  Source code, R package and the CSV data files corresponding each
figure are archived in the supplement [@Boettiger_2014].

[^1]: We abbreviate Gaussian Process as GP, which refers to 
the statistical model we use to approximate the population dynamics, and we use the
term Gaussian Process Dynamic Programming (GPDP), to refer to the _use_ of a
GP as the underlying process model when solving a Dynamic Programming equation.
Hence we will refer to the models as: GP, Ricker, Allen, etc, and the novel method we put forward here as GPDP.

### Requirements of dynamic optimization

Dynamic optimization requires characterizing the dynamics of a state
variable (or variables), a control action, and a value function.
For simplicity, we consider only a single state variable.  This is
a best-case scenario for the parametric models because we simulate
underlying dynamics from one of the three parametric models, whereas
in the natural world we never know the "true" model.  In addition, by
choosing one-dimensional models with just a few parameters, we limit
the chance that poor performance will be due to inability to estimate
parameters accurately, something that becomes a more severe problem for
higher-dimensional parametric models. Finally, the parametric models we
consider are commonly used in modeling stock-recruitment dynamics or to
model sudden transitions between alternative stable states.

We let $X(t)$ denote the size (numbers or biomass) of the focal population
at time $t$ and assume that in the absence of take its dynamics are:

\begin{equation}
X(t+1) = Z(t) f(X(t), \mathbf{p}) \label{eq1} 
\end{equation}

Where Z(t) is log-normally distributed process stochasticity [@Reed1979]
and $\mathbf{p}$ is a vector of parameters to be estimated from the
data.  We describe the three choices for $f(X(t),\mathbf{p})$ in the
next section.  The control action is a harvest or take, $h(t)$, measured
in the same units as $X$, at time $t$.  Thus, in the presence of take,
the population size on the right hand side of Eqn 1 is replaced by $S(t)
= X(t) - h(t)$.

To construct the value function, we consider a return when $X(t) = x(t)$
and harvest $h(t) = h$ denoted as the reward, $R(x(t), h)$. For example,
if the return is the harvest at time $t$, then $R(x(t), h(t)) = \min(x(t),
h(t))$. We assume that future harvests are discounted relative to current
ones at a constant rate of discount $\delta$ and ask for the harvest
policy that maximizes total discounted harvest between the current time
$t$ and a final time $T$. That is, we seek to maximize over choices of
harvest $E [ \sum_{t = 0}^{T}  R(X(t), h(t), t) \delta^t]$, where the
state dynamics are given by Eqn 1 and $E$ denotes the expectation over
future population states.

In order to find that policy, we introduce the value function $V(x(t),
t)$ representing the total discounted catch from time $t$ onwards given
that $X(t) = x(t)$.  This value function satisfies an equation of SDP
[@Mangel1988; @Clark2000; @Clark2009 ;@Mangel2014],

\begin{equation}
V(x(t), t) = \max_{h(t)} \lbrace R(h(t), x(t)) + \delta \cdot \mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] \rbrace
\end{equation}

where expectation is taken over all possible values of the next state,
$X(t+1)$, and maximized over all possible choices of harvest, $h(t)$.
That is, at time $t$, when population size is $x(t)$ and harvest $h(t)$
is applied, the immediate return is $R(h(t), x(t))$. When the sole
source of uncertainty is the process stochasticity term, $Z$, and thus
the expectation in Eqn 2 is equivalent to taking expectations over
$Z(t)$. That is

\begin{equation}
\mathbf{\mathrm{E}}_{X(t+1)} \left[ V( X(t+1),t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{Z(t)} \left[V( Z(t) f(x(t) - h(t))|\mathbf{p}), t+1 | x(t), h(t) \right]
\end{equation}

where the population size after the take is $x(t) - h(t)$, which is
then translated into $X(t+1)$ by Eqn 1 (that is, we replace $X(t+1)$
by $Z(t) f(x(t)-h(t)|\mathbf{p})$).

When the parameters governing the dynamics are also uncertain, we take
the expectation over the posterior distribution for the parameters:

\begin{equation}
\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{\mathbf{p}|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z(t) | \mathbf{p}, \mathrm{data}} \left[ V(Z(t) f(x(t) - h(t)|\mathbf{p}), t)  \right] \}
\end{equation}

When the underlying population dynamics are unknown (the case of
structural uncertainty), the function $f$ itself is uncertain and
the expectation for the next state includes uncertainty in $f$ as
well. That is

\begin{equation}
\mathbf{\mathrm{E}}_{X(t+1)} \left[ V(X(t+1), t+1) | x(t), h(t) \right] = \mathbf{\mathrm{E}}_{\mathbf{p}|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z(t) | \mathbf{p}, \mathrm{data}} \left[ V( Z(t) f(x(t) - h(t)| \mathbf{p}), t) \right] \}
\end{equation}

We consider the finite time problem with $T=$ `r MaxT`, which we
solve using the standard value iteration algorithm [see @Mangel1988;
@Clark2000].

### Parametric Models ###

We consider three candidate parametric models for the population dynamics:
The Ricker model, the Allen model [@Allen2005], and the Myers model
[@Myers1995], Eqns \eqref{ricker}-\eqref{myers}. In all three, we let
$K$ denote the carrying capacity and $r$ the maximum per capita growth
rate. The Ricker model has two parameters and the right hand side of
Eqn 1 is

\begin{equation}
f(S(t)|r,K) = S(t) e^{r \left(1 - \frac{S(t)}{K} \right) } \label{ricker}
\end{equation}

The Allen model has three parameters

\begin{equation}
f(S(t)|r, K, X_c) = S(t) e^{r \left(1 - \frac{S(t)}{K}\right)\left(S(t) - X_c\right)} \label{allen}
\end{equation}

where $X_c$ denotes the location of the unstable steady state (i.e., the tipping point).

The Myers model also has three parameters:

\begin{equation}
f(S(t) | r, K, \theta)  = \frac{r S(t)^{\theta}}{1 + \frac{S(t)^\theta}{K}} \label{myers}
\end{equation}

where $\theta = 1$ corresponds to Beverton-Holt dynamics and $\theta$ > 2 leads to Allee effects and multiple stable states.

The Ricker model does not lead to multiple steady states.  The Allen
model resembles the Ricker dynamics with an added Allee effect parameter
[@Courchamp2008], below which the population cannot persist.  The Myers
model also has three parameters and contains an Allee threshold,
but unlike the Ricker model saturates  at high population size.
The multiplicative log-normal stochasticity in Eqn 1 introduces one
additional parameter $\sigma$ that must be estimated.

Because of our interest in management performance in the presence of
tipping points, all of our simulations are based on the Allen model.
The Allen model is thus the  state of nature  and is expected to provide
the best-case scenario. The Ricker model is a reasonable approximation
of these dynamics far from the Allee threshold (but lacks threshold
dynamics), while the Myers model shares the essential feature of a
threshold but differs in structure from the Allen model.  Throughout,
we refer to the "True" model when the underlying parameters _are known
without error_, and refer to the "Allen" model when these parameters
have been estimated from the sample data.

We consider a period of `r Tobs` in which data are obtained to estimate
the parameters or the GP. This is long enough that the estimates do not
depend on the particular realization, and longer times are not likely
to provide substantial improvement.  Each of the models is fit to the
same data (Figure 1).

We inferred posterior distributions for the parameters of each model by
Gibbs sampling (@Gelman2003 implemented in R [@RTeam] using `jags`,
[@R2jags]).  We choose uniform priors for all parameters of the
parametric models (See appendix Tables S1-S3; R code provided). We show
one-step-ahead predictions of these model fits in Figure 1.  We tested
each chain for Gelman-Rubin convergence and results were robust to
longer runs.  For each simulation we also applied several commonly used
model selection criteria (AIC, BIC, DIC, see @Burnham2002) to identify
the best fitting model.

Additionally, we compute the maximum likelihood estimate (MLE, as
we will refer to this model in the figures) of the parameters for
the (structurally correct) Allen model.  Comparing this to using the
posterior distribution of parameters inferred from MCMC for the same
model gives some indication of the importance of this uncertainty in
the dynamic programming.

### The Gaussian Process model

The core difference for our purpose between the estimated GP and
the estimated parametric models is that the estimated GP model is
defined explicitly in reference to the observed data.  As a result,
uncertainty arises in the GP model not only from uncertainty in the
parameters, but is also increases in regions farther from the observed
states, such as low population sizes in the example illustrated here.
The estimated parametric models, by contrast, are completely specified
by the parameters.

The use of GPs to characterize dynamical systems is relatively new
[@Kocijan2005], and was first introduced in the context ecological
modeling and fisheries management in @Munch2005.  GP models have
subsequently been used to test for the presence of Allee effects
[@Sugeno2013a], estimate the maximum reproductive rate [@Sugeno2013b],
determine temporal variation in food availability [@Sigourney2012], and
provide a basis for identifying model-misspecification [@Thorson2014].
An accessible and thorough introduction to the formulation and use of
GPs can be found in @Rasmussen2006.

A GP is a stochastic process for which any realization consisting of n
points follows a multivariate normal distribution of dimension $n$. To
characterize the GP we need a mean function and a covariance function. We
proceed as follows.

As before, we assume that the data $X(t)$ are observed with process
stochasticity around a mean function $g(X(t))$

\begin{equation}
X(t+1) = g(X(t)) + \varepsilon,
\end{equation}

where $\varepsilon$ are IID normal random variables with zero-mean
and variance $\sigma^2$. Note that we have chosen to assume additive
stochasticity.  While we could consider log-normal stochasticity as in
the parametric models, we make this choice to emphasize that the Gaussian
process approach need not have structurally correct stochasticity to
be effective.

In order to make predictions, we update the GP based on the observed set
of transitions.  To do so, we collect the time series of observed states
into a vector of "current" states, $\mathbf{X}_{\textrm{obs}} = \{X(1),
\dots, X(T-1)\}$ and a vector of "next" states $\mathbf{Y}_{\textrm{obs}}
= \{X(2),\dots,X(T)\}$ where $T$ is the time of the final observation.
Conditional on these observations, the predicted next state, $g(X_p)$,
for any given "current" state, $X_p$ follows a normal distribution
with mean $E$ and variance $C$ determined using the standard rules for
conditioning in multivariate normals, i.e.

\begin{equation}
E = K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbf{I}_n \right)^{-1} \mathbf{Y}_{\textrm{obs}}
\end{equation}

and

\begin{equation}
C = K(X_p, X_p) - K(X_p, \mathbf{X}_{\textrm{obs}}) \left(K(\mathbf{X}_{\textrm{obs}},\mathbf{X}_{\textrm{obs}}) - \sigma \mathbf{I} \right)^{-1} K(\mathbf{X}_{\textrm{obs}}, X_p)
\end{equation}

Here $\mathbf{I}_n$ is the $n$ by $n$ identity matrix (i.e. a matrix with
ones down the diagonal and zeros elsewhere) and $K$ is the ‘covariance
kernel.’  The covariance kernel controls how much influence one
observation has on another.  In the present application we use the
squared-exponential kernel which, when evaluated over a pair of vectors,
say $\mathbf{x}$ and $\mathbf{y}$,  generates a covariance matrix whose
$i,j$th element is given by

\begin{equation}
K_{i,j}(\mathbf{x}, \mathbf{y}) = \exp\left( \frac{ -(x_i - y_j)^2}{2 \ell^2} \right)
\end{equation}

so that $\ell$ gives the characteristic length-scale over which
correlation between two observations decays. See @Rasmussen2006 for
other choices of covariance kernels and their properties.  Note that this
simple formulation assumes a prior mean of zero. For the parameters we
use inverse Gamma priors on both the length-scale $\ell$ and $\sigma$,
thus for example

\begin{equation}
f(\ell; \alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} \ell^{-\alpha - 1}\exp\left(-\frac{\beta}{\ell}\right)
\end{equation}

For the prior on  $\ell$,
$\alpha =$ `r d.p[1]` and 
$\beta =$ `r d.p[2]`. 
The prior  on $\sigma$,
$\alpha =$ `r s2.p[1]`
and $\beta =$ `r s2.p[2]`.

We use a Metropolis-Hastings Markov Chain Monte Carlo (@Gelman2003) to
infer posterior distributions of the parameters of the GP (Figure
S4, code in appendix). Since the posterior distributions differ
substantially from the priors (Figure S4), most of the information 
in the posterior comes from the data rather than the
prior belief.

The method of Gaussian Process Dynamic Programming (GPDP)
--------------------------------------------------

We  derive the harvest policy from the estimated GP by inserting it into
a SDP algorithm. Given the GP posteriors, we construct the transition
matrix representing the probability of going to each state $X(t+1)$
given any current state $x(t)$ and any harvest $h(t)$ (See the function
`gp_transition_matrix()` in the provided R package). Given this transition
matrix, we use the same value iteration algorithm as in the parametric
case to determine the optimal policy. In doing so, the uncertainty in the
future state under the GP, $X(t+1)$, includes both process uncertainty
(based on the estimation of $\sigma$) and structural uncertainty of the
posterior collection of curves.

Results
=======

Parametric and GP models for population dynamics
------------------------------------------------

To ensure our results are robust to the choice of parameters, we will
consider 96 different scenarios.  To help better understand the process,
we first describe in detail the results of a single scenario.

```{r figure_1, dependson=c("assemble-models", "par-fns", "plot-options"), fig.cap="Points show the training data of stock-size over time.  Curves show the expectations over the posterior step-ahead predictions based on each of the estimated models. Observe that all models are fitting the data reasonably well.)", fig.width=8, fig.height=6}
step_ahead_posteriors <- function(x){
  gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
  df_post <- melt(lapply(sample(100, 30),  
  function(i){
    data.frame(time = 1:length(x), stock = x, 
                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
                True = step_ahead(x,f,p),  
                MLE = step_ahead(x,f,est$p), 
                Allen = step_ahead(x, allen_f, pardist[i,]), 
                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
  }), id=c("time", "stock"))
}
df_post <- step_ahead_posteriors(x)
library("dplyr")
df_post_means <- summarize(group_by(df_post, time,variable,stock), value = mean(value))
ggplot(df_post_means,aes(time, value)) +
  geom_point(aes(time, stock), size=2) + 
  geom_line(aes(time, value, col=variable, group=variable),lwd=1) + 
#  geom_line(aes(time, value, col=variable, group=interaction(L1,variable))) + 
  facet_wrap(~variable) + 
  scale_colour_manual(values=colorkey) +  
  theme(legend.position="none")

write.csv(df_post_means, "components/data/figure1.csv")
```

All of the models fit the observed data rather closely and with relatively
small uncertainty. In Figure 1, we show the posterior predictive curves.
The training data of stock sizes observed over time are points,
overlaid with the step-ahead predictions of each estimated model using
the parameters sampled from their posterior distributions. Compared to
the true model most estimates appear to over-fit, predicting patterns
that are actually due purely to stochasticity.  Model selection criteria
(Table 1) penalize more complex models and show a preference for the
simpler Ricker model over the  models with alternative stable states
(Allen and Myers). Supplement provides details on the model estimates.

```{r deviances, dependson=c("dic_calc"), include = FALSE, echo = FALSE}
allen_deviance  <- - posterior.mode(pardist[,'deviance'])
ricker_deviance <- - posterior.mode(ricker_pardist[,'deviance'])
myers_deviance  <- - posterior.mode(myers_pardist[,'deviance'])
true_deviance   <- 2*estf(c(p, sigma_g))
mle_deviance    <- 2*estf(c(est$p, est$sigma_g))
aictable <- data.frame(Allen = allen_deviance + 2*(1+length(bayes_pars)),  # +1 for noise parameter
                       Ricker = ricker_deviance + 2*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + 2*(1+length(myers_bayes_pars)), 
                       row.names = c("AIC"))
bictable <- data.frame(Allen = allen_deviance + log(length(x))*(1+length(bayes_pars)), 
                       Ricker = ricker_deviance + log(length(x))*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + log(length(x))*(1+length(myers_bayes_pars)), 
                       row.names = c("BIC"))
pandoc.table(rbind(dictable, aictable, bictable))
```

```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE}
pandoc.table(rbind(dictable, aictable, bictable), caption="Model selection scores for several common criteria (DIC: Deviance Information Criterion, AIC: Akaike Information Criterion, BIC: Bayesian Information Criterion) all select the wrong model. As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion. This highlights the danger of model choice when the selected model will be used outside of the observed range of the data.")
```

```{r figure_2, dependson=c("assemble-models", "plot-options"), fig.cap="The inferred Gaussian process compared to the true process and maximum-likelihood estimated (MLE) process.  We show the expected value for the function $f$ under each model.  Plots show replicates drawn from the posterior distributions in order to convey uncertainty of the estimates.  Note the MLE is a point estimate of parameters and so reflects no uncertainty in the distribution. The training data are also shown as black points.  The GP is conditioned on (0,0), shown as a pseudo-data point.", fig.width=8, fig.height=6}
# subset data first, don't care about seeing dynamics for stock >> K
x_grid_short <- x_grid[1:max(which(x_grid<12))]
gp_short <- gp_predict(gp, x_grid_short, burnin=1e4, thin=300)

# sample from subset data
models_posteriors <- 
  melt(lapply(sample(100, 50), 
              function(i){
    sample_gp <- mvrnorm(1, 
                            gp_short$Ef_posterior[,i],         
                            gp_short$Cf_posterior[[i]])
    data.frame(stock = x_grid_short, 
               GP = sample_gp,
               y = sample_gp,
               ymin = sample_gp - 2 * sqrt(gp_short$E_Vf), 
               ymax = sample_gp + 2 * sqrt(gp_short$E_Vf), 
               True = sapply(x_grid_short,f,0, p),  
               MLE = sapply(x_grid_short,f,0, est$p), 
               Allen = sapply(x_grid_short, allen_f, 0, pardist[i,]), 
               Ricker = sapply(x_grid_short, ricker_f, 0, ricker_pardist[i,]), 
               Myers = sapply(x_grid_short, myers_f, 0, myers_pardist[i,]))
             }), 
       id=c("stock", "y", "ymin", "ymax"))
ggplot(models_posteriors) + 
#    geom_ribbon(aes(x=stock, y=y, ymin=ymin, ymax=ymax, group=L1), 
#                  fill = "gray80", 
#                  data=subset(models_posteriors, variable == "GP")) + 
    geom_line(aes(stock, value, col = variable, 
                  group=interaction(L1,variable)), alpha=0.4) + 
    geom_point(data = obs, aes(x,y), size=2) + 
    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
    facet_wrap(~variable) + 
    scale_colour_manual(values=colorkey) +  
    theme(legend.position="none")

write.csv(models_posteriors, "components/data/figure2.csv")
write.csv(obs, "components/data/training_data.csv")
```

We show the  mean inferred population dynamics of each model relative
to the true model used to generate the data in Figure 2, predicting the
relationship between observed population size (x-axis) to the population
size after recruitment the following year.  In addition to the raw data,
the GP is conditioned on going through the point 0,0 without error.
All parametric models also make this assumption. Conditioning on (0,0)
is equivalent to making the assumption that the population is closed,
so that once it hits 0 it stays at 0, despite the lack of any data in
the observed sequence to justify this.  This assumption illustrates how
the GP can capture common-sense biology without having to assume more
explicit details about the dynamics at low population numbers that have
never been observed. If the population were not closed, one could repeat
the entire analysis without this assumption.  Unlike parametric models,
the GP corresponds to a distribution of curves, of which this plot only
shows the means. Uncertainty in the parameters of the GP (not shown)
further widens the band of possible population sizes.  In Figure S1
(see supplement), we show the performance of the models outside the
observed training data.

```{r figure_3, dependson = c("assemble-opt"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.  Policies are shown in terms of target escapement, $S(t)$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979] Several policies show a numerical jitter due to the discretization of states in the dynamic programming algorithm; doubling the number of grid points did not qualitatively change the results.", fig.width=8, fig.height=6}
policies <- melt(data.frame(stock=x_grid, sapply(OPT, function(x) x_grid[x])), id="stock")
names(policies) <- c("stock", "method", "value")

ggplot(policies, aes(stock, stock - value, color=method)) +
  geom_line() + 
  # stat_smooth(lwd=1.2, method="loess", degree=1, span=0.2, level=0, n=50) + 
  facet_wrap(~method) +
  xlab("stock size, x(t)") + 
  ylab("escapement, S(t)")  +
  scale_colour_manual(values=colorkey, guide=FALSE)

write.csv(policies, "components/data/figure3.csv")
```

Despite the similarities in model fits to the observed data, the policies
inferred under each model differ widely (Figure 3).  Policies are
shown in terms of target escapement, $S(t) = x_t - h$.  Under models
such as this a constant escapement policy is expected to be optimal
[@Reed1979], whereby population levels below a certain size $S$ are
unharvested, while above that size the harvest strategy aims to return
the population to $S$. Whenever a model predicts that the population
will not persist below a certain threshold, the optimal solution is to
harvest the entire population immediately, resulting in an escapement
$S=0$, as seen in the true (correct form, exact parameters) model, the
Allen model (correct form, estimated parameters) and the GP. Only the
structurally correct model (Allen model) and the GP produce policies
close to the true optimum policy.

```{r figure_4, dependson=c("sim"), fig.cap="In the management context, GPDP outperforms approaches based on parametric models. We show 100 replicate simulations of the stock dynamics (Eqn 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", fig.width=6, fig.height=6}
ggplot(sims_data) + 
  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=0.1) +
  scale_colour_manual(values=colorkey, guide = FALSE) +
  xlab("stock size, x(t)") +
  facet_wrap(~method) + 
  guides(legend.position="none", color=FALSE)

write.csv(sims_data, "components/data/figure4.csv")
```

In Figure 4, we show the consequences of managing 100 replicate
realizations of the simulated fishery under policies derived from each
model.  The structurally correct model under-harvests, leaving the stock
to vary around its unfished optimum.  The structurally incorrect Ricker
model over-harvests the population past the tipping point consistently,
resulting in the immediate crash of the stock and thus leads to minimal
long-term catch.

The results across replicate stochastic simulations are most easily
compared by using the relative differences in net present value
realized by each of the model (Figure 5).  Although not perfect, the
GPDP consistently realizes a value close to the optimal solution, and
avoids ever driving the system across the tipping point, which results
in the near-zero value cases in the parametric models.

```{r figure_5, dependson=c("profits"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates. For each data set, the three models are estimated as described above.", fig.width=6, fig.height=6, cache=FALSE}

ggplot(actual_over_optimal, aes(value)) + geom_histogram(aes(fill=variable)) + 
  facet_wrap(~variable, ncol=2)  + 
  guides(legend.position = "none") +
  xlab("Net present value") + 
  scale_fill_manual(values=colorkey, guide=FALSE) # density plots fail when delta fn

write.csv(actual_over_optimal, "components/data/figure5.csv")
```

Sensitivity Analysis
--------------------

These results are not sensitive to the modeling details of the simulation.
The GPDP estimate remains very close to the optimal solution (obtained by
knowing the true model) across changes to the training simulation, scale
of stochasticity, parameters or structure of the underlying model. In
the Supplement, we consider both a Latin hypercube approach and a more
focused investigation of the effects of the relative distance to the
Allee threshold and the variance of process stochasticity.

The GPDP is only weakly influenced by increasing stochasticity or
increasing Allee effects over much of the range (Figure S2). Larger
$\sigma$ or higher Allee levels make even the optimal solution without
any model or parameter uncertainty unable to harvest the population
effectively (e.g. the stochasticity is large enough to violate the
self-sustaining criterion of @Reed1979).

Discussion 
==========

Simple, mechanistically motivated models offer the
potential to increase our basic understanding of ecological processes
[@Cuddington2013; @Geritz2012]. But such models can be both inaccurate and
misleading when used in a decision making framework.
In this paper we tackled two aspects of uncertainty that are 
common to many ecological decision-making problems and fundamentally
challenging to existing approaches that largely rely on parametric
models:

1. We do not know the correct models for ecological systems.
2. We have limited data from which to estimate the model.

We have illustrated how the use of non-parametric methods provides
more reliable solutions in the sequential decision-making problem.

### Traditional model-choice approaches can be positively misleading

Our results illustrate that model-choice approaches can be absolutely
misleading -- by providing support to models that cannot capture tipping point
dynamics because they have fewer parameters and the data are far from the tipping point.  That is, when the data 
come from around the stable steady state,
all the parametric models are approximately linear and approximately
identical.  Thus, it is intuitive that all model selection methods choose the simplest model.
In a complex world, the result is suboptimal. But in a world that might
contain tipping points, the result could be disastrous.

Many managers both in fisheries and beyond face a similar situation: they
have not observed the population dynamics at all possible densities.  A
lack of comprehensive data at all population sizes, combined with the
inability to formulate accurate population models for low population sizes
in the absence of data, makes this situation the rule more than the exception.
Relying on parametric models and model choice processes that favor simplicity
ignores this basic reality. For a long time,  Carl Walters [e.g. @Walters1978] has argued that if we began
by fishing any newly exploited population down to very low levels and
then let it recover, we would be much better at estimating population
dynamics and thus predicting the optimal harvest levels.  While certainly true,
this presents a rather risky policy in the face of potential tipping points. 
The GPDP offers a risk-adverse alternative.

### GPDP population dynamics capture larger uncertainty in regions where the data are poor

Parametric models perform most poorly when we seek a management strategy
outside the range of the observed data. The GPDP, 
in contrast, leads to a predictive model that expresses a great deal of uncertainty
about the probable dynamics _outside_ the range of the observed data, while retaining very
good predictive accuracy _inside_ the range.  The management policy 
based on by the GPDP balances uncertainty outside the range of the observed data against
the immediate value of the harvest, and acts to stabilize the population 
dynamics in a region of state space in which the predictions are
reliably reflected by the data.

Such problems are 
ubiquitous across ecological decision-making
and conservation where the greatest concerns involve decisions that lead to population sizes that have never been observed and for which we do not know the response
-- whether this is the collapse of a fishery, the spread
of an invasive, or the loss of habitat.

### The role of the prior

Outside of the observed range of the data, the GP reverts to
the prior, and consequently the choice of the prior can also play a
significant role in determining the optimal policy.
In the examples shown here we have selected a prior that is both
relatively uninformative (due to the broad priors placed on its parameters
$\ell$ and $\sigma$ and simple (the choice of our covariance function, Eqns 12 and 13 ).
In practice, these
should be chosen to confer particular biological properties.
In principle, this may allow a manager to improve the performance of the
GPDP by adding detail as is justified.
For instance, it would be possible to use a linear or a Ricker-shaped
mean in the prior without making the much stronger assumption that the
Ricker is the structurally correct model [@Sugeno2013a]. One fruitful avenue of future research is  identifying criteria to ensure the prior and the reward function
are chosen appropriately for the problem at hand.

Acknowledgments
===============

This work was partially supported by NOAA-IAM grant to SM and Alec McCall and administered
through the Center for Stock Assessment Research, a partnership between
the University of California Santa Cruz and the Fisheries Ecology
Division, Southwest Fisheries Science Center, Santa Cruz, CA and by NSF
grant EF-0924195 to MM and NSF grant DBI-1306697 to CB.